### 2005-07-26 paper/main.tex
Start writing the paper.
Write the search procedure in pseudo-code as a test

\documentclass[a4paper,10pt]{article}
\usepackage[boxruled,vlined,portugues,figure,linesnumbered,longend]{algorithm2e}

\newcommand*{\Assign}{\ensuremath\longleftarrow}

%opening
\title{Approximate Text Indexing}
\author{L-P Coelho \and A. Oliveira }

\begin{document}

\maketitle

\begin{abstract}

\end{abstract}

\section{Introduction}
\section{The Structure}
\subsection{Search}

\input{search-algo}

\end{document}

### 2005-07-27 paper/main.tex
Fill in almost all the needed definitions. Start a BibTeX file Write the construction algorithm down (yeah, just three lines)

@@
 \newcommand*{\Assign}{\ensuremath\longleftarrow}
+\newcommand{\putstring}[1]{\textsl{#1}}
+\newtheorem{definition}{Definition}
 
-%opening
-\title{Approximate Text Indexing}
-\author{L-P Coelho \and A. Oliveira }
+\title{Dotted Suffix Trees\\A Structure for Approximate Text Indexing}
+\author{L. P. Coelho \and A. Oliveira}
 
@@
 \section{Introduction}
-\section{The Structure}
-\subsection{Search}
+
+\subsection{Related Work}
+\subsection{Structure of the Paper}
+
+In section~\ref{sec:structure} we present the structure and the way to use it for searching.
+In section~\ref{sec:build-1} we show the way to build a structure to handle one error only and analize this version of the algorithm.
+Then, we show, in section~\ref{sec:build-k} how to generalize this procedure for any number of errors and how the previous analisys also applies to any number of errors.
+
+\section{The Indexing Structure}
+
+\subsection{Intuition}
+Suffix Trees have become a well known structure since their introduction in 1973~\cite{weiner}. In Figure~\ref{fig:suffix-0} we show the suffix tree for the string \putstring{mississipi}.
+
+We add to each internal node in the structure a new link, which we call the \emph{error link}.
+
+\subsection{Formal definition}
+
+\begin{definition}[Character, string]
+Given a set $\Sigma$, we say that $s$ is a \emph{string over $\Sigma$} if $s$ is a possibly empty sequence of elements of $\Sigma$.
+The elements of the set $\Sigma$ are also called \emph{character}.
+The lenght of the string, denoted by $|s|$ is the number of elements it contains.
+We shall write $s_i$ for the $i$th element of the string.
+For the empty string, we write $\epsilon$.
+
+The set of all string is denoted by $\Sigma^*$ and $\Sigma^+=\Sigma^*-\{\epsilon\}$.
+\end{definition}
+
+For denoting character we shall use letters from the begginning of the roman alphabet ($a$, $b$, $c$,\ldots) and for strings, we shall use letters from the end of the alphabet ($w$, $z$, \ldots).
+
+\begin{definition}
+$wx$ will denote the string formed by all the character of $w$ followed by all the characters of $x$ (the usual concatenation operation).
+$aw$ ($wa$) shall the denote the strings by the caracter $a$ before (after) the characters of $w$.
+
+If $s = wxy$, then $w$ is a \emph{prefix} of $s$, $x$ is a \emph{substring} of $s$ (at position $s$) and $y$ is a \emph{suffix} of $s$ (at position $p$).
+\end{definition}
+
+\begin{definition}[$\Sigma^+$ tree]
+$T$ is a $\Sigma^+$ tree if $T$ is a rooted tree with edge labels from $\Sigma^+$. For each $a \in \Sigma$ and every node $n$ in $T$, there is at most an edge leaving $n$ whose label is $aw$.
+
+$\Sigma^+$ trees are also called a PATRICIA trees~\cite{pat}.
+
+Each node in a $\Sigma^+$ tree has a path leading to it which forms a string. If the node $n$ has the leading path $w$, we shall also refer to $n$ as $\overline{w}$.
+\end{definition}
+
+In what follows we will assume that there are two symbols (\$ and $.$) which are not part of $\Sigma$.
+
+\begin{definition}[Aproximate Match]
+We say that the string $s$ matches the string $t$ at position $p$ with $k$ errors if we can make $k$ substitutions in $s$ such that $s$ is a substring of $t$ at position $p$.
+\end{definition}
+
+\begin{definition}[Pattern]
+A pattern is a string over the extended alphabet $\Sigma\cup\{.\}$.
+We say that a pattern $s$ matches a string $t$ at position $p$, if for each dot in the pattern we can replace it by a character in the original alphabet to obtain $s'$ which is a substring of $t$ at position $p$.
+\end{definition}
+
+\begin{definition}[Suffix Tree]
+A \emph{suffix tree} for a string $S$ is a $\Sigma^+$ whose leaf nodes (those without children) have paths corresponding to suffixes of the string $S\$ $.
+\end{definition}
+
+\begin{definition}[Occurence Set]
+Given a node $\overline{w}$ in a suffix tree, we call it's \emph{occurence set} the set of indexes in the original string where the string $w$ occurs.
+\end{definition}
+
+\begin{definition}{Position Set}
+Given a node $\overline{w}$ in a suffix tree, it's \emph{position set} is the set formed by taking it's occurence set and adding the lenght of $w$ to each element.
+\end{definition}
+
+For example, in the node \overline{issi} in the suffix tree of Figure~\ref{fig:mississipi-0}, the occurence set is $\{5, 2\}$ and its position set is $\{6, 9\}$.
+In a sense, one can say that being at node \overline{issi} is being at positions 6 and~9 simultaneously.
+Note that for each node, the tree below it is the $\Sigma^+$ tree of the suffixes starting at the positions in its position set.
+
+In a leaf, the position set is a singleton, and we label the leaf by its element.
+
+\begin{definition}[Error Tree]
+For any node $w$ its error tree is the $\Sigma^+$ tree formed by taking it's position set, adding one to each element and forming the $\Sigma^+$ tree of the suffixes of the suffixes starting at those positions.
+If the position set includes the end of the string, that element is removed.
+
+The leafs are labeled by the position of the string in which their paths occur minus $|w| + 1$.
+\end{definition}
+
+For example, in the above mentioned node \overline{issi}, the error tree is formed by taking the strings starting at positions $\{7, 10\}$ (ie \putstring{sipi\$} and \putstring{i\$}) in a $\Sigma^+$ tree.
+
+\begin{definition}[1-error dotted Tree]
+We define a \emph{1-error dotted tree} as the tree which add to each node in a suffix tree, a new edge labeled by \putstring{.} which points to its error tree.
+The edge labeled \putstring{.} shall be called a \emph{dot link}.
+\end{definition}
+
+The one-error dotted tree for \putstring{mississipi} is shown in Figure~\ref{fig:mississipi-1}.
+
+The paths in the dotted tree are paths in the extended alphabet $\Sigma\cup\{.,\$\}$.
+The notions of \emph{occurence set}, \emph{position set} and \emph{error tree} are valid for all nodes in a dotted tree considering that a node path may contain dot elements which match any other character.
+
+\begin{definition}[k-error dotted tree]
+We define a k-error dotted tree as the tree obtained by adding error trees to each node in the (k-1)-error dotted tree.
+\end{definition}
+
+\subsection{Space Considerations}
+
+How much space does a k-error dotted tree take?
+Suffix Trees have a number of nodes proportional to the size of the string (which we will refer to as $n$).
+One-error dotted trees have more nodes.
+The worst case presented in Figure~\ref{fig:aaaa} is a string of the form \putstring{aa\ldots{}aa} which generates a one-dotted tree taking $O(n^2)$ nodes.
+For a larger number of errors $k$, it is easy to see that this leads to $O(n^{k+1})$ nodes.
+
+We will reason by induction, starting with the case of a one-error dotted tree to show that this bound can also be expressed as $O(nh^k)$ where $h$ is the average height of the tree (number of nodes between the root and the leaf).
+
+The number of nodes in a Patricia tree is $2n-1$~\cite{patricia} (as a special case, a suffix tree is the Patricia tree of all the suffixes in a string).
+The number of nodes in the error tree of a node is therefore bounded by double the number of leafs in the subtree at that node.
+
+We define $L(w)$ as the number of leafs in the subtree rooted at $n$.
+The total number of nodes for all the error trees in the dotted tree is therefore equal to $\Sum_{w \in \mathcal{T}}L(n)$ where $\mathcal{T}$ is the original suffix tree.
+It is a general property of trees that, for any tree $\mathcal{T}$, $\Sum_{w \in \mathcal{T}}L(w) = O(nh)$~\cite{whatever}.
+
+We also remark that at each node, the error tree cannot be deeper than the subtree rooted at that node. Therefore the average height of the tree is preserved. FIXME: show this
+
+For the (k+1)-error tree, we need to add error trees to each node in the k-error tree (which we assume inductively to have $O(nh^k)$ nodes).
+Each of these error trees will have $O(L(w))$ nodes.
+Again, given that the k-error tree has height $h$, we have $O(nh^{k+1})$ nodes as desired.
+
+Sofar, we have not really obtained much since, in the worst case, $h=O(n)$. However, the average expected case, is $h=O(\log n)$~\cite{devroye:note,szpankowski:unexpected}.
+
+\subsection{Searching}
+
+To search a string with at maximum k errors in a k-error dotted tree, we need to descend the tree.
+Whenever we find a node we follow both the dot link and the edge which matches the current position of the string we are searching.
+If we have a mismatch in the middle of an edge, we continue, decreasing the amount of errors permitted.
 
@@
 
+The search algorithm is presented in Figure~\ref{algo:search}.
+We have taken the liberty of using $s+j$ where $s$ is a string and $j$ an integer to mean the suffix of $s$ starting at $j$.
+Reporting of the leafs below a certain node can be done by a simple Depth First Search reporting any leafs found.
+If one is interested in knowing only whether the string occurs, then the lines which report the matches can be substituted by simple early exist returning true.
+
+\section{Constructing the Dotted Tree}
+
+\subsection{One-Error Dotted Tree}
+
+\bibliographystyle{alpha}
+\bibliography{biblio}
 \end{document}

### 2005-07-27 paper/main.tex
Almost complete description of the algorithm and almost complete analysis

@@
 
+\begin{definition}{Suffix Link}
+A suffix link in a suffix tree is a link from the node $\overline{aw}$ to the node $\overline{w}$. This node has the label $a$.
+\end{definition}
+
+In a suffix tree, it is possible to define a suffix link for each internal node.
+Both Ukkonen's~\cite{ukkonen} and McCreight's~\cite{mccreight} algorithms are linear time algorithms for constructing a suffix tree which includes suffix links.
+
 \begin{definition}[Occurence Set]
@@
 
-For example, in the node \overline{issi} in the suffix tree of Figure~\ref{fig:mississipi-0}, the occurence set is $\{5, 2\}$ and its position set is $\{6, 9\}$.
-In a sense, one can say that being at node \overline{issi} is being at positions 6 and~9 simultaneously.
+For example, in the node $\overline{issi}$ in the suffix tree of Figure~\ref{fig:mississipi-0}, the occurence set is $\{5, 2\}$ and its position set is $\{6, 9\}$.
+In a sense, one can say that being at node $\overline{issi}$ is being at positions 6 and~9 simultaneously.
 Note that for each node, the tree below it is the $\Sigma^+$ tree of the suffixes starting at the positions in its position set.
@@
 
-For example, in the above mentioned node \overline{issi}, the error tree is formed by taking the strings starting at positions $\{7, 10\}$ (ie \putstring{sipi\$} and \putstring{i\$}) in a $\Sigma^+$ tree.
+For example, in the above mentioned node $\overline{issi}$, the error tree is formed by taking the strings starting at positions $\{7, 10\}$ (ie \putstring{sipi\$} and \putstring{i\$}) in a $\Sigma^+$ tree.
 
@@
 We define $L(w)$ as the number of leafs in the subtree rooted at $n$.
-The total number of nodes for all the error trees in the dotted tree is therefore equal to $\Sum_{w \in \mathcal{T}}L(n)$ where $\mathcal{T}$ is the original suffix tree.
-It is a general property of trees that, for any tree $\mathcal{T}$, $\Sum_{w \in \mathcal{T}}L(w) = O(nh)$~\cite{whatever}.
+The total number of nodes for all the error trees in the dotted tree is therefore equal to $\sum_{w \in \mathcal{T}}L(n)$ where $\mathcal{T}$ is the original suffix tree.
+It is a general property of trees that, for any tree $\mathcal{T}$, $\sum_{w \in \mathcal{T}}L(w) = O(nh)$~\cite{whatever}.
 
@@
 \subsection{One-Error Dotted Tree}
+\subsubsection{The Algorithm}
+
+Constructing the tree is actually incredibly simple. We will start with a suffix tree which includes suffix links and show how to add error trees to it.
+
+First, we construct the error tree for the root. This tree is almost a copy of the entire tree, except for two properties:
+
+\begin{enumerate}
+\item It does not have the leaf labeled 1 in the original tree.
+\item All other leafs are labeled with value which is the original leaf's value minus one.
+\end{enumerate}
+
+It is easy to see why this is so.
+The leaf labeled 1 would have resulted in the string \putstring{.s\$} which does not occur in original string (it is one character too large).
+For any other leaf $\overline{w\$}$ which occured at position $p$ in the string, we have a new leaf $\overline{.w\$}$ which occurs at position $p-1$ in the string.
+
+For any other node $\overline{aw}$, the error tree is a filtered copy of the error tree at node $\overline{w}$.
+
+\begin{enumerate}
+\item The leaf labeled 1 in the original error tree is not included
+\item A leaf labeled $p$ is only included if $s_{p-1}$ is $a$ (the label of the suffix link)
+\item All leafs in the copy have a label which is the original label's value minus one
+\end{enumerate}
+
+Note that these conditions are very similiar to the conditions for the root and both can be implemented as shown in Figure~\ref{algo:copy-subtree}.
+
+\input{copy-subtree-algo}
+
+The only point to note is line~\ref{algoline:copy-subtree:merge}.
+This is motivated by noting that since we filter some leafs, without a merging procedure, it would be possible to have nodes which possess only one child.
+These need to be removed.
+
+It is very simple to remove these nodes.
+As discussed above (FIXME: not), suffix trees are normally implemented such that each node is actually just a pair of indices into the original string.
+As shown in Figure~\ref{fig:merge} to merge a child with its parent, we just need to replace both by a node whose starting index is the starting index of the original child minus the lenght of the parent edge.
+
+\input{adddotlink-algo}
+
+One can construct the error tree for any node if the tree for the node pointed to by the current node's suffix link has an error tree.
+This leads to a recursive definition as presented in Figure~\ref{algo:addDotLink}.
+
+\input{construct-algo}
+
+To add dot links at every node, we need to add the dot link at the root and then for every node call the procedure addDotLink.
+This is shown in Figure~\ref{algo:construct}.
+The \textit{foreach} loop can be implemented using Depth First Search on the tree.
+In fact, the procedure addDotLink only needs to be called for certain nodes (those who do not posess an incoming suffix link).
+However, there is no readily available way to access only these nodes and the procedure above does little extra work.
+
+\subsubsection{Time cost}
+
+If the number of nodes in the final tree is $N$ and the alphabet $\Sigma$, then the above algorithm runs in time $O(N|\Sigma|)$.
+
+The root's error tree is created in time proportional to the number of nodes it contains.
+Every other error tree is constructed by looking at an existing one and copying it.
+Each of these copies need time proportional to the original error tree.
+Since there are at maximum $|\Sigma|$ incoming suffix links to a node~\cite{}, each error tree is \emph{looked at} a maximum of $|\Sigma|$ times.
+The sum of all these operations is therefore bounded by $|\Sigma|N$.
+
+\subsection{Extension to any number of errors}
+
+The above algorithm can be used to construct trees with any number of errors by iterating it in the following way:
+
+To construct the (k+1)-error tree from the k-error tree, make a copy of the tree as above (adjusting leafs, filtering the leafs labeled one) and make this the root's error tree.
+Then, for every other node, remove the current error tree (it has a level too few).
+Finally, for every node except the root, construct its error tree as above (adjusting leafs, filtering the leafs labeled one or which do not correspond to positions in the suffix link's label).
+
+\input{copy-subtree-with-dot-link-algo}
+
+The code for the copy is shown in Figure~\ref{algo:copy-subtree-with-dot-link} which differs from Figure~\ref{algo:copy-subtree} in that the dot link is also followed while copying the tree.
+
+\input{construct-k-algo}
+
+The final algorithm for constructing k-error trees is shown in Figure~\ref{algo:construct-k}.
+
+We note that the method of removing the existing error trees can be improved in practice (the assymptotical bound remains the same) the following way: While making a copy of the tree for construction of the root's error tree, instead of following a dot link to make a copy, move the error tree to its destination and make the necessary adjustments inplace (this move will probably be just an adjustment of pointer depending on the exact implementation).
+Using this method only leafs labeled one need to be removed (as well as any internal node which becomes unnecessary).
 

### 2005-07-28 paper/abstract.tex
Add abstract & introduction Ignore generated files Fill in biblio

The problem we are addressing is the one of text indexing for approximate matching. We consider that we are given a text $\mathcal{T}$ which undergoes some preprocessing to generate an index. We can later query this index to identify the places where a string occurs up to a certain number of allowed substitutions $k$. We present a structure for this indexing which occupies space $O(n\log^kn)$ in the average case, indepedent of alphabet size, $n$ being the text size. This structure support searching in $O(m^{k+1})$ time, for patterns of size $m$, again independently of alphabet size. The construction of the structure has time bounded by $O(N|\Sigma|)$, where $N$ is the number of nodes in the index and $|\Sigma|$ the alphabet size.

### 2005-07-28 paper/introduction.tex
Add abstract & introduction Ignore generated files Fill in biblio

Since their introduction~\cite{weiner}, suffix trees have been one of the methods of choice for text indexing. They allow one to solve the following problem: given a text which has been preprocessed, and a pattern, efficiently find the places in the text where this pattern occurs.  However, for many real-life problems this statement is too restrictive. In these problems, one is interested in finding places in the text where an approximate form of the pattern occurs. Several algorithms have been proposed to solve this problem. \cite{navarro:guided} presents a survey of the problem and existing approaches to solving it at the time that work was written. Recently there have been some further work in the field, notably~\cite{maass:text} which also contains an updated overview of methods for solving the problem in addition to presenting one itself.

In this work we present an approach which solves this problem based on extending of suffix trees. The main advantage of this approach is that both the search and the index size are \emph{alphabet independent}, although the indexing time is not.

\subsection{Related Work}

Our work is related to efforts to index the text for approximate matching such as~\cite{maass:text,mosche}. The structure presented is superficially very similiar to the one presented in~\cite{inexact} as an \textit{inexact suffix tree} though that work is concerned with solving a different set of problems.

\subsection{Structure of the Paper}

In section~\ref{sec:structure} we present the structure and the way to use it for searching. In section~\ref{sec:build-1} we show the way to build a structure to handle one error only and analize this version of the algorithm. Then, we show, in section~\ref{sec:build-k} how to generalize this procedure for any number of errors and how the previous analisys also applies to any number of errors.

### 2005-07-28 paper/main.tex
Add abstract & introduction Ignore generated files Fill in biblio

@@
-\documentclass[a4paper,10pt]{article}
+\documentclass[letter,10pt]{article}
 \usepackage[boxruled,vlined,portugues,figure,linesnumbered,longend]{algorithm2e}
+\input{figures/transfig}
+\usepackage[]{graphicx}
 
@@
 \newtheorem{definition}{Definition}
+\newtheorem{lemma}{Lemma}
 
@@
 \begin{abstract}
-
+\input{abstract}
 \end{abstract}
@@
 \section{Introduction}
-
-\subsection{Related Work}
-\subsection{Structure of the Paper}
-
-In section~\ref{sec:structure} we present the structure and the way to use it for searching.
-In section~\ref{sec:build-1} we show the way to build a structure to handle one error only and analize this version of the algorithm.
-Then, we show, in section~\ref{sec:build-k} how to generalize this procedure for any number of errors and how the previous analisys also applies to any number of errors.
+\input{introduction}
 
@@
 \subsection{Intuition}
-Suffix Trees have become a well known structure since their introduction in 1973~\cite{weiner}. In Figure~\ref{fig:suffix-0} we show the suffix tree for the string \putstring{mississipi}.
 
-We add to each internal node in the structure a new link, which we call the \emph{error link}.
+\begin{figure}
+\includegraphics{figures/mississipi-0}
+\caption{Suffix Tree for the string \putstring{mississipi}}%
+\label{fig:mississipi-0}
+\end{figure}
+
+Suffix Trees have become a well known structure since their introduction in 1973~\cite{weiner}. In Figure~\ref{fig:mississipi-0} we show the suffix tree for the string \putstring{mississipi}.
+
+Being at a node in the suffix tree corresponds to being in several points in the string at the same time.
+We now add to each node an extra edge which corresponds to moving one character forward in all these points.
 
 \subsection{Formal definition}
+
+We assume that the reader is familiar with the basic concepts of suffix trees and strings and present the definitions below mainly to introduce the particular notation and terminology used in this paper.
 
@@
 
-\begin{definition}[$\Sigma^+$ tree]
-$T$ is a $\Sigma^+$ tree if $T$ is a rooted tree with edge labels from $\Sigma^+$. For each $a \in \Sigma$ and every node $n$ in $T$, there is at most an edge leaving $n$ whose label is $aw$.
-
-$\Sigma^+$ trees are also called a PATRICIA trees~\cite{pat}.
+\begin{definition}[Patricia tree]
+$T$ is a Patricia tree if $T$ is a rooted tree with edge labels from $\Sigma^+$. For each $a \in \Sigma$ and every node $n$ in $T$, there is at most an edge leaving $n$ whose label is $aw$.
 
@@
 
+The name Patricia is sometimes spelt PATRICIA, since it was formed from the initial of Practical Algorithm to Retrieve Information Coded in Alphanumeric~\cite{morrison:patricia}.
+
 In what follows we will assume that there are two symbols (\$ and $.$) which are not part of $\Sigma$.
+
+In this paper, we will be interested in substitution of Hamming distance:
 
@@
 
-\begin{definition}{Suffix Link}
+\begin{definition}[Suffix Link]
 A suffix link in a suffix tree is a link from the node $\overline{aw}$ to the node $\overline{w}$. This node has the label $a$.
@@
 
-\begin{definition}{Position Set}
+\begin{definition}[Position Set]
 Given a node $\overline{w}$ in a suffix tree, it's \emph{position set} is the set formed by taking it's occurence set and adding the lenght of $w$ to each element.
@@
 
+\begin{figure}
+\includegraphics{figures/aaaa}
+\caption{$a^n$: worst case space consumption}%
+\label{fig:aaaa}
+\end{figure}
+
 How much space does a k-error dotted tree take?
@@
 One-error dotted trees have more nodes.
-The worst case presented in Figure~\ref{fig:aaaa} is a string of the form \putstring{aa\ldots{}aa} which generates a one-dotted tree taking $O(n^2)$ nodes.
+The worst case presented in Figure~\ref{fig:aaaa} is a string of the form $a^n = \underbrace{aaa\cdots}_{n \mathit{times }a}$ which generates a one-dotted tree taking $O(n^2)$ nodes.
 For a larger number of errors $k$, it is easy to see that this leads to $O(n^{k+1})$ nodes.
+
+\begin{definition}[weight under a node, weight of a tree]
+The weight under a node $\mathit{Weight}(n)$ is the number of leafs in the subtree rooted at that node.
+
+The total weight of a tree is the sum of the weights of all the nodes.
+\end{definition}
+
+\begin{definition}[height of a node, average height]
+The height of a node $\mathit{Height}(n)$ is the number of nodes in the path leading to that node from the root. The average height of the tree $\hat{h}$, is the average height of its nodes.
+
+\[ \hat{h} = \frac{1}{n} \sum_{x \mathit{node of \mathcal{T}}} \mathit{Height}(x), n \mathrm{being the number of nodes} \]
+\end{definition}
+
+\begin{lemma}
+Given a tree (any tree, not just suffix or similiar tree), the sum of the weights of all the nodes in the tree is equal to the average height of the node times the number of nodes.
+\end{lemma}
+
+Consider the tree being built node by node, in any order, as long as the property that it remains a tree during the process is maintained.
+Each node added, adds a certain amount of weight to the tree.
+In particular it adds one for each node that sits above it in the tree.
+This is the same as saying that each node adds its height (which does not change).
+The weight of the final tree is the sum of all these values.
+
+\[ \mathit{Weight}(\mathcal{T}) = \sum_{x \in \mathcal{T}}\mathit{Height}(x) = n\frac{1}{n}\sum_{x \in \mathcal{T}}\mathit{Height}(x) = n\hat{h} \]
+
+FIXME
 
@@
 
+\section{Experimental Results}
+
+\section{Future Work, Open Problems}
+
+In the example for the string \putstring{mississipi}, presented in Figure~\ref{fig:mississipi-1}, one can see that the tree below \putstring{s.i} and \putstring{ssi} are exactly the same.
+Whether such occurences are the basis for a significant space saving and what algorithms mightexploit them is an open question.
+Along the same lines, the structure's definition might be extended to structures such as the suffix-DAG presented in~\cite[7.7]{gusfield:algorithms}.
+An algorithm to efficiently construct it over these suffix-DAGs is another open problem.
+
 \bibliographystyle{alpha}

### 2005-07-28 src/suffix_tree.cpp
Greatly simplify the code for searching Fix stupid typo in copy_subtree()

@@ -602,23 +602,16 @@ DBL_WORD find_string( SUFFIX_TREE* tree, NODE* node, const char* W, DBL_WORD P, 
 	if ( !node ) return ST_ERROR;
 
 	DBL_WORD j = 0;
-	unsigned k = node->edge_label_start;
-	const unsigned label_end = get_node_label_end( tree, node );
+	const unsigned start = node->edge_label_start;
+	const unsigned len = std::min( P, get_node_label_length( tree, node ) );
 	while ( d >= 0 ) {
-		while ( j < P &&
-				k <= label_end &&
-				W[ j ] == tree->tree_string[ k ] ) {
+		while ( j < len && W[ j ] == tree->tree_string[ start + j ] ) {
 			++j;
-			++k;
 		}
-		// Check what caused the break:
-		if ( j == P ) return node->path_position;
-
-		if ( k > label_end ) {
+		if ( j == len ) {
 			return find_string( tree, node, W + j, P - j, d );
 		}
 		--d;
-		++k;
 		++j;
 	}
 	return ST_ERROR;
@@ -1248,7 +1241,7 @@ NODE* copy_subtree( SUFFIX_TREE* tree, NODE* start, copy_filter_type ftype, char
 			res->right_sibling = copy_subtree( tree, start->right_sibling, ftype, x );
 			if ( res->right_sibling ) res->right_sibling->left_sibling = res;
 		}
-		res->dot_link = copy_subtree( tree, res->dot_link, ftype, x );
+		res->dot_link = copy_subtree( tree, start->dot_link, ftype, x );
 	}
 	return res;
 }

### 2005-07-29 paper/construction.tex
Break into little tex files

--- 2005-07-27 paper/main.tex
@@
-\documentclass[a4paper,10pt]{article}
-\usepackage[boxruled,vlined,portugues,figure,linesnumbered,longend]{algorithm2e}
-
-\newcommand*{\Assign}{\ensuremath\longleftarrow}
-\newcommand{\putstring}[1]{\textsl{#1}}
-\newtheorem{definition}{Definition}
-
-\title{Dotted Suffix Trees\\A Structure for Approximate Text Indexing}
-\author{L. P. Coelho \and A. Oliveira}
-
-\begin{document}
-
-\maketitle
-
-\begin{abstract}
-
-\end{abstract}
-
-\section{Introduction}
-
-\subsection{Related Work}
-\subsection{Structure of the Paper}
-
-In section~\ref{sec:structure} we present the structure and the way to use it for searching.
-In section~\ref{sec:build-1} we show the way to build a structure to handle one error only and analize this version of the algorithm.
-Then, we show, in section~\ref{sec:build-k} how to generalize this procedure for any number of errors and how the previous analisys also applies to any number of errors.
-
-\section{The Indexing Structure}
-
-\subsection{Intuition}
-Suffix Trees have become a well known structure since their introduction in 1973~\cite{weiner}. In Figure~\ref{fig:suffix-0} we show the suffix tree for the string \putstring{mississipi}.
-
-We add to each internal node in the structure a new link, which we call the \emph{error link}.
-
-\subsection{Formal definition}
-
-\begin{definition}[Character, string]
-Given a set $\Sigma$, we say that $s$ is a \emph{string over $\Sigma$} if $s$ is a possibly empty sequence of elements of $\Sigma$.
-The elements of the set $\Sigma$ are also called \emph{character}.
-The lenght of the string, denoted by $|s|$ is the number of elements it contains.
-We shall write $s_i$ for the $i$th element of the string.
-For the empty string, we write $\epsilon$.
-
-The set of all string is denoted by $\Sigma^*$ and $\Sigma^+=\Sigma^*-\{\epsilon\}$.
-\end{definition}
-
-For denoting character we shall use letters from the begginning of the roman alphabet ($a$, $b$, $c$,\ldots) and for strings, we shall use letters from the end of the alphabet ($w$, $z$, \ldots).
-
-\begin{definition}
-$wx$ will denote the string formed by all the character of $w$ followed by all the characters of $x$ (the usual concatenation operation).
-$aw$ ($wa$) shall the denote the strings by the caracter $a$ before (after) the characters of $w$.
-
-If $s = wxy$, then $w$ is a \emph{prefix} of $s$, $x$ is a \emph{substring} of $s$ (at position $s$) and $y$ is a \emph{suffix} of $s$ (at position $p$).
-\end{definition}
-
-\begin{definition}[$\Sigma^+$ tree]
-$T$ is a $\Sigma^+$ tree if $T$ is a rooted tree with edge labels from $\Sigma^+$. For each $a \in \Sigma$ and every node $n$ in $T$, there is at most an edge leaving $n$ whose label is $aw$.
-
-$\Sigma^+$ trees are also called a PATRICIA trees~\cite{pat}.
-
-Each node in a $\Sigma^+$ tree has a path leading to it which forms a string. If the node $n$ has the leading path $w$, we shall also refer to $n$ as $\overline{w}$.
-\end{definition}
-
-In what follows we will assume that there are two symbols (\$ and $.$) which are not part of $\Sigma$.
-
-\begin{definition}[Aproximate Match]
-We say that the string $s$ matches the string $t$ at position $p$ with $k$ errors if we can make $k$ substitutions in $s$ such that $s$ is a substring of $t$ at position $p$.
-\end{definition}
-
-\begin{definition}[Pattern]
-A pattern is a string over the extended alphabet $\Sigma\cup\{.\}$.
-We say that a pattern $s$ matches a string $t$ at position $p$, if for each dot in the pattern we can replace it by a character in the original alphabet to obtain $s'$ which is a substring of $t$ at position $p$.
-\end{definition}
-
-\begin{definition}[Suffix Tree]
-A \emph{suffix tree} for a string $S$ is a $\Sigma^+$ whose leaf nodes (those without children) have paths corresponding to suffixes of the string $S\$ $.
-\end{definition}
-
-\begin{definition}{Suffix Link}
-A suffix link in a suffix tree is a link from the node $\overline{aw}$ to the node $\overline{w}$. This node has the label $a$.
-\end{definition}
-
-In a suffix tree, it is possible to define a suffix link for each internal node.
-Both Ukkonen's~\cite{ukkonen} and McCreight's~\cite{mccreight} algorithms are linear time algorithms for constructing a suffix tree which includes suffix links.
-
-\begin{definition}[Occurence Set]
-Given a node $\overline{w}$ in a suffix tree, we call it's \emph{occurence set} the set of indexes in the original string where the string $w$ occurs.
-\end{definition}
-
-\begin{definition}{Position Set}
-Given a node $\overline{w}$ in a suffix tree, it's \emph{position set} is the set formed by taking it's occurence set and adding the lenght of $w$ to each element.
-\end{definition}
-
-For example, in the node $\overline{issi}$ in the suffix tree of Figure~\ref{fig:mississipi-0}, the occurence set is $\{5, 2\}$ and its position set is $\{6, 9\}$.
-In a sense, one can say that being at node $\overline{issi}$ is being at positions 6 and~9 simultaneously.
-Note that for each node, the tree below it is the $\Sigma^+$ tree of the suffixes starting at the positions in its position set.
-
-In a leaf, the position set is a singleton, and we label the leaf by its element.
-
-\begin{definition}[Error Tree]
-For any node $w$ its error tree is the $\Sigma^+$ tree formed by taking it's position set, adding one to each element and forming the $\Sigma^+$ tree of the suffixes of the suffixes starting at those positions.
-If the position set includes the end of the string, that element is removed.
-
-The leafs are labeled by the position of the string in which their paths occur minus $|w| + 1$.
-\end{definition}
-
-For example, in the above mentioned node $\overline{issi}$, the error tree is formed by taking the strings starting at positions $\{7, 10\}$ (ie \putstring{sipi\$} and \putstring{i\$}) in a $\Sigma^+$ tree.
-
-\begin{definition}[1-error dotted Tree]
-We define a \emph{1-error dotted tree} as the tree which add to each node in a suffix tree, a new edge labeled by \putstring{.} which points to its error tree.
-The edge labeled \putstring{.} shall be called a \emph{dot link}.
-\end{definition}
-
-The one-error dotted tree for \putstring{mississipi} is shown in Figure~\ref{fig:mississipi-1}.
-
-The paths in the dotted tree are paths in the extended alphabet $\Sigma\cup\{.,\$\}$.
-The notions of \emph{occurence set}, \emph{position set} and \emph{error tree} are valid for all nodes in a dotted tree considering that a node path may contain dot elements which match any other character.
-
-\begin{definition}[k-error dotted tree]
-We define a k-error dotted tree as the tree obtained by adding error trees to each node in the (k-1)-error dotted tree.
-\end{definition}
-
-\subsection{Space Considerations}
-
-How much space does a k-error dotted tree take?
-Suffix Trees have a number of nodes proportional to the size of the string (which we will refer to as $n$).
-One-error dotted trees have more nodes.
-The worst case presented in Figure~\ref{fig:aaaa} is a string of the form \putstring{aa\ldots{}aa} which generates a one-dotted tree taking $O(n^2)$ nodes.
-For a larger number of errors $k$, it is easy to see that this leads to $O(n^{k+1})$ nodes.
-
-We will reason by induction, starting with the case of a one-error dotted tree to show that this bound can also be expressed as $O(nh^k)$ where $h$ is the average height of the tree (number of nodes between the root and the leaf).
-
-The number of nodes in a Patricia tree is $2n-1$~\cite{patricia} (as a special case, a suffix tree is the Patricia tree of all the suffixes in a string).
-The number of nodes in the error tree of a node is therefore bounded by double the number of leafs in the subtree at that node.
-
-We define $L(w)$ as the number of leafs in the subtree rooted at $n$.
-The total number of nodes for all the error trees in the dotted tree is therefore equal to $\sum_{w \in \mathcal{T}}L(n)$ where $\mathcal{T}$ is the original suffix tree.
-It is a general property of trees that, for any tree $\mathcal{T}$, $\sum_{w \in \mathcal{T}}L(w) = O(nh)$~\cite{whatever}.
-
-We also remark that at each node, the error tree cannot be deeper than the subtree rooted at that node. Therefore the average height of the tree is preserved. FIXME: show this
-
-For the (k+1)-error tree, we need to add error trees to each node in the k-error tree (which we assume inductively to have $O(nh^k)$ nodes).
-Each of these error trees will have $O(L(w))$ nodes.
-Again, given that the k-error tree has height $h$, we have $O(nh^{k+1})$ nodes as desired.
-
-Sofar, we have not really obtained much since, in the worst case, $h=O(n)$. However, the average expected case, is $h=O(\log n)$~\cite{devroye:note,szpankowski:unexpected}.
-
-\subsection{Searching}
-
-To search a string with at maximum k errors in a k-error dotted tree, we need to descend the tree.
-Whenever we find a node we follow both the dot link and the edge which matches the current position of the string we are searching.
-If we have a mismatch in the middle of an edge, we continue, decreasing the amount of errors permitted.
-
-\input{search-algo}
-
-The search algorithm is presented in Figure~\ref{algo:search}.
-We have taken the liberty of using $s+j$ where $s$ is a string and $j$ an integer to mean the suffix of $s$ starting at $j$.
-Reporting of the leafs below a certain node can be done by a simple Depth First Search reporting any leafs found.
-If one is interested in knowing only whether the string occurs, then the lines which report the matches can be substituted by simple early exist returning true.
-
-\section{Constructing the Dotted Tree}
-
 \subsection{One-Error Dotted Tree}
@@
 
-Constructing the tree is actually incredibly simple. We will start with a suffix tree which includes suffix links and show how to add error trees to it.
+Constructing the tree is actually very simple. We will start with a suffix tree which includes suffix links and show how to add error trees to it.
 
@@
 
-For any other node $\overline{aw}$, the error tree is a filtered copy of the error tree at node $\overline{w}$.
+For any other node $\overline{aw}$, the error tree is a filtered copy of the error tree at node $\overline{w}$ (the node pointed to by node $\overline{aw}$'s suffix link.
 
@@
 It is very simple to remove these nodes.
-As discussed above (FIXME: not), suffix trees are normally implemented such that each node is actually just a pair of indices into the original string.
-As shown in Figure~\ref{fig:merge} to merge a child with its parent, we just need to replace both by a node whose starting index is the starting index of the original child minus the lenght of the parent edge.
+As discussed above, suffix trees are normally implemented such that each node is actually just a pair of indices into the original string as shown in Figure~\ref{fig:mississipi-nodes}.
+To merge a child with its parent, we just need to replace both by a node whose starting index is the starting index of the original child minus the lenght of the parent edge.
+This process is illustrated in Figure~\ref{fig:merge}
 
@@
 Using this method only leafs labeled one need to be removed (as well as any internal node which becomes unnecessary).
-
-\bibliographystyle{alpha}
-\bibliography{biblio}
-\end{document}

### 2005-07-29 paper/experimental.tex
Break into little tex files

\subsection{Implementation and Data Sets}

The code was implemented in the C~Language, starting from the suffix tree implementation by FIXME.

Three data sets were used:

\begin{enumerate}
\item ``Democracy in America'' by Alexis de Tocqueville as natural language data ser
\item The DNA of S. Cervesiae
\item Randomly generated text
\end{enumerate}

### 2005-07-29 paper/main.tex
Break into little tex files

--- 2005-07-27 paper/main.tex
@@
-\documentclass[a4paper,10pt]{article}
+\documentclass[letter,10pt]{article}
 \usepackage[boxruled,vlined,portugues,figure,linesnumbered,longend]{algorithm2e}
+\input{figures/transfig}
+\usepackage[]{graphicx}
+\usepackage{QED}
+\usepackage{amssymb}
 
@@
 \section{Introduction}
-
-\subsection{Related Work}
-\subsection{Structure of the Paper}
-
-In section~\ref{sec:structure} we present the structure and the way to use it for searching.
-In section~\ref{sec:build-1} we show the way to build a structure to handle one error only and analize this version of the algorithm.
-Then, we show, in section~\ref{sec:build-k} how to generalize this procedure for any number of errors and how the previous analisys also applies to any number of errors.
+\input{introduction}
 
 \section{The Indexing Structure}
+\input{structure}
+\section{Space Considerations}
+\input{space}
+\section{Searching}
+\input{search}
+\section{Constructing the Dotted Tree}
+\input{construction}
+\section{Experimental Results}
+\input{experimental}
+\section{Future Work, Open Problems}
 
-\subsection{Intuition}
-Suffix Trees have become a well known structure since their introduction in 1973~\cite{weiner}. In Figure~\ref{fig:suffix-0} we show the suffix tree for the string \putstring{mississipi}.
-
-We add to each internal node in the structure a new link, which we call the \emph{error link}.
-
-\subsection{Formal definition}
-
-\begin{definition}[Character, string]
-Given a set $\Sigma$, we say that $s$ is a \emph{string over $\Sigma$} if $s$ is a possibly empty sequence of elements of $\Sigma$.
-The elements of the set $\Sigma$ are also called \emph{character}.
-The lenght of the string, denoted by $|s|$ is the number of elements it contains.
-We shall write $s_i$ for the $i$th element of the string.
-For the empty string, we write $\epsilon$.
-
-The set of all string is denoted by $\Sigma^*$ and $\Sigma^+=\Sigma^*-\{\epsilon\}$.
-\end{definition}
-
-For denoting character we shall use letters from the begginning of the roman alphabet ($a$, $b$, $c$,\ldots) and for strings, we shall use letters from the end of the alphabet ($w$, $z$, \ldots).
-
-\begin{definition}
-$wx$ will denote the string formed by all the character of $w$ followed by all the characters of $x$ (the usual concatenation operation).
-$aw$ ($wa$) shall the denote the strings by the caracter $a$ before (after) the characters of $w$.
-
-If $s = wxy$, then $w$ is a \emph{prefix} of $s$, $x$ is a \emph{substring} of $s$ (at position $s$) and $y$ is a \emph{suffix} of $s$ (at position $p$).
-\end{definition}
-
-\begin{definition}[$\Sigma^+$ tree]
-$T$ is a $\Sigma^+$ tree if $T$ is a rooted tree with edge labels from $\Sigma^+$. For each $a \in \Sigma$ and every node $n$ in $T$, there is at most an edge leaving $n$ whose label is $aw$.
-
-$\Sigma^+$ trees are also called a PATRICIA trees~\cite{pat}.
-
-Each node in a $\Sigma^+$ tree has a path leading to it which forms a string. If the node $n$ has the leading path $w$, we shall also refer to $n$ as $\overline{w}$.
-\end{definition}
-
-In what follows we will assume that there are two symbols (\$ and $.$) which are not part of $\Sigma$.
-
-\begin{definition}[Aproximate Match]
-We say that the string $s$ matches the string $t$ at position $p$ with $k$ errors if we can make $k$ substitutions in $s$ such that $s$ is a substring of $t$ at position $p$.
-\end{definition}
-
-\begin{definition}[Pattern]
-A pattern is a string over the extended alphabet $\Sigma\cup\{.\}$.
-We say that a pattern $s$ matches a string $t$ at position $p$, if for each dot in the pattern we can replace it by a character in the original alphabet to obtain $s'$ which is a substring of $t$ at position $p$.
-\end{definition}
-
-\begin{definition}[Suffix Tree]
-A \emph{suffix tree} for a string $S$ is a $\Sigma^+$ whose leaf nodes (those without children) have paths corresponding to suffixes of the string $S\$ $.
-\end{definition}
-
-\begin{definition}[Occurence Set]
-Given a node $\overline{w}$ in a suffix tree, we call it's \emph{occurence set} the set of indexes in the original string where the string $w$ occurs.
-\end{definition}
-
-\begin{definition}{Position Set}
-Given a node $\overline{w}$ in a suffix tree, it's \emph{position set} is the set formed by taking it's occurence set and adding the lenght of $w$ to each element.
-\end{definition}
-
-For example, in the node \overline{issi} in the suffix tree of Figure~\ref{fig:mississipi-0}, the occurence set is $\{5, 2\}$ and its position set is $\{6, 9\}$.
-In a sense, one can say that being at node \overline{issi} is being at positions 6 and~9 simultaneously.
-Note that for each node, the tree below it is the $\Sigma^+$ tree of the suffixes starting at the positions in its position set.
-
-In a leaf, the position set is a singleton, and we label the leaf by its element.
-
-\begin{definition}[Error Tree]
-For any node $w$ its error tree is the $\Sigma^+$ tree formed by taking it's position set, adding one to each element and forming the $\Sigma^+$ tree of the suffixes of the suffixes starting at those positions.
-If the position set includes the end of the string, that element is removed.
-
-The leafs are labeled by the position of the string in which their paths occur minus $|w| + 1$.
-\end{definition}
-
-For example, in the above mentioned node \overline{issi}, the error tree is formed by taking the strings starting at positions $\{7, 10\}$ (ie \putstring{sipi\$} and \putstring{i\$}) in a $\Sigma^+$ tree.
-
-\begin{definition}[1-error dotted Tree]
-We define a \emph{1-error dotted tree} as the tree which add to each node in a suffix tree, a new edge labeled by \putstring{.} which points to its error tree.
-The edge labeled \putstring{.} shall be called a \emph{dot link}.
-\end{definition}
-
-The one-error dotted tree for \putstring{mississipi} is shown in Figure~\ref{fig:mississipi-1}.
-
-The paths in the dotted tree are paths in the extended alphabet $\Sigma\cup\{.,\$\}$.
-The notions of \emph{occurence set}, \emph{position set} and \emph{error tree} are valid for all nodes in a dotted tree considering that a node path may contain dot elements which match any other character.
-
-\begin{definition}[k-error dotted tree]
-We define a k-error dotted tree as the tree obtained by adding error trees to each node in the (k-1)-error dotted tree.
-\end{definition}
-
-\subsection{Space Considerations}
-
-How much space does a k-error dotted tree take?
-Suffix Trees have a number of nodes proportional to the size of the string (which we will refer to as $n$).
-One-error dotted trees have more nodes.
-The worst case presented in Figure~\ref{fig:aaaa} is a string of the form \putstring{aa\ldots{}aa} which generates a one-dotted tree taking $O(n^2)$ nodes.
-For a larger number of errors $k$, it is easy to see that this leads to $O(n^{k+1})$ nodes.
-
-We will reason by induction, starting with the case of a one-error dotted tree to show that this bound can also be expressed as $O(nh^k)$ where $h$ is the average height of the tree (number of nodes between the root and the leaf).
-
-The number of nodes in a Patricia tree is $2n-1$~\cite{patricia} (as a special case, a suffix tree is the Patricia tree of all the suffixes in a string).
-The number of nodes in the error tree of a node is therefore bounded by double the number of leafs in the subtree at that node.
-
-We define $L(w)$ as the number of leafs in the subtree rooted at $n$.
-The total number of nodes for all the error trees in the dotted tree is therefore equal to $\Sum_{w \in \mathcal{T}}L(n)$ where $\mathcal{T}$ is the original suffix tree.
-It is a general property of trees that, for any tree $\mathcal{T}$, $\Sum_{w \in \mathcal{T}}L(w) = O(nh)$~\cite{whatever}.
-
-We also remark that at each node, the error tree cannot be deeper than the subtree rooted at that node. Therefore the average height of the tree is preserved. FIXME: show this
-
-For the (k+1)-error tree, we need to add error trees to each node in the k-error tree (which we assume inductively to have $O(nh^k)$ nodes).
-Each of these error trees will have $O(L(w))$ nodes.
-Again, given that the k-error tree has height $h$, we have $O(nh^{k+1})$ nodes as desired.
-
-Sofar, we have not really obtained much since, in the worst case, $h=O(n)$. However, the average expected case, is $h=O(\log n)$~\cite{devroye:note,szpankowski:unexpected}.
-
-\subsection{Searching}
-
-To search a string with at maximum k errors in a k-error dotted tree, we need to descend the tree.
-Whenever we find a node we follow both the dot link and the edge which matches the current position of the string we are searching.
-If we have a mismatch in the middle of an edge, we continue, decreasing the amount of errors permitted.
-
-\input{search-algo}
-
-The search algorithm is presented in Figure~\ref{algo:search}.
-We have taken the liberty of using $s+j$ where $s$ is a string and $j$ an integer to mean the suffix of $s$ starting at $j$.
-Reporting of the leafs below a certain node can be done by a simple Depth First Search reporting any leafs found.
-If one is interested in knowing only whether the string occurs, then the lines which report the matches can be substituted by simple early exist returning true.
-
-\section{Constructing the Dotted Tree}
-
-\subsection{One-Error Dotted Tree}
+In the example for the string \putstring{mississipi}, presented in Figure~\ref{fig:mississipi-1}, one can see that the tree below \putstring{s.i} and \putstring{ssi} are exactly the same.
+Whether such occurences are the basis for a significant space saving and what algorithms mightexploit them is an open question.
+Along the same lines, the structure's definition might be extended to structures such as the suffix-DAG presented in~\cite[7.7]{gusfield:algorithms}.
+An algorithm to efficiently construct it over these suffix-DAGs is another open problem.
 

### 2005-07-29 paper/search.tex
Break into little tex files

To search a string with at maximum k errors in a k-error dotted tree, we need to descend the tree. Whenever we find a node we follow both the dot link and the edge which matches the current position of the string we are searching. If we have a mismatch in the middle of an edge, we continue, decreasing the amount of errors permitted.

\begin{figure}
\includegraphics{figures/mississipi-nodes}
\caption{Suffix Tree for the string \putstring{mississipi}}%
\label{fig:mississipi-nodes}
\end{figure}

To understand the pseudo-code, it is important to understand that suffix trees are normally implemented as shown in Figure~\ref{fig:mississipi-nodes} where each node contains the start and end indices of its incoming edge. When presenting the algorithm, we take the liberty of using $s+j$ where $s$ is a string and $j$ an integer to mean the suffix of $s$ starting at position $j$.

\input{search-algo}

The search algorithm is presented in Figure~\ref{algo:search}. Reporting of the leafs below a certain node can be done by a simple Depth First Search reporting any leafs found. If one is interested in knowing only whether the string occurs, then the lines which report the matches can be substituted by simple early exist returning true.

This algorithm runs in time $O(m^{k+1})$.

### 2005-07-29 paper/space.tex
Break into little tex files

\begin{figure}
\includegraphics{figures/aaaa}
\caption{$a^n$: worst case space consumption}%
\label{fig:aaaa}
\end{figure}

How much space does a k-error dotted tree take? Suffix Trees have a number of nodes proportional to the size of the string (which we will refer to as $n$). One-error dotted trees have more nodes. The worst case presented in Figure~\ref{fig:aaaa} is a string of the form $a^n = \underbrace{aaa\cdots}_{n \mathit{times }a}$ which generates a one-dotted tree taking $O(n^2)$ nodes. For a larger number of errors $k$, it is easy to see that this leads to $O(n^{k+1})$ nodes.

\begin{definition}[weight under a node, weight of a tree]
The weight under a node $\mathit{Weight}(n)$ is the number of leafs in the subtree rooted at that node.

The total weight of a tree is the sum of the weights of all the nodes.
\end{definition}

\begin{definition}[height of a node, average height]
The height of a node $\mathit{Height}(n)$ is the number of nodes in the path leading to that node from the root. The average height of the tree $\hat{h}$, is the average height of its nodes.

\[ \hat{h} = \frac{1}{n} \sum_{x \mathit{node of \mathcal{T}}} \mathit{Height}(x), n \mathrm{being the number of nodes} \]
\end{definition}

\begin{lemma}
Given a tree (any tree, not just suffix or similiar tree), the sum of the weights of all the nodes in the tree is equal to the average height of the node times the number of nodes.
\end{lemma}

Consider the tree being built node by node, in any order, as long as the property that it remains a tree during the process is maintained. Each node added, adds a certain amount of weight to the tree. In particular it adds one for each node that sits above it in the tree. This is the same as saying that each node adds its height (which does not change). The weight of the final tree is the sum of all these values.

\[ \mathit{Weight}(\mathcal{T}) = \sum_{x \in \mathcal{T}}\mathit{Height}(x) = n\frac{1}{n}\sum_{x \in \mathcal{T}}\mathit{Height}(x) = n\hat{h} \qed \]

We will reason by induction, starting with the case of a one-error dotted tree to show, that the bound on the number of nodes in an k-error dotted tree can also be expressed as $O(n\hat{h}^k)$ where $\hat{h}$ is the average height of the suffix tree.

The number of nodes in a Patricia tree with $c$ strings ($c$ leafs) is $O(c)$~\cite{patricia} (as a special case, a suffix tree is the Patricia tree of all the suffixes in a string). The number of nodes in the error tree of a node is therefore assymptotically bound by the number of leafs in the subtree at that node times a constant.

We define $\mathit{Weight}(\overline{w})$ as above. The total number of nodes for all the error trees in the dotted tree is therefore equal to $\sum_{\overline{w} \in \mathcal{T}}\mathit{Weight}(\overline{w})$ where $\mathcal{T}$ is the original suffix tree. Given the lemma above, $\sum_{\overline{w} \in \mathcal{T}}\mathit{Weight}(\overline{w}) = O(n\hat{h})$.

The difference between a k-error and a (k+1)-error tree is that all the

Sofar, we have not really obtained much since, in the worst case, $h=O(n)$. However, the average expected case, is $h=O(\log n)$~\cite{devroye:note,szpankowski:unexpected}.

### 2005-07-29 paper/structure.tex
Break into little tex files

\subsection{Intuition}

\begin{figure}
\includegraphics{figures/mississipi-0}
\caption{Suffix Tree for the string \putstring{mississipi}}%
\label{fig:mississipi-0}
\end{figure}

Suffix Trees have become a well known structure since their introduction in 1973~\cite{weiner}. In Figure~\ref{fig:mississipi-0} we show the suffix tree for the string \putstring{mississipi}.

Being at a node in the suffix tree corresponds to being in several points in the string at the same time. We now add to each node an extra edge which corresponds to moving one character forward in all these points.

\subsection{Formal definition}

We assume that the reader is familiar with the basic concepts of suffix trees and strings and present the definitions below mainly to introduce the particular notation and terminology used in this paper.

\begin{definition}[Character, string]
Given a set $\Sigma$, we say that $s$ is a \emph{string over $\Sigma$} if $s$ is a possibly empty sequence of elements of $\Sigma$. The elements of the set $\Sigma$ are also called \emph{character}. The lenght of the string, denoted by $|s|$ is the number of elements it contains. We shall write $s_i$ for the $i$th element of the string. For the empty string, we write $\epsilon$.

The set of all string is denoted by $\Sigma^*$ and $\Sigma^+=\Sigma^*-\{\epsilon\}$.
\end{definition}

For denoting character we shall use letters from the begginning of the roman alphabet ($a$, $b$, $c$,\ldots) and for strings, we shall use letters from the end of the alphabet ($w$, $z$, \ldots).

\begin{definition}
$wx$ will denote the string formed by all the character of $w$ followed by all the characters of $x$ (the usual concatenation operation). $aw$ ($wa$) shall the denote the strings by the caracter $a$ before (after) the characters of $w$.

If $s = wxy$, then $w$ is a \emph{prefix} of $s$, $x$ is a \emph{substring} of $s$ (at position $s$) and $y$ is a \emph{suffix} of $s$ (at position $p$).
\end{definition}

\begin{definition}[Patricia tree]
$T$ is a Patricia tree if $T$ is a rooted tree with edge labels from $\Sigma^+$. For each $a \in \Sigma$ and every node $n$ in $T$, there is at most an edge leaving $n$ whose label is $aw$.

Each node in a $\Sigma^+$ tree has a path leading to it which forms a string. If the node $n$ has the leading path $w$, we shall also refer to $n$ as $\overline{w}$.
\end{definition}

The name Patricia is sometimes spelt PATRICIA, since it was formed from the initial of Practical Algorithm to Retrieve Information Coded in Alphanumeric~\cite{morrison:patricia}.

In what follows we will assume that there are two symbols (\$ and $.$) which are not part of $\Sigma$.

In this paper, we will be interested in substitution of Hamming distance:

\begin{definition}[Aproximate Match]
We say that the string $s$ matches the string $t$ at position $p$ with $k$ errors if we can make $k$ substitutions in $s$ such that $s$ is a substring of $t$ at position $p$.
\end{definition}

\begin{definition}[Pattern]
A pattern is a string over the extended alphabet $\Sigma\cup\{.\}$. We say that a pattern $s$ matches a string $t$ at position $p$, if for each dot in the pattern we can replace it by a character in the original alphabet to obtain $s'$ which is a substring of $t$ at position $p$.
\end{definition}

\begin{definition}[Suffix Tree]
A \emph{suffix tree} for a string $S$ is a $\Sigma^+$ whose leaf nodes (those without children) have paths corresponding to suffixes of the string $S\$ $.
\end{definition}

\begin{definition}[Suffix Link]
A suffix link in a suffix tree is a link from the node $\overline{aw}$ to the node $\overline{w}$. This node has the label $a$.
\end{definition}

In a suffix tree, it is possible to define a suffix link for each internal node. Both Ukkonen's~\cite{ukkonen} and McCreight's~\cite{mccreight} algorithms are linear time algorithms for constructing a suffix tree which includes suffix links.

\begin{definition}[Occurence Set]
Given a node $\overline{w}$ in a suffix tree, we call it's \emph{occurence set} the set of indexes in the original string where the string $w$ occurs.
\end{definition}

\begin{definition}[Position Set]
Given a node $\overline{w}$ in a suffix tree, it's \emph{position set} is the set formed by taking it's occurence set and adding the lenght of $w$ to each element.
\end{definition}

For example, in the node $\overline{issi}$ in the suffix tree of Figure~\ref{fig:mississipi-0}, the occurence set is $\{5, 2\}$ and its position set is $\{6, 9\}$. In a sense, one can say that being at node $\overline{issi}$ is being at positions 6 and~9 simultaneously. Note that for each node, the tree below it is the $\Sigma^+$ tree of the suffixes starting at the positions in its position set.

In a leaf, the position set is a singleton, and we label the leaf by its element.

\begin{definition}[Error Tree]
For any node $w$ its error tree is the $\Sigma^+$ tree formed by taking it's position set, adding one to each element and forming the $\Sigma^+$ tree of the suffixes of the suffixes starting at those positions. If the position set includes the end of the string, that element is removed.

The leafs are labeled by the position of the string in which their paths occur minus $|w| + 1$.
\end{definition}

For example, in the above mentioned node $\overline{issi}$, the error tree is formed by taking the strings starting at positions $\{7, 10\}$ (ie \putstring{sipi\$} and \putstring{i\$}) in a $\Sigma^+$ tree.

\begin{definition}[1-error dotted Tree]
We define a \emph{1-error dotted tree} as the tree which add to each node in a suffix tree, a new edge labeled by \putstring{.} which points to its error tree. The edge labeled \putstring{.} shall be called a \emph{dot link}.
\end{definition}

The one-error dotted tree for \putstring{mississipi} is shown in Figure~\ref{fig:mississipi-1}.

The paths in the dotted tree are paths in the extended alphabet $\Sigma\cup\{.,\$\}$. The notions of \emph{occurence set}, \emph{position set} and \emph{error tree} are valid for all nodes in a dotted tree considering that a node path may contain dot elements which match any other character.

\begin{definition}[k-error dotted tree]
We define a k-error dotted tree as the tree obtained by adding error trees to each node in the (k-1)-error dotted tree.
\end{definition}

### 2005-07-29 paper/construction.tex
Polish it a bit Still not proved O(nh^k)!

@@
 Note that these conditions are very similiar to the conditions for the root and both can be implemented as shown in Figure~\ref{algo:copy-subtree}.
+This is an expression of Lemma~\ref{lemma:suffix-error}.
 
@@
 
-To construct the (k+1)-error tree from the k-error tree, make a copy of the tree as above (adjusting leafs, filtering the leafs labeled one) and make this the root's error tree.
+To construct the (k+1)-error tree from the k-error tree, make a copy of the tree as above (adjusting leafs and filtering the leafs labeled one) and make this the new root error tree.
 Then, for every other node, remove the current error tree (it has a level too few).
 Finally, for every node except the root, construct its error tree as above (adjusting leafs, filtering the leafs labeled one or which do not correspond to positions in the suffix link's label).
+These error trees will now have the correct number of nodes.
 

### 2005-07-29 paper/space.tex
Polish it a bit Still not proved O(nh^k)!

@@
 
-\[ \hat{h} = \frac{1}{n} \sum_{x \mathit{node of \mathcal{T}}} \mathit{Height}(x), n \mathrm{being the number of nodes} \]
+\[ \hat{h} = \frac{1}{|\mathcal{T}|} \sum_{x \in \mathit{\mathcal{T}}} \mathit{Height}(x) \]
 \end{definition}
@@
 
-\[ \mathit{Weight}(\mathcal{T}) = \sum_{x \in \mathcal{T}}\mathit{Height}(x) = n\frac{1}{n}\sum_{x \in \mathcal{T}}\mathit{Height}(x) = n\hat{h} \qed \]
+\[ \mathit{Weight}(\mathcal{T}) = \sum_{x \in \mathcal{T}}\mathit{Height}(x) = |\mathcal{T}|\frac{1}{|\mathcal{T}|}\sum_{x \in \mathcal{T}}\mathit{Height}(x) = |\mathcal{T}|\hat{h} \qed \]
 
-We will reason by induction, starting with the case of a one-error dotted tree to show, that the bound on the number of nodes in an k-error dotted tree can also be expressed as $O(n\hat{h}^k)$ where $\hat{h}$ is the average height of the suffix tree.
+We will reason by induction, starting with the case of a zero-error dotted tree to show, that the bound on the number of nodes in an k-error dotted tree can also be expressed as $O(n\hat{h}^k)$ where $\hat{h}$ is the average height of the suffix tree.
+The base case is a suffix tree and is obvious.
 
@@
 
-We define $\mathit{Weight}(\overline{w})$ as above.
-The total number of nodes for all the error trees in the dotted tree is therefore equal to $\sum_{\overline{w} \in \mathcal{T}}\mathit{Weight}(\overline{w})$ where $\mathcal{T}$ is the original suffix tree.
+%Assume the k-error suffix tree has $O(h\hat{h}^k)$ nodes.
+
+%We define $\mathit{Weight}(\overline{w})$ as above.
+For any node, the number of leafs in its error tree will be $O(\mathit{Weight}(\overline{w}))$.
+The total number of nodes for all the error trees in the dotted tree is therefore equal to $O(\sum_{\overline{w} \in \mathcal{T}}\mathit{Weight}(\overline{w}))$ where $\mathcal{T}$ is the original suffix tree.
 Given the lemma above, $\sum_{\overline{w} \in \mathcal{T}}\mathit{Weight}(\overline{w}) = O(n\hat{h})$.
 
-The difference between a k-error and a (k+1)-error tree is that all the
+%The difference between a k-error and a (k+1)-error tree is that all the
 

### 2005-07-29 paper/structure.tex
Polish it a bit Still not proved O(nh^k)!

@@
 
-Each node in a $\Sigma^+$ tree has a path leading to it which forms a string. If the node $n$ has the leading path $w$, we shall also refer to $n$ as $\overline{w}$.
+Each node in a Patricia tree has a path leading to it which forms a string. If the node $n$ has the leading path $w$, we shall also refer to $n$ as $\overline{w}$.
 \end{definition}
@@
 In a sense, one can say that being at node $\overline{issi}$ is being at positions 6 and~9 simultaneously.
-Note that for each node, the tree below it is the $\Sigma^+$ tree of the suffixes starting at the positions in its position set.
+Note that for each node, the tree below it is the Patricia tree of the suffixes starting at the positions in its position set.
 
-In a leaf, the position set is a singleton, and we label the leaf by its element.
+In a leaf, the occurence set is a singleton, and we label the leaf by its element.
+
+\begin{lemma}[Position set at the suffix node]\label{lemma:suffix-error}
+Given two nodes $\overline{aw}$ and $\overline{w}$ (connected by a suffix link, labeled $a$), if one takes position set of $\overline{w}$, substracts one to each element, one obtains a superset of the position set of $\overline{w}$.
+The items shared by both sets are those positions of the string which contain an $a$.
+\end{lemma}
+
+The lemma is fairly obvious given that the position set of $\overline{aw}$ contains all the positions where $aw$ occurs which will be exactly one minus those positions where $w$ occurs following $a$.
 
@@
 
-For example, in the above mentioned node $\overline{issi}$, the error tree is formed by taking the strings starting at positions $\{7, 10\}$ (ie \putstring{sipi\$} and \putstring{i\$}) in a $\Sigma^+$ tree.
+For example, in the above mentioned node $\overline{issi}$, the error tree is formed by taking the strings starting at positions $\{7, 10\}$ (ie \putstring{sipi\$} and \putstring{i\$}) in a Patricia tree.
 
@@
 
-The one-error dotted tree for \putstring{mississipi} is shown in Figure~\ref{fig:mississipi-1}.
+\begin{figure}
+\includegraphics{figures/mississipi-1}
+\caption{1-error dotted tree for \putstring{mississipi}}%
+\label{fig:mississipi-1}
+\end{figure}
+
+The one-error dotted tree for \putstring{mississipi} is shown in Figure~\ref{fig:mississipi-1}. The nodes are connected to their error trees by diagonal links.
 
@@
 \begin{definition}[k-error dotted tree]
-We define a k-error dotted tree as the tree obtained by adding error trees to each node in the (k-1)-error dotted tree.
+We define a k-error dotted tree as the tree obtained by adding error trees to each node in the (k-1)-error dotted tree which does not already contain one.
 \end{definition}

### 2005-08-11 paper/construction.tex
Small fixes. Longer, more detailed explanations.

@@
 
+These conditions are an expression of Lemma~\ref{lemma:suffix-error}.
 Note that these conditions are very similiar to the conditions for the root and both can be implemented as shown in Figure~\ref{algo:copy-subtree}.
-This is an expression of Lemma~\ref{lemma:suffix-error}.
 
@@
 To add dot links at every node, we need to add the dot link at the root and then for every node call the procedure addDotLink.
-This is shown in Figure~\ref{algo:construct}.
-The \textit{foreach} loop can be implemented using Depth First Search on the tree.
+This is shown in Figure~\ref{algo:construct-1}.
+The \textit{foreach} loop at line~\ref{algoline:construct-1:foreach} can be implemented using Depth First Search on the tree.
 In fact, the procedure addDotLink only needs to be called for certain nodes (those who do not posess an incoming suffix link).

### 2005-08-11 paper/introduction.tex
Small fixes. Longer, more detailed explanations.

@@
 Since their introduction~\cite{weiner}, suffix trees have been one of the methods of choice for text indexing.
-They allow one to solve the following problem: given a text which has been preprocessed, and a pattern, efficiently find the places in the text where this pattern occurs.
-However, for many real-life problems this statement is too restrictive.
-In these problems, one is interested in finding places in the text where an approximate form of the pattern occurs.
+Suffix trees solve the following problem: given a text which has been preprocessed, and a pattern, efficiently find the places in the text where this pattern occurs.
+However, for many real-life problems this is too restrictive.
+In this class of problems, one is interested in finding places in the text where an approximate form of the pattern occurs.
 Several algorithms have been proposed to solve this problem.
@@
 
-In this work we present an approach which solves this problem based on extending of suffix trees.
+In this work we present an approach which solves this problem based on an extention of suffix trees.
 The main advantage of this approach is that both the search and the index size are \emph{alphabet independent}, although the indexing time is not.

### 2005-08-11 paper/search.tex
Small fixes. Longer, more detailed explanations.

@@
 \begin{figure}
-\includegraphics{figures/mississipi-nodes}
+\includegraphics[width=\textwidth]{figures/mississipi-nodes}
 \caption{Suffix Tree for the string \putstring{mississipi}}%
@@
 
-This algorithm runs in time $O(m^{k+1})$.
+This algorithm runs in time $O(m^{k+1})$:
+
+The recursive call at line~\ref{algoline:search:followDot} means that a dot link was followed.
+There are at most $O(m^k)$ places where a dot link can be followed while descending a string~\footnote{For $k=1$ each point in the string.
+For $k>1$, each path for $k-1$, can now have at most $m$ extra branches.}.
+Following one path in the tree takes $O(m)$ time.
+Therefore, the total time is $O(m^{k+1})$.

### 2005-08-11 paper/space.tex
Small fixes. Longer, more detailed explanations.

@@
 One-error dotted trees have more nodes.
-The worst case presented in Figure~\ref{fig:aaaa} is a string of the form $a^n = \underbrace{aaa\cdots}_{n \mathit{times }a}$ which generates a one-dotted tree taking $O(n^2)$ nodes.
+The worst case presented in Figure~\ref{fig:aaaa} is a string of the form $a^n = aaa\cdots$ which generates a one-dotted tree taking $O(n^2)$ nodes.
 For a larger number of errors $k$, it is easy to see that this leads to $O(n^{k+1})$ nodes.
+
+We can do better, however, by considering the expected case. First we need a pair of auxiliar definitions:
 
@@
 \end{definition}
+
+Note that the weight of a tree \emph{is not the same} as the weight at the root.
 
@@
 
-Sofar, we have not really obtained much since, in the worst case, $h=O(n)$. However, the average expected case, is $h=O(\log n)$~\cite{devroye:note,szpankowski:unexpected}.
+Sofar, we have not really obtained much since, in the worst case, $h=O(n)$.
+However, the average expected case, is $h=O(\log n)$~\cite{devroye:note,szpankowski:unexpected}.
+This allows us to conclude that, in the average case, we will have $O(n\log^k n)$ nodes in our tree.

### 2005-08-11 paper/structure.tex
Small fixes. Longer, more detailed explanations.

@@
 
-Being at a node in the suffix tree corresponds to being in several points in the string at the same time.
+To extend the suffix tree, we consider that a node in the suffix tree corresponds to several points in the string at the same time (those points where the string on the path to the node occurs).
 We now add to each node an extra edge which corresponds to moving one character forward in all these points.
@@
 
-In this paper, we will be interested in substitution of Hamming distance:
+In this paper, we will be interested in substitution distance (aka Hamming distance):
 
@@
 Both Ukkonen's~\cite{ukkonen} and McCreight's~\cite{mccreight} algorithms are linear time algorithms for constructing a suffix tree which includes suffix links.
+The suffix links are shown in Figure~\ref{fig:mississipi-0} as dotted links.
 
@@
 \begin{definition}[1-error dotted Tree]
-We define a \emph{1-error dotted tree} as the tree which add to each node in a suffix tree, a new edge labeled by \putstring{.} which points to its error tree.
+We define a \emph{1-error dotted tree} as the tree which is formed by adding to each node in a suffix tree, a new edge labeled by \putstring{.} which points to its error tree.
 The edge labeled \putstring{.} shall be called a \emph{dot link}.
@@
 \begin{figure} 
-\includegraphics{figures/mississipi-1}
+\includegraphics[width=\textwidth]{figures/mississipi-1.eps}
+% mississipi-1.eps: 300dpi, width=6.08cm, height=1.63cm, bb=0 0 718 193
 \caption{1-error dotted tree for \putstring{mississipi}}%

### 2005-08-11 paper/abstract.tex
Spell check

@@
 We can later query this index to identify the places where a string occurs up to a certain number of allowed substitutions $k$.
-We present a structure for this indexing which occupies space $O(n\log^kn)$ in the average case, indepedent of alphabet size, $n$ being the text size.
+We present a structure for this indexing which occupies space $O(n\log^kn)$ in the average case, independent of alphabet size, $n$ being the text size.
 This structure support searching in $O(m^{k+1})$ time, for patterns of size $m$, again independently of alphabet size.

### 2005-08-11 paper/construction.tex
Spell check

@@
 The leaf labeled 1 would have resulted in the string \putstring{.s\$} which does not occur in original string (it is one character too large).
-For any other leaf $\overline{w\$}$ which occured at position $p$ in the string, we have a new leaf $\overline{.w\$}$ which occurs at position $p-1$ in the string.
+For any other leaf $\overline{w\$}$ which occurred at position $p$ in the string, we have a new leaf $\overline{.w\$}$ which occurs at position $p-1$ in the string.
 
@@
 These conditions are an expression of Lemma~\ref{lemma:suffix-error}.
-Note that these conditions are very similiar to the conditions for the root and both can be implemented as shown in Figure~\ref{algo:copy-subtree}.
+Note that these conditions are very similar to the conditions for the root and both can be implemented as shown in Figure~\ref{algo:copy-subtree}.
 
@@
 As discussed above, suffix trees are normally implemented such that each node is actually just a pair of indices into the original string as shown in Figure~\ref{fig:mississipi-nodes}.
-To merge a child with its parent, we just need to replace both by a node whose starting index is the starting index of the original child minus the lenght of the parent edge.
+To merge a child with its parent, we just need to replace both by a node whose starting index is the starting index of the original child minus the length of the parent edge.
 This process is illustrated in Figure~\ref{fig:merge}
@@
 The \textit{foreach} loop at line~\ref{algoline:construct-1:foreach} can be implemented using Depth First Search on the tree.
-In fact, the procedure addDotLink only needs to be called for certain nodes (those who do not posess an incoming suffix link).
+In fact, the procedure addDotLink only needs to be called for certain nodes (those who do not posses an incoming suffix link).
 However, there is no readily available way to access only these nodes and the procedure above does little extra work.
@@
 
-We note that the method of removing the existing error trees can be improved in practice (the assymptotical bound remains the same) the following way: While making a copy of the tree for construction of the root's error tree, instead of following a dot link to make a copy, move the error tree to its destination and make the necessary adjustments inplace (this move will probably be just an adjustment of pointer depending on the exact implementation).
+We note that the method of removing the existing error trees can be improved in practice (the asymptotic bound remains the same) the following way: While making a copy of the tree for construction of the root's error tree, instead of following a dot link to make a copy, move the error tree to its destination and make the necessary adjustments inplace (this move will probably be just an adjustment of pointer depending on the exact implementation).
 Using this method only leafs labeled one need to be removed (as well as any internal node which becomes unnecessary).

### 2005-08-11 paper/introduction.tex
Spell check

@@
 
-In this work we present an approach which solves this problem based on an extention of suffix trees.
+In this work we present an approach which solves this problem based on an extension of suffix trees.
 The main advantage of this approach is that both the search and the index size are \emph{alphabet independent}, although the indexing time is not.
@@
 Our work is related to efforts to index the text for approximate matching such as~\cite{maass:text,mosche}.
-The structure presented is superficially very similiar to the one presented in~\cite{inexact} as an \textit{inexact suffix tree} though that work is concerned with solving a different set of problems.
+The structure presented is superficially very similar to the one presented in~\cite{inexact} as an \textit{inexact suffix tree} though that work is concerned with solving a different set of problems.
 
@@
 In section~\ref{sec:structure} we present the structure and the way to use it for searching.
-In section~\ref{sec:build-1} we show the way to build a structure to handle one error only and analize this version of the algorithm.
-Then, we show, in section~\ref{sec:build-k} how to generalize this procedure for any number of errors and how the previous analisys also applies to any number of errors.
+In section~\ref{sec:build-1} we show the way to build a structure to handle one error only and analyze this version of the algorithm.
+Then, we show, in section~\ref{sec:build-k} how to generalize this procedure for any number of errors and how the previous analysis also applies to any number of errors.

### 2005-08-11 paper/space.tex
Spell check

@@
 
-We can do better, however, by considering the expected case. First we need a pair of auxiliar definitions:
+We can do better, however, by considering the expected case. First we need a pair of definitions:
 
@@
 \begin{lemma}
-Given a tree (any tree, not just suffix or similiar tree), the sum of the weights of all the nodes in the tree is equal to the average height of the node times the number of nodes.
+Given a tree (any tree, not just suffix or similar tree), the sum of the weights of all the nodes in the tree is equal to the average height of the node times the number of nodes.
 \end{lemma}
@@
 The number of nodes in a Patricia tree with $c$ strings ($c$ leafs) is $O(c)$~\cite{patricia} (as a special case, a suffix tree is the Patricia tree of all the suffixes in a string).
-The number of nodes in the error tree of a node is therefore assymptotically bound by the number of leafs in the subtree at that node times a constant.
+The number of nodes in the error tree of a node is therefore asymptomatically bound by the number of leafs in the subtree at that node times a constant.
 
@@
 
-Sofar, we have not really obtained much since, in the worst case, $h=O(n)$.
+So far, we have not really obtained much since, in the worst case, $h=O(n)$.
 However, the average expected case, is $h=O(\log n)$~\cite{devroye:note,szpankowski:unexpected}.

### 2005-08-11 paper/structure.tex
Spell check

@@
 The elements of the set $\Sigma$ are also called \emph{character}.
-The lenght of the string, denoted by $|s|$ is the number of elements it contains.
+The length of the string, denoted by $|s|$ is the number of elements it contains.
 We shall write $s_i$ for the $i$th element of the string.
@@
 
-For denoting character we shall use letters from the begginning of the roman alphabet ($a$, $b$, $c$,\ldots) and for strings, we shall use letters from the end of the alphabet ($w$, $z$, \ldots).
+For denoting character we shall use letters from the beginning of the roman alphabet ($a$, $b$, $c$,\ldots) and for strings, we shall use letters from the end of the alphabet ($w$, $z$, \ldots).
 
@@
 $wx$ will denote the string formed by all the character of $w$ followed by all the characters of $x$ (the usual concatenation operation).
-$aw$ ($wa$) shall the denote the strings by the caracter $a$ before (after) the characters of $w$.
+$aw$ ($wa$) shall the denote the strings by the character $a$ before (after) the characters of $w$.
 
@@
 
-The name Patricia is sometimes spelt PATRICIA, since it was formed from the initial of Practical Algorithm to Retrieve Information Coded in Alphanumeric~\cite{morrison:patricia}.
+The name Patricia is sometimes spelled PATRICIA, since it was formed from the initial of Practical Algorithm to Retrieve Information Coded in Alphanumeric~\cite{morrison:patricia}.
 
@@
 \begin{definition}[Occurence Set]
-Given a node $\overline{w}$ in a suffix tree, we call it's \emph{occurence set} the set of indexes in the original string where the string $w$ occurs.
+Given a node $\overline{w}$ in a suffix tree, we call it's \emph{occurrence set} the set of indexes in the original string where the string $w$ occurs.
 \end{definition}
@@
 \begin{definition}[Position Set]
-Given a node $\overline{w}$ in a suffix tree, it's \emph{position set} is the set formed by taking it's occurence set and adding the lenght of $w$ to each element.
+Given a node $\overline{w}$ in a suffix tree, it's \emph{position set} is the set formed by taking it's occurrence set and adding the length of $w$ to each element.
 \end{definition}
 
-For example, in the node $\overline{issi}$ in the suffix tree of Figure~\ref{fig:mississipi-0}, the occurence set is $\{5, 2\}$ and its position set is $\{6, 9\}$.
+For example, in the node $\overline{issi}$ in the suffix tree of Figure~\ref{fig:mississipi-0}, the occurrence set is $\{5, 2\}$ and its position set is $\{6, 9\}$.
 In a sense, one can say that being at node $\overline{issi}$ is being at positions 6 and~9 simultaneously.
@@
 
-In a leaf, the occurence set is a singleton, and we label the leaf by its element.
+In a leaf, the occurrence set is a singleton, and we label the leaf by its element.
 
 \begin{lemma}[Position set at the suffix node]\label{lemma:suffix-error}
-Given two nodes $\overline{aw}$ and $\overline{w}$ (connected by a suffix link, labeled $a$), if one takes position set of $\overline{w}$, substracts one to each element, one obtains a superset of the position set of $\overline{w}$.
+Given two nodes $\overline{aw}$ and $\overline{w}$ (connected by a suffix link, labeled $a$), if one takes position set of $\overline{w}$, subtracts one to each element, one obtains a superset of the position set of $\overline{w}$.
 The items shared by both sets are those positions of the string which contain an $a$.
@@
 The paths in the dotted tree are paths in the extended alphabet $\Sigma\cup\{.,\$\}$.
-The notions of \emph{occurence set}, \emph{position set} and \emph{error tree} are valid for all nodes in a dotted tree considering that a node path may contain dot elements which match any other character.
+The notions of \emph{occurrence set}, \emph{position set} and \emph{error tree} are valid for all nodes in a dotted tree considering that a node path may contain dot elements which match any other character.
 

### 2005-08-12 paper/main.tex
Back in Europe
Small fixes. Longer, more detailed explanations.
Spell check
fix references

@@
-\documentclass[letter,10pt]{article}
+\documentclass[a4paper,10pt]{article}
 \usepackage[boxruled,vlined,portugues,figure,linesnumbered,longend]{algorithm2e}
@@
 \usepackage[]{graphicx}
+\usepackage{amssymb}
 \usepackage{QED}
-\usepackage{amssymb}
 
@@
 \newtheorem{lemma}{Lemma}
+\graphicspath{{figures/}}
 
@@
 
-\section{Introduction}
+\section{Introduction}\label{sec:introduction}
 \input{introduction}
 
-\section{The Indexing Structure}
+\section{The Indexing Structure}\label{sec:structure}
 \input{structure}
-\section{Space Considerations}
+
+\section{Space Considerations}\label{sec:space}
 \input{space}
-\section{Searching}
+
+\section{Searching}\label{sec:search}
 \input{search}
-\section{Constructing the Dotted Tree}
+
+\section{Constructing the Dotted Tree}\label{sec:construction}
 \input{construction}
-\section{Experimental Results}
+
+\section{Experimental Results}\label{sec:experimental}
 \input{experimental}
-\section{Future Work, Open Problems}
+
+\section{Future Work, Open Problems}\label{sec:future}
 
 In the example for the string \putstring{mississipi}, presented in Figure~\ref{fig:mississipi-1}, one can see that the tree below \putstring{s.i} and \putstring{ssi} are exactly the same.
-Whether such occurences are the basis for a significant space saving and what algorithms mightexploit them is an open question.
+Whether such occurrences are the basis for a significant space saving and what algorithms might exploit them is an open question.
 Along the same lines, the structure's definition might be extended to structures such as the suffix-DAG presented in~\cite[7.7]{gusfield:algorithms}.

### 2005-08-12 paper/space.tex
fix references

@@
 \[ \hat{h} = \frac{1}{|\mathcal{T}|} \sum_{x \in \mathit{\mathcal{T}}} \mathit{Height}(x) \]
+
+\[ h_{max} = \mathit{max}_{x \in \mathcal{T}} \mathit{Height}(x) \]
+
 \end{definition}
@@
 
+\Proof
 Consider the tree being built node by node, in any order, as long as the property that it remains a tree during the process is maintained.
@@
 \[ \mathit{Weight}(\mathcal{T}) = \sum_{x \in \mathcal{T}}\mathit{Height}(x) = |\mathcal{T}|\frac{1}{|\mathcal{T}|}\sum_{x \in \mathcal{T}}\mathit{Height}(x) = |\mathcal{T}|\hat{h} \qed \]
+
+\begin{lemma}
+Given a tree (any tree), we have
+
+\[ h_{max} \leq 2\hat{h} \]
+\end{lemma}
+
+\Proof Assume $h_{max} > 2\hat{h}$.
+Then there is a path, ending in a leaf which lies at depth $h_{max}$.
+Over this path, there are $h_{max}$ nodes at heights 1, 2, \dots $h_{max}$.
+The average height over this path is $\frac{1}{h_{max}}\sum_{i=1}^{h_{max}}i=\frac{h_{max}-1}{2}$.
 

### 2005-08-30 paper/structure.tex
fix references
s/mississipi/mississippi/

@@
 \begin{figure}
-\includegraphics{figures/mississipi-0}
-\caption{Suffix Tree for the string \putstring{mississipi}}%
-\label{fig:mississipi-0}
+\includegraphics{figures/mississippi-0}
+\caption{Suffix Tree for the string \putstring{mississippi}}%
+\label{fig:mississippi-0}
 \end{figure}
 
-Suffix Trees have become a well known structure since their introduction in 1973~\cite{weiner}. In Figure~\ref{fig:mississipi-0} we show the suffix tree for the string \putstring{mississipi}.
+Suffix Trees have become a well known structure since their introduction in 1973~\cite{weiner}. In Figure~\ref{fig:mississippi-0} we show the suffix tree for the string \putstring{mississippi}.
 
@@
 Both Ukkonen's~\cite{ukkonen} and McCreight's~\cite{mccreight} algorithms are linear time algorithms for constructing a suffix tree which includes suffix links.
-The suffix links are shown in Figure~\ref{fig:mississipi-0} as dotted links.
+The suffix links are shown in Figure~\ref{fig:mississippi-0} as dotted links.
 
@@
 
-For example, in the node $\overline{issi}$ in the suffix tree of Figure~\ref{fig:mississipi-0}, the occurrence set is $\{5, 2\}$ and its position set is $\{6, 9\}$.
+For example, in the node $\overline{issi}$ in the suffix tree of Figure~\ref{fig:mississippi-0}, the occurrence set is $\{5, 2\}$ and its position set is $\{6, 9\}$.
 In a sense, one can say that being at node $\overline{issi}$ is being at positions 6 and~9 simultaneously.
@@
 
-For example, in the above mentioned node $\overline{issi}$, the error tree is formed by taking the strings starting at positions $\{7, 10\}$ (ie \putstring{sipi\$} and \putstring{i\$}) in a Patricia tree.
+For example, in the above mentioned node $\overline{issi}$, the error tree is formed by taking the strings starting at positions $\{7, 10\}$ (ie \putstring{sippi\$} and \putstring{i\$}) in a Patricia tree.
 
@@
 \begin{figure} 
-\includegraphics[width=\textwidth]{figures/mississipi-1.eps}
-% mississipi-1.eps: 300dpi, width=6.08cm, height=1.63cm, bb=0 0 718 193
-\caption{1-error dotted tree for \putstring{mississipi}}%
-\label{fig:mississipi-1}
+\includegraphics[width=\textwidth]{figures/mississippi-1.eps}
+% mississippi-1.eps: 300dpi, width=6.08cm, height=1.63cm, bb=0 0 718 193
+\caption{1-error dotted tree for \putstring{mississippi}}%
+\label{fig:mississippi-1}
 \end{figure}
 
-The one-error dotted tree for \putstring{mississipi} is shown in Figure~\ref{fig:mississipi-1}. The nodes are connected to their error trees by diagonal links.
+The one-error dotted tree for \putstring{mississippi} is shown in Figure~\ref{fig:mississippi-1}. The nodes are connected to their error trees by thick diagonal links.
 

### 2005-08-30 paper/space.tex
Make a better argument that this is O(n\log^k n)

@@
 \begin{definition}[weight under a node, weight of a tree]
-The weight under a node $\mathit{Weight}(n)$ is the number of leafs in the subtree rooted at that node.
+The weight under a node $\mathit{Weight}(n)$ is the number of nodes in the subtree rooted at that node.
 
@@
 
+\begin{definition}[Height of error tree]
+Given an integer k, we define the height $h_k$ as the average height of all the error trees which exist in the k-error dotted tree and not in the (k-1)-error tree.
+$h_0$ is the average height of the original suffix tree.
+\end{definition}
+
 \begin{lemma}
-Given a tree (any tree), we have
-
-\[ h_{max} \leq 2\hat{h} \]
+The number of nodes of an k-error dotted tree is bounded by $O(nh_0h_1\dots{}h_{k-1})$.
 \end{lemma}
 
-\Proof Assume $h_{max} > 2\hat{h}$.
-Then there is a path, ending in a leaf which lies at depth $h_{max}$.
-Over this path, there are $h_{max}$ nodes at heights 1, 2, \dots $h_{max}$.
-The average height over this path is $\frac{1}{h_{max}}\sum_{i=1}^{h_{max}}i=\frac{h_{max}-1}{2}$.
+\Proof
+By induction.
+For zero errors, the number of nodes of a suffix tree is known to be $O(n)$~\cite{}.
+Taking a (k-1)-error dotted tree (which we assume, by induction, to have $O(nh_0\dots{}h_{k-2})$ nodes), we can add error trees to this to obtain a k-error dotted tree.
+These error trees will be added to the (k-1)-error trees, whose average height is $h_{k-1}$.
+The maximum total weight of these trees is the number of nodes times the average height, therefore $O(nh_0\dots{}h_{k-1})$.
+At each node, to which we add an error tree, that error tree has a number of nodes bounded by a multiple of the number of leafs of the node, which is in turn bounded by the weight at the node.
+The total number of added nodes is therefore bounded by the total weight $O(nh_0\dots{}h_{k-1})$ \qed.
 
-We will reason by induction, starting with the case of a zero-error dotted tree to show, that the bound on the number of nodes in an k-error dotted tree can also be expressed as $O(n\hat{h}^k)$ where $\hat{h}$ is the average height of the suffix tree.
-The base case is a suffix tree and is obvious.
-
-The number of nodes in a Patricia tree with $c$ strings ($c$ leafs) is $O(c)$~\cite{patricia} (as a special case, a suffix tree is the Patricia tree of all the suffixes in a string).
-The number of nodes in the error tree of a node is therefore asymptomatically bound by the number of leafs in the subtree at that node times a constant.
-
-%Assume the k-error suffix tree has $O(h\hat{h}^k)$ nodes.
-
-%We define $\mathit{Weight}(\overline{w})$ as above.
-For any node, the number of leafs in its error tree will be $O(\mathit{Weight}(\overline{w}))$.
-The total number of nodes for all the error trees in the dotted tree is therefore equal to $O(\sum_{\overline{w} \in \mathcal{T}}\mathit{Weight}(\overline{w}))$ where $\mathcal{T}$ is the original suffix tree.
-Given the lemma above, $\sum_{\overline{w} \in \mathcal{T}}\mathit{Weight}(\overline{w}) = O(n\hat{h})$.
-
-%The difference between a k-error and a (k+1)-error tree is that all the
-
-So far, we have not really obtained much since, in the worst case, $h=O(n)$.
-However, the average expected case, is $h=O(\log n)$~\cite{devroye:note,szpankowski:unexpected}.
+So far, we have not really obtained much since, in the worst case, $h_i=O(n)$.
+However, these are all Patricia trees of suffixes of the original text, and the average expected case, is $h_i=O(\log n)$~\cite{}.
 This allows us to conclude that, in the average case, we will have $O(n\log^k n)$ nodes in our tree.

### 2005-08-30 scripts/get-fileportion.py
Do repeated searches on the index

@@ -1,0 +1,31 @@
+#!/usr/bin/env python
+from sys import argv
+import random
+import os
+
+inputfile=argv[1]
+filelen=os.stat(inputfile)[6]
+
+lenght=int(argv[3])
+k=int(argv[4])
+
+if argv[2] == '.':
+	begin = random.randrange(filelen-lenght)
+else:
+	begin=int(argv[2])
+	if begin + lenght > filelen:
+		raise Exception('File not large enough')
+
+f=open(inputfile,'r')
+f.seek(begin)
+string = f.read(lenght)
+
+changed = []
+for point in random.sample(range(0,len(string)),k):
+	# '#' is a character which does not show up in any of our data files
+	if point > 0:
+		string = string[0:point-1] + '#' + string[point:]
+	else:
+		string = '#' + string[1:]
+
+print string

### 2005-08-30 paper/conclusion.tex
Make conclusion a real section.

We have presented an indexing structure for approximate text matching which takes $O(n\log^k n)$ space in the average case (this can be seen either in a theoretical or experimentally). This structure allows matches in $O(m^{k+1})$ time and can be constructed in $O(N|\Sigma|)$ time ($N$ being the actual number of nodes). The structure and the algorithms to construct it are very simple and should be easily implemented.

The biggest drawback of this structure is the amount of space it takes which might limit its applicability. Future research along these lines might address this issue (see below).

\subsection{Future Work, Open Problems}\label{sec:future}

In the example for the string \putstring{mississippi}, presented in Figure~\ref{fig:mississippi-1}, one can see that the tree below \putstring{s.i} and \putstring{ssi} are exactly the same. Whether such occurrences are the basis for a significant space saving and what algorithms might exploit them is an open question.

Along the same lines, the structure's definition might be extended to structures such as the suffix-DAG presented in~\cite[7.7]{gusfield:algorithms}. An algorithm to efficiently construct it over these suffix-DAGs is another open problem.

\subsection{Acknowledgements}

We thank Y for making their suffix tree implementation available~\cite{} over which we implemented our algorithms.

### 2005-08-30 paper/main.tex
Make conclusion a real section.

@@
 \newtheorem{lemma}{Lemma}
-\graphicspath{{figures/}}
+\graphicspath{{figures/}{graphics/}}
 
@@
 
-\section{Future Work, Open Problems}\label{sec:future}
-
-In the example for the string \putstring{mississipi}, presented in Figure~\ref{fig:mississipi-1}, one can see that the tree below \putstring{s.i} and \putstring{ssi} are exactly the same.
-Whether such occurrences are the basis for a significant space saving and what algorithms might exploit them is an open question.
-Along the same lines, the structure's definition might be extended to structures such as the suffix-DAG presented in~\cite[7.7]{gusfield:algorithms}.
-An algorithm to efficiently construct it over these suffix-DAGs is another open problem.
+\section{Conclusions}\label{sec:conclusion}
+\input{conclusion}
 

### 2005-08-30 paper/experimental.tex
Describe results

@@
 \begin{enumerate}
-\item ``Democracy in America'' by Alexis de Tocqueville as natural language data ser
-\item The DNA of S. Cervesiae
+\item ``Democracy in America'' by Alexis de Tocqueville as natural language example
+\item The DNA of S. Cervesiae (first chromossome)
 \item Randomly generated text
 \end{enumerate}
+
+The machine used for testing was an Intel Pentium~IV at 2.40GHz with 4Gb of main memory.
+
+\subsection{Indexing Time and Size}
+
+We report the indexing time and size (as a count of the nodes in the structure) as a function of the number of character (by using larger and larger suffixes of the datasets).
+
+\begin{figure}%
+\includegraphics{ratio-1}%
+\includegraphics{ratio-2}%
+\includegraphics{ratio-3}%
+\caption{Ratios of k-error and (k+1)-error dotted trees}\label{fig:ratios}%
+\end{figure}
+
+To experimentally verify the average case, we show in Figure~\ref{fig:ratios} the ratios between the k-error and the (k+1)-error dotted trees.
+We can easily see that for all three data sets, the experimental values do resemble a logarithm.
+
+\subsection{Search Time}
+
+For searching, the full textes were first indexed and then online searches were performed on top of the structure.
+To generate input patterns, randomly chosen bits of the text were taken.
+These patterns were then altered by substituting a character at random.
+To generate mismatches, one character too many was altered~\footnote{Since the characters used were not present in the original text, there is no risk of generating a string which \emph{does appear} in the text.}.
+The search algorithm performed an early exit, without reporting of occurences (ie, it only reported whether the string exists in the text).
+Therefore, the number of occurences had no influence on the search time.

### 2005-08-30 paper/search.tex
s/mississipi/mississippi/

@@
 \begin{figure}
-\includegraphics[width=\textwidth]{figures/mississipi-nodes}
-\caption{Suffix Tree for the string \putstring{mississipi}}%
-\label{fig:mississipi-nodes}
+\includegraphics[width=\textwidth]{figures/mississippi-nodes}
+\caption{Suffix Tree for the string \putstring{mississippi}}%
+\label{fig:mississippi-nodes}
 \end{figure}
 
-To understand the pseudo-code, it is important to understand that suffix trees are normally implemented as shown in Figure~\ref{fig:mississipi-nodes} where each node contains the start and end indices of its incoming edge.
+To understand the pseudo-code, it is important to understand that suffix trees are normally implemented as shown in Figure~\ref{fig:mississippi-nodes} where each node contains the start and end indices of its incoming edge.
 When presenting the algorithm, we take the liberty of using $s+j$ where $s$ is a string and $j$ an integer to mean the suffix of $s$ starting at position $j$.

### 2005-08-30 paper/construction.tex
fix references
s/mississipi/mississippi/

@@
 
+\begin{figure}
+\includegraphics{merge}
+\caption{Node Merging demonstration}%
+\label{fig:merge}
+\end{figure}
+
 It is very simple to remove these nodes.
-As discussed above, suffix trees are normally implemented such that each node is actually just a pair of indices into the original string as shown in Figure~\ref{fig:mississipi-nodes}.
+As discussed above, suffix trees are normally implemented such that each node is actually just a pair of indices into the original string as shown in Figure~\ref{fig:mississippi-nodes}.
 To merge a child with its parent, we just need to replace both by a node whose starting index is the starting index of the original child minus the length of the parent edge.
-This process is illustrated in Figure~\ref{fig:merge}
+This process is illustrated in Figure~\ref{fig:merge}.
+
+The construction of the tree using either Ukonnen or McCreight's algorithm assures one that this operation is always correct due to the way that the indices are created.
 

### 2005-09-03 paper/space.tex
Finally a correct proof for the space requirements.

--- 2005-08-11 paper/space.tex
@@
 One-error dotted trees have more nodes.
-The worst case presented in Figure~\ref{fig:aaaa} is a string of the form $a^n = aaa\cdots$ which generates a one-dotted tree taking $O(n^2)$ nodes.
-For a larger number of errors $k$, it is easy to see that this leads to $O(n^{k+1})$ nodes.
+The worst case presented in Figure~\ref{fig:aaaa} is a string of the form $a^n = aaa\cdots$ which generates a one-dotted tree taking $\mathcal{O}(n^2)$ nodes.
+For a larger number of errors $k$, it is easy to see that this leads to $\mathcal{O}(n^{k+1})$ nodes.
 
-We can do better, however, by considering the expected case. First we need a pair of definitions:
+We can do better, however, by considering the expected case.
 
-\begin{definition}[weight under a node, weight of a tree]
-The weight under a node $\mathit{Weight}(n)$ is the number of leafs in the subtree rooted at that node.
+\begin{definition}
+We define $N_k$ to be the number of nodes of the k-error dotted tree.
 
-The total weight of a tree is the sum of the weights of all the nodes.
+$N_0$ is the number of nodes of the suffix tree.
 \end{definition}
 
-Note that the weight of a tree \emph{is not the same} as the weight at the root.
+\begin{definition}[Maximum Path Lenght]
+$l$ is the lenght of the deepest node in the tree (which isn't a leaf)
 
-\begin{definition}[height of a node, average height]
-The height of a node $\mathit{Height}(n)$ is the number of nodes in the path leading to that node from the root. The average height of the tree $\hat{h}$, is the average height of its nodes.
-
-\[ \hat{h} = \frac{1}{|\mathcal{T}|} \sum_{x \in \mathit{\mathcal{T}}} \mathit{Height}(x) \]
+\[ l = \max_{\overline{w} \in \mathcal{T}} |w| \mathrm{and \overline{w} isn't a leaf} \]
 \end{definition}
 
-\begin{lemma}
-Given a tree (any tree, not just suffix or similar tree), the sum of the weights of all the nodes in the tree is equal to the average height of the node times the number of nodes.
-\end{lemma}
+\begin{definition}[Suffix-Depth]
+For any node $\overline{w}$ its suffix-depth is the number of nodes which must be followed to reach the root.
+\end{definition}
 
-Consider the tree being built node by node, in any order, as long as the property that it remains a tree during the process is maintained.
-Each node added, adds a certain amount of weight to the tree.
-In particular it adds one for each node that sits above it in the tree.
-This is the same as saying that each node adds its height (which does not change).
-The weight of the final tree is the sum of all these values.
+For a node $\overline{w}$, its suffix-depth is $|w|$. It is obvious that the $l$ is a maximum on the suffix-depth of nodes.
 
-\[ \mathit{Weight}(\mathcal{T}) = \sum_{x \in \mathcal{T}}\mathit{Height}(x) = |\mathcal{T}|\frac{1}{|\mathcal{T}|}\sum_{x \in \mathcal{T}}\mathit{Height}(x) = |\mathcal{T}|\hat{h} \qed \]
+We can now show that $N_k = \mathcal{O}(nl^k)$ by induction.
+It is a known fact that $N_0 = \mathcal{O}(n)$.
+The algorithm for turning a k-error into a (k+1)-error dotted tree, can be looked at the following way~\footnote{We are here considering that the \textit{foreach} loop of Figure~\ref{algo:construct-k} runs in a specific order which is in fact difficult to code for.
+However, as an analysis tool, it is a valid assumption.}: First it makes an almost exact copy of the k-error dotted tree and sets it as the root's error tree and it clears all the other error trees.
+Then it makes a (possibly incomplete copy) broken amongst all the nodes whose suffix links point at the root (nodes at suffix-depth 1).
+It proceeds to the nodes at suffix-depth 2 and on to increasing suffix-depths up to $l$.
+At each such step, the number of nodes is increased by a maximum of $N_k$.
+Therefore, we start with $N_k$ nodes, make an almost full copy, and copy that at most $l$ times.
+We have $N_{k+1}=\mathcal{O}(N_kl)$.
+Assuming $N_k=\mathcal{O}(nl^k)$ by induction we conclude $N_{k+1}=\mathcal{O}(nl^{k+1})$.
 
-We will reason by induction, starting with the case of a zero-error dotted tree to show, that the bound on the number of nodes in an k-error dotted tree can also be expressed as $O(n\hat{h}^k)$ where $\hat{h}$ is the average height of the suffix tree.
-The base case is a suffix tree and is obvious.
-
-The number of nodes in a Patricia tree with $c$ strings ($c$ leafs) is $O(c)$~\cite{patricia} (as a special case, a suffix tree is the Patricia tree of all the suffixes in a string).
-The number of nodes in the error tree of a node is therefore asymptomatically bound by the number of leafs in the subtree at that node times a constant.
-
-%Assume the k-error suffix tree has $O(h\hat{h}^k)$ nodes.
-
-%We define $\mathit{Weight}(\overline{w})$ as above.
-For any node, the number of leafs in its error tree will be $O(\mathit{Weight}(\overline{w}))$.
-The total number of nodes for all the error trees in the dotted tree is therefore equal to $O(\sum_{\overline{w} \in \mathcal{T}}\mathit{Weight}(\overline{w}))$ where $\mathcal{T}$ is the original suffix tree.
-Given the lemma above, $\sum_{\overline{w} \in \mathcal{T}}\mathit{Weight}(\overline{w}) = O(n\hat{h})$.
-
-%The difference between a k-error and a (k+1)-error tree is that all the
-
-So far, we have not really obtained much since, in the worst case, $h=O(n)$.
-However, the average expected case, is $h=O(\log n)$~\cite{devroye:note,szpankowski:unexpected}.
-This allows us to conclude that, in the average case, we will have $O(n\log^k n)$ nodes in our tree.
+So far, we have achieved little since in the worst case $l=n-1$ as shown above.
+However, the expected case is $l=\mathcal{O}(\log n)$~\cite{apostolico92selfalignments} and we have $N_k=\mathcal{O}(n\log^k n)$.

### 2005-09-03 paper/abstract.tex
I like $\mathcal{O}$ for big-O notation

--- 2005-07-28 paper/abstract.tex
@@
 We can later query this index to identify the places where a string occurs up to a certain number of allowed substitutions $k$.
-We present a structure for this indexing which occupies space $O(n\log^kn)$ in the average case, indepedent of alphabet size, $n$ being the text size.
-This structure support searching in $O(m^{k+1})$ time, for patterns of size $m$, again independently of alphabet size.
-The construction of the structure has time bounded by $O(N|\Sigma|)$, where $N$ is the number of nodes in the index and $|\Sigma|$ the alphabet size.
+We present a structure for this indexing which occupies space $\mathcal{O}(n\log^kn)$ in the average case, independent of alphabet size, $n$ being the text size.
+This structure support searching in $\mathcal{O}(m^{k+1})$ time, for patterns of size $m$, again independently of alphabet size.
+The construction of the structure has time bounded by $\mathcal{O}(N|\Sigma|)$, where $N$ is the number of nodes in the index and $|\Sigma|$ the alphabet size.

### 2005-09-03 paper/conclusion.tex
I like $\mathcal{O}$ for big-O notation

@@
-We have presented an indexing structure for approximate text matching which takes $O(n\log^k n)$ space in the average case (this can be seen either in a theoretical or experimentally).
-This structure allows matches in $O(m^{k+1})$ time and can be constructed in $O(N|\Sigma|)$ time ($N$ being the actual number of nodes).
+We have presented an indexing structure for approximate text matching which takes $\mathcal{O}(n\log^k n)$ space in the average case (this can be seen either in a theoretical or experimentally).
+This structure allows matches in $\mathcal{O}(m^{k+1})$ time and can be constructed in $\mathcal{O}(N|\Sigma|)$ time ($N$ being the actual number of nodes).
 The structure and the algorithms to construct it are very simple and should be easily implemented.

### 2005-09-03 paper/space.tex
Better description Spelling fix

@@
 \begin{definition}[Maximum Path Lenght]
-$l$ is the lenght of the deepest node in the tree (which isn't a leaf)
+$l$ is the length of the deepest node in the tree (which isn't a leaf)
 
@@
 However, as an analysis tool, it is a valid assumption.}: First it makes an almost exact copy of the k-error dotted tree and sets it as the root's error tree and it clears all the other error trees.
-Then it makes a (possibly incomplete copy) broken amongst all the nodes whose suffix links point at the root (nodes at suffix-depth 1).
+Then it proceeds in stages, making a (possibly incomplete) copy of this tree spread amongst the nodes at suffix-depth 1.
 It proceeds to the nodes at suffix-depth 2 and on to increasing suffix-depths up to $l$.
@@
 Therefore, we start with $N_k$ nodes, make an almost full copy, and copy that at most $l$ times.
-We have $N_{k+1}=\mathcal{O}(N_kl)$.
+We have $N_{k+1}=\mathcal{O}(N_k(l+1))$.
 Assuming $N_k=\mathcal{O}(nl^k)$ by induction we conclude $N_{k+1}=\mathcal{O}(nl^{k+1})$.

### 2005-09-03 paper/introduction.tex
Make the bibliography correct

@@
-Since their introduction~\cite{weiner}, suffix trees have been one of the methods of choice for text indexing.
+Since their introduction~\cite{DBLP:conf/focs/Weiner73}, suffix trees have been one of the methods of choice for text indexing.
 Suffix trees solve the following problem: given a text which has been preprocessed, and a pattern, efficiently find the places in the text where this pattern occurs.
@@
 
-Our work is related to efforts to index the text for approximate matching such as~\cite{maass:text,mosche}.
-The structure presented is superficially very similar to the one presented in~\cite{inexact} as an \textit{inexact suffix tree} though that work is concerned with solving a different set of problems.
+Our work is related to efforts to index the text for approximate matching such as~\cite{maass:text,amir00oneerror}.
+The structure presented is superficially very similar to the one presented in~\cite{DBLP:journals/tcs/ChattarajP05} as an \textit{inexact suffix tree} though that work is concerned with solving a different set of problems and the algorithms presented are fundamentally different.
 

### 2005-09-03 paper/structure.tex
Make the bibliography correct

@@
 
-Suffix Trees have become a well known structure since their introduction in 1973~\cite{weiner}. In Figure~\ref{fig:mississippi-0} we show the suffix tree for the string \putstring{mississippi}.
+Suffix Trees have become a well known structure since their introduction in 1973~\cite{DBLP:conf/focs/Weiner73}.
+In Figure~\ref{fig:mississippi-0} we show the suffix tree for the string \putstring{mississippi}.
 
@@
 
-The name Patricia is sometimes spelled PATRICIA, since it was formed from the initial of Practical Algorithm to Retrieve Information Coded in Alphanumeric~\cite{morrison:patricia}.
+The name Patricia is sometimes spelled PATRICIA, since it was formed from the initial of Practical Algorithm to Retrieve Information Coded in Alphanumeric~\cite{DBLP:journals/jacm/Morrison68}.
 
@@
 In a suffix tree, it is possible to define a suffix link for each internal node.
-Both Ukkonen's~\cite{ukkonen} and McCreight's~\cite{mccreight} algorithms are linear time algorithms for constructing a suffix tree which includes suffix links.
+Both Ukkonen's~\cite{ukkonen92constructing} and McCreight's~\cite{mccreight} algorithms are linear time algorithms for constructing a suffix tree which includes suffix links.
 The suffix links are shown in Figure~\ref{fig:mississippi-0} as dotted links.

### 2005-09-03 paper/abstract.tex
Make the text better

--- 2005-07-28 paper/abstract.tex
@@
-The problem we are addressing is the one of text indexing for approximate matching.
+The problem we address is text indexing for approximate matching.
 We consider that we are given a text $\mathcal{T}$ which undergoes some preprocessing to generate an index.
 We can later query this index to identify the places where a string occurs up to a certain number of allowed substitutions $k$.
-We present a structure for this indexing which occupies space $O(n\log^kn)$ in the average case, indepedent of alphabet size, $n$ being the text size.
-This structure support searching in $O(m^{k+1})$ time, for patterns of size $m$, again independently of alphabet size.
-The construction of the structure has time bounded by $O(N|\Sigma|)$, where $N$ is the number of nodes in the index and $|\Sigma|$ the alphabet size.
+We present a structure for indexing which occupies space $\mathcal{O}(n\log^kn)$ in the average case, independent of alphabet size, $n$ being the text size.
+This structure support searching in $\mathcal{O}(m^{k+1})$ time, for patterns of size $m$, again independent of alphabet size.
+The construction of the structure has time bound by $\mathcal{O}(N|\Sigma|)$, where $N$ is the number of nodes in the index and $|\Sigma|$ the alphabet size.
+
+\textbf{Keywords:} string algorithms, suffix trees, approximate text matching, text indexing.

### 2005-09-03 paper/structure.tex
Call it Patricia tree Small fixes

@@
 \begin{figure}
-\includegraphics{figures/mississippi-0}
+\includegraphics[width=\textwidth]{figures/mississippi-0}
 \caption{Suffix Tree for the string \putstring{mississippi}}%
@@
 Given a set $\Sigma$, we say that $s$ is a \emph{string over $\Sigma$} if $s$ is a possibly empty sequence of elements of $\Sigma$.
-The elements of the set $\Sigma$ are also called \emph{character}.
+The elements of the set $\Sigma$ are also called \emph{characters}.
 The length of the string, denoted by $|s|$ is the number of elements it contains.
@@
 
-If $s = wxy$, then $w$ is a \emph{prefix} of $s$, $x$ is a \emph{substring} of $s$ (at position $s$) and $y$ is a \emph{suffix} of $s$ (at position $p$).
+If $s = wxy$, then $w$ is a \emph{prefix} of $s$, $x$ is a \emph{substring} of $s$ (at position $|w|$) and $y$ is a \emph{suffix} of $s$ (at position $|wx|$).
 \end{definition}
@@
 \begin{definition}[Suffix Tree]
-A \emph{suffix tree} for a string $S$ is a $\Sigma^+$ whose leaf nodes (those without children) have paths corresponding to suffixes of the string $S\$ $.
+A \emph{suffix tree} for a string $S$ is a Patricia tree whose leaf nodes (those without children) have paths corresponding to all suffixes of the string $S\$ $.
 \end{definition}
@@
 \begin{definition}[Occurence Set]
-Given a node $\overline{w}$ in a suffix tree, we call it's \emph{occurrence set} the set of indexes in the original string where the string $w$ occurs.
+Given a node $\overline{w}$ in a suffix tree, we call its \emph{occurrence set} the set of indexes in the original string where the string $w$ occurs.
 \end{definition}
@@
 \begin{definition}[Position Set]
-Given a node $\overline{w}$ in a suffix tree, it's \emph{position set} is the set formed by taking it's occurrence set and adding the length of $w$ to each element.
+Given a node $\overline{w}$ in a suffix tree, its \emph{position set} is the set formed by taking its occurrence set and adding the length of $w$ to each element.
 \end{definition}
@@
 \begin{definition}[Error Tree]
-For any node $w$ its error tree is the $\Sigma^+$ tree formed by taking it's position set, adding one to each element and forming the $\Sigma^+$ tree of the suffixes of the suffixes starting at those positions.
+For any node $w$ its error tree is the Patricia tree formed by taking its position set, adding one to each element and forming the Patricia tree of the suffixes of the suffixes starting at those positions.
 If the position set includes the end of the string, that element is removed.

### 2005-09-03 paper/introduction.tex
Remove "structure" section

@@
 The structure presented is superficially very similar to the one presented in~\cite{DBLP:journals/tcs/ChattarajP05} as an \textit{inexact suffix tree} though that work is concerned with solving a different set of problems and the algorithms presented are fundamentally different.
-
-\subsection{Structure of the Paper}
-
-In section~\ref{sec:structure} we present the structure and the way to use it for searching.
-In section~\ref{sec:build-1} we show the way to build a structure to handle one error only and analyze this version of the algorithm.
-Then, we show, in section~\ref{sec:build-k} how to generalize this procedure for any number of errors and how the previous analysis also applies to any number of errors.

### 2005-09-03 paper/search.tex
Use \mathcal{O} for big-O Specify that reporting can be done in linear time

@@
 The recursive call at line~\ref{algoline:search:followDot} means that a dot link was followed.
-There are at most $O(m^k)$ places where a dot link can be followed while descending a string~\footnote{For $k=1$ each point in the string.
+There are at most $\mathcal{O}(m^k)$ places where a dot link can be followed while descending a string~\footnote{For $k=1$ each point in the string.
 For $k>1$, each path for $k-1$, can now have at most $m$ extra branches.}.
-Following one path in the tree takes $O(m)$ time.
-Therefore, the total time is $O(m^{k+1})$.
+Following one path in the tree takes $\mathcal{O}(m)$ time.
+Therefore, the total time is $\mathcal{O}(m^{k+1})$ without the reporting of occurences (which can be done in time linear to the number of occurences).

### 2005-09-03 paper/construction.tex
I like $\mathcal{O}$ for big-O notation
Small corrections

@@
 
-For any other node $\overline{aw}$, the error tree is a filtered copy of the error tree at node $\overline{w}$ (the node pointed to by node $\overline{aw}$'s suffix link.
+For any other node $\overline{aw}$, the error tree is a filtered copy of the error tree at node $\overline{w}$ (the node pointed to by node $\overline{aw}$'s suffix link).
 
@@
 \item The leaf labeled 1 in the original error tree is not included
-\item A leaf labeled $p$ is only included if $s_{p-1}$ is $a$ (the label of the suffix link)
+\item A leaf labeled $p$ is included only if $s_{p-1}$ equals $a$ (the label of the suffix link)
 \item All leafs in the copy have a label which is the original label's value minus one
@@
 \begin{figure}
-\includegraphics{merge}
+\centering
+\includegraphics[width=.5\textwidth]{merge}
 \caption{Node Merging demonstration}%
@@
 
-If the number of nodes in the final tree is $N$ and the alphabet $\Sigma$, then the above algorithm runs in time $O(N|\Sigma|)$.
+If the number of nodes in the final tree is $N$ and the alphabet $\Sigma$, then the above algorithm runs in time $\mathcal{O}(N|\Sigma|)$.
 
@@
 
-To construct the (k+1)-error tree from the k-error tree, make a copy of the tree as above (adjusting leafs and filtering the leafs labeled one) and make this the new root error tree.
+To construct the (k+1)-error tree from the k-error tree, make an adjusted copy of the tree as above (adjusting leafs and filtering the leafs labeled one) and make this the new root error tree.
 Then, for every other node, remove the current error tree (it has a level too few).
@@
 Using this method only leafs labeled one need to be removed (as well as any internal node which becomes unnecessary).
+
+The analysis for the time cost above remains valid.
+Taking $N_i$ to mean the number of nodes of the i-error tree, we now have that the time cost is $\mathcal{O}(N_1|\Sigma|+N_2|\Sigma|+\dots+N_k|\Sigma)=\mathcal{O}((N_1+N_2+\dots+N_k)|\Sigma|)=\mathcal{O}(N_k|\Sigma|)=\mathcal{O}(N|\Sigma|)$ since $N_k \le N_i$ ($N_k = N$).

### 2005-09-05 src/stats.cpp
If you have nothing to say, keep quiet

@@ -56,10 +56,12 @@ void stats::print() {
 }
 
 void stats::print( std::ostream& out ) {
-	out << "Total matches: "<< matches.total
-		<< ", of which " << matches.positive
-		<< " ( " << double( matches.positive ) / matches.total
-		<< " ) were positive\n";
+	if ( matches.total ) {
+		out << "Total matches: "<< matches.total
+			<< ", of which " << matches.positive
+			<< " ( " << double( matches.positive ) / matches.total
+			<< " ) were positive\n";
+	}
 	for ( std::map< std::string, proportioninfo>::const_iterator f = proportions.begin(), e = proportions.end();
 			f != e;
 			++f ) {

### 2005-09-05 src/timer.cpp
more precise

@@ -35,7 +35,7 @@ void Timer::start() {
 	struct tms tmp;
 	if ( times( &tmp ) != clock_t( -1 ) ) {
 		begin = tmp.tms_utime;
-		if ( !Clk_Tck ) Clk_Tck = sysconf( _SC_CLK_TCK );
+		if ( !Clk_Tck ) Clk_Tck = sysconf( _SC_CLK_TCK )/100;
 	} else {
 		std::cerr << "Error: " << strerror( errno ) << '\n';
 		exit( 1 );
@@ -106,7 +106,7 @@ void CummulativeTimer::start() {
 	struct tms tmp;
 	times( &tmp );
 	begin = tmp.tms_utime;
-	if ( !Clk_Tck ) Clk_Tck = sysconf( _SC_CLK_TCK );
+	if ( !Clk_Tck ) Clk_Tck = sysconf( _SC_CLK_TCK )/100;
 }
 
 

### 2005-09-05 src/main.cpp
More stats

@@ -1,5 +1,6 @@
 #include "suffix_tree.h"
 #include "timer.h"
+#include "stats.h"
 #include <stdio.h>
 #include <stdlib.h>
 #include <string.h>
@@ -22,10 +23,12 @@ void PrintUsage()
 	exit(0);
 }
 
-void find_substring( SUFFIX_TREE* tree, const char* string ) {
+void find_substring( SUFFIX_TREE* tree, const char* string, CummulativeTimer* m = 0 ) {
 	Timer match( "match" );
+	if ( m ) m->start();
 	unsigned i = ST_FindSubstringWithErrors(tree, string, strlen(string));
 	match.stop();
+	if ( m ) m->stop();
 
 	if(i == ST_ERROR)
 		printf("\nResults: String is not a substring.\n\n");
@@ -79,11 +82,19 @@ int main(int argc, char* argv[])
 		find_substring( tree, argv[ 3 ] );
 	} else {
 		std::string tmp;
-		while ( std::getline( std::cin, tmp ) ) find_substring( tree, tmp.c_str() );
+		CummulativeTimer* cummul = 0;
+		while ( std::getline( std::cin, tmp ) ) {
+			if ( tmp.substr( 0, 5 ) == "timer" ) {
+				delete cummul;
+				cummul = new CummulativeTimer( tmp.substr( 6,std::string::npos ).c_str() );
+			} else find_substring( tree, tmp.c_str(), cummul );
+		}
 	}
 	std::cout << boost::format( "String size: %s \n" ) % ( tree->length - 1 );
 	std::cout << boost::format( "Nodes without dot links: %s \n" ) % ST_CountNodes( tree, false );
 	std::cout << boost::format( "Nodes with dot links: %s \n" ) % ST_CountNodes( tree, true );
+
+	stats::print();
 
 	free(str);
 	ST_DeleteTree(tree);

### 2005-09-16 paper/abstract.tex
Many small changes and it's almost done

@@
 This structure support searching in $\mathcal{O}(m^{k+1})$ time, for patterns of size $m$, again independent of alphabet size.
-The construction of the structure has time bound by $\mathcal{O}(N|\Sigma|)$, where $N$ is the number of nodes in the index and $|\Sigma|$ the alphabet size.
+The construction of the structure has time bound by $\mathcal{O}(kN|\Sigma|)$, where $N$ is the number of nodes in the index and $|\Sigma|$ the alphabet size.
 

### 2005-09-16 paper/conclusion.tex
Many small changes and it's almost done

@@
 We have presented an indexing structure for approximate text matching which takes $\mathcal{O}(n\log^k n)$ space in the average case (this can be seen either in a theoretical or experimentally).
-This structure allows matches in $\mathcal{O}(m^{k+1})$ time and can be constructed in $\mathcal{O}(N|\Sigma|)$ time ($N$ being the actual number of nodes).
+This structure allows matches in $\mathcal{O}(m^{k+1})$ time and can be constructed in $\mathcal{O}(kN|\Sigma|)$ time ($N$ being the actual number of nodes).
 The structure and the algorithms to construct it are very simple and should be easily implemented.
@@
 
-\subsection{Acknowledgements}
+\subsection{Acknowledgments}
 
-We thank Y for making their suffix tree implementation available~\cite{} over which we implemented our algorithms.
+We thank Lu\'{\i}s~Russo and Sara~Madeira for several productive discussions.

### 2005-09-16 paper/construction.tex
Many small changes and it's almost done

@@
 
-The construction of the tree using either Ukonnen or McCreight's algorithm assures one that this operation is always correct due to the way that the indices are created.
+The construction of the tree using either Ukonnen's or McCreight's algorithm assures one that this operation is always correct due to the way that the indices are created.
 
@@
 
-\subsection{Extension to any number of errors}
+\subsection{Extension to any number of errors}\label{subsec:construct-k-larger-1}
 
@@
 
-We note that the method of removing the existing error trees can be improved in practice (the asymptotic bound remains the same) the following way: While making a copy of the tree for construction of the root's error tree, instead of following a dot link to make a copy, move the error tree to its destination and make the necessary adjustments inplace (this move will probably be just an adjustment of pointer depending on the exact implementation).
+We note that the method of removing the existing error trees can be improved in practice (the asymptotic bound remains the same) the following way: While making a copy of the tree for construction of the root's error tree, instead of following a dot link to make a copy, move the error tree to its destination and make the necessary adjustments in place (this move will probably be just an adjustment of pointer depending on the exact implementation).
 Using this method only leafs labeled one need to be removed (as well as any internal node which becomes unnecessary).
 
+\begin{definition}
+We define $N_k$ to be the number of nodes of the k-error dotted tree.
+
+$N_0$ is the number of nodes of the suffix tree.
+
+For a known $k$, we set $N=N_k$.
+\end{definition}
+
 The analysis for the time cost above remains valid.
-Taking $N_i$ to mean the number of nodes of the i-error tree, we now have that the time cost is $\mathcal{O}(N_1|\Sigma|+N_2|\Sigma|+\dots+N_k|\Sigma)=\mathcal{O}((N_1+N_2+\dots+N_k)|\Sigma|)=\mathcal{O}(N_k|\Sigma|)=\mathcal{O}(N|\Sigma|)$ since $N_k \le N_i$ ($N_k = N$).
+We now have that the time cost is $\mathcal{O}(N_1|\Sigma|+N_2|\Sigma|+\dots+N_k|\Sigma)=\mathcal{O}((N_1+N_2+\dots+N_k)|\Sigma|)=\mathcal{O}(kN|\Sigma|)$.

### 2005-09-16 paper/experimental.tex
Many small changes and it's almost done

@@
 
-The code was implemented in the C~Language, starting from the suffix tree implementation by FIXME.
+The code was implemented in the C~Language, starting from an open-source implementation~\cite{ansicimplementation}.
 
@@
 \begin{enumerate}
-\item ``Democracy in America'' by Alexis de Tocqueville as natural language example
-\item The DNA of S. Cervesiae (first chromossome)
-\item Randomly generated text
+\item The English translation of ``Democracy in America'' by Alexis de Tocqueville as natural language example
+\item The DNA of S. Cervesiae (first chromosome)
+\item Randomly generated text by a uniform distribution
 \end{enumerate}
@@
 \begin{figure}%
-\includegraphics{ratio-1}%
-\includegraphics{ratio-2}%
-\includegraphics{ratio-3}%
+\centering
+\subfigure[English Text]{
+\includegraphics[width=.25\textwidth]{ratio-in=english-0-1.eps}%
+\includegraphics[width=.25\textwidth]{ratio-in=english-1-2.eps}%
+\includegraphics[width=.25\textwidth]{ratio-in=english-2-3.eps}%
+}
+\\
+\subfigure[DNA]{
+\includegraphics[width=.25\textwidth]{ratio-in=dna-0-1.eps}%
+\includegraphics[width=.25\textwidth]{ratio-in=dna-0-1.eps}%
+\includegraphics[width=.25\textwidth]{ratio-in=dna-1-2.eps}%
+}
+\\
+\subfigure[Random Text]{
+\includegraphics[width=.25\textwidth]{ratio-in=random-0-1.eps}%
+\includegraphics[width=.25\textwidth]{ratio-in=random-0-1.eps}%
+\includegraphics[width=.25\textwidth]{ratio-in=random-1-2.eps}%
+}
 \caption{Ratios of k-error and (k+1)-error dotted trees}\label{fig:ratios}%
@@
 
-To experimentally verify the average case, we show in Figure~\ref{fig:ratios} the ratios between the k-error and the (k+1)-error dotted trees.
-We can easily see that for all three data sets, the experimental values do resemble a logarithm.
+To experimentally verify the average case, we show in Figures~\ref{fig:ratios} the ratios between the k-error and the (k+1)-error dotted trees.
+We can easily see that for all three data sets, the experimental values do resemble a logarithm as predicted.
 
@@
 
-For searching, the full textes were first indexed and then online searches were performed on top of the structure.
+For searching, the full texts were first indexed and then online searches were performed on top of the structure.
 To generate input patterns, randomly chosen bits of the text were taken.
@@
 To generate mismatches, one character too many was altered~\footnote{Since the characters used were not present in the original text, there is no risk of generating a string which \emph{does appear} in the text.}.
-The search algorithm performed an early exit, without reporting of occurences (ie, it only reported whether the string exists in the text).
-Therefore, the number of occurences had no influence on the search time.
+The search algorithm performed an early exit, without reporting of occurrences (ie, it only reported whether the string exists in the text).
+Therefore, the number of occurrences had no influence on the search time.
+
+\begin{figure}
+\includegraphics[width=\textwidth]{search-time-vary-m.k=2.in=english-100k.eps}
+\caption{Search time versus pattern size}\label{fig:search-vary-m}
+\end{figure}
+
+Figure~\ref{fig:search-vary-m} shows the search time for growing pattern lenghts.
+It shows the case where the pattern exists and where it does not (an extra error is added).
+The fact that the search is done using an early-exit strategy explains why non existing patterns take longer than existing patterns.
+The value presented is the number of character comparisons the algorithm makes.

### 2005-09-16 paper/search.tex
Many small changes and it's almost done

@@
 
-To understand the pseudo-code, it is important to understand that suffix trees are normally implemented as shown in Figure~\ref{fig:mississippi-nodes} where each node contains the start and end indices of its incoming edge.
-When presenting the algorithm, we take the liberty of using $s+j$ where $s$ is a string and $j$ an integer to mean the suffix of $s$ starting at position $j$.
+To understand the pseudo-code, it is important to understand that suffix trees are normally implemented so that each node tracks the string on the edge leading to it by keeping the start and end indices on the original string (all strings on the trees presented here are substrings of the original one).
+On Figure~\ref{fig:mississippi-nodes} we show an example of this structure (for the now familiar \putstring{mississippi}) where the string each node references is shown explicitly.
+
+When presenting the algorithm, we take the liberty of using $s+j$ where $s$ is a string and $j$ an integer to mean the suffix of $s$ starting at position $j$ as happens in the C~Programming Language.
 
@@
 
-This algorithm runs in time $O(m^{k+1})$:
+This algorithm runs in time $\mathcal{O}(m^{k+1})$:
 
@@
 Following one path in the tree takes $\mathcal{O}(m)$ time.
-Therefore, the total time is $\mathcal{O}(m^{k+1})$ without the reporting of occurences (which can be done in time linear to the number of occurences).
+Therefore, the total time is $\mathcal{O}(m^{k+1})$ without the reporting of occurrences (which can be done in time linear to the number of occurrences).

### 2005-09-16 paper/space.tex
Make the aaaa figure smaller
Many small changes and it's almost done

@@
 \begin{figure}
-\includegraphics{figures/aaaa}
+\centering
+\includegraphics[height=.2\textheight]{figures/aaaa}
 \caption{$a^n$: worst case space consumption}%
@@
 One-error dotted trees have more nodes.
-The worst case presented in Figure~\ref{fig:aaaa} is a string of the form $a^n = aaa\cdots$ which generates a one-dotted tree taking $\mathcal{O}(n^2)$ nodes.
+The worst case presented in Figure~\ref{fig:aaaa} is a string of the form $a^n = aaa\dots$ which generates a one-dotted tree taking $\mathcal{O}(n^2)$ nodes.
 For a larger number of errors $k$, it is easy to see that this leads to $\mathcal{O}(n^{k+1})$ nodes.
@@
 
-\begin{definition}
-We define $N_k$ to be the number of nodes of the k-error dotted tree.
-
-$N_0$ is the number of nodes of the suffix tree.
-\end{definition}
-
 \begin{definition}[Maximum Path Lenght]
@@
 
-\[ l = \max_{\overline{w} \in \mathcal{T}} |w| \mathrm{and \overline{w} isn't a leaf} \]
+\[ l = \max_{\overline{w} \in \mathcal{T}} |w| \mbox{ and $\overline{w}$ isn't a leaf} \]
 \end{definition}
@@
 So far, we have achieved little since in the worst case $l=n-1$ as shown above.
-However, the expected case is $l=\mathcal{O}(\log n)$~\cite{apostolico92selfalignments} and we have $N_k=\mathcal{O}(n\log^k n)$.
+However, the expected case is $l=\mathcal{O}(\log n)$~\cite{apostolico92selfalignments,szpankowski:unexpected} and we have $N_k=\mathcal{O}(n\log^k n)$.

### 2005-09-16 paper/conclusion.tex
Corrections by aml, the easy ones

@@
-We have presented an indexing structure for approximate text matching which takes $\mathcal{O}(n\log^k n)$ space in the average case (this can be seen either in a theoretical or experimentally).
+We have presented an indexing structure for approximate text matching which takes $\mathcal{O}(n\log^k n)$ space in the average case.
+This complexity was analyzed theoretically and observed experimentally.
 This structure allows matches in $\mathcal{O}(m^{k+1})$ time and can be constructed in $\mathcal{O}(kN|\Sigma|)$ time ($N$ being the actual number of nodes).

### 2005-09-16 paper/construction.tex
Corrections by aml, the easy ones

@@
 The \textit{foreach} loop at line~\ref{algoline:construct-1:foreach} can be implemented using Depth First Search on the tree.
-In fact, the procedure addDotLink only needs to be called for certain nodes (those who do not posses an incoming suffix link).
+In fact, the procedure addDotLink only needs to be called for certain nodes (those which do not possess an incoming suffix link).
 However, there is no readily available way to access only these nodes and the procedure above does little extra work.
@@
 
-The root's error tree is created in time proportional to the number of nodes it contains.
+The error tree at the root is created in time proportional to the number of nodes it contains.
 Every other error tree is constructed by looking at an existing one and copying it.
@@
 
-We note that the method of removing the existing error trees can be improved in practice (the asymptotic bound remains the same) the following way: While making a copy of the tree for construction of the root's error tree, instead of following a dot link to make a copy, move the error tree to its destination and make the necessary adjustments in place (this move will probably be just an adjustment of pointer depending on the exact implementation).
+We note that the method of removing the existing error trees can be improved in practice (the asymptotic bound remains the same) the following way: While making a copy of the tree for construction of the error tree at the root, instead of following a dot link to make a copy, move the error tree to its destination and make the necessary adjustments in place (this move will probably be just an adjustment of pointer depending on the exact implementation).
 Using this method only leafs labeled one need to be removed (as well as any internal node which becomes unnecessary).

### 2005-09-16 paper/introduction.tex
Corrections by aml, the easy ones

@@
 Since their introduction~\cite{DBLP:conf/focs/Weiner73}, suffix trees have been one of the methods of choice for text indexing.
-Suffix trees solve the following problem: given a text which has been preprocessed, and a pattern, efficiently find the places in the text where this pattern occurs.
+Suffix trees can be used solve the following problem: given a text which has been preprocessed, and a pattern, efficiently find the places in the text where this pattern occurs.
 However, for many real-life problems this is too restrictive.
@@
 In this work we present an approach which solves this problem based on an extension of suffix trees.
-The main advantage of this approach is that both the search and the index size are \emph{alphabet independent}, although the indexing time is not.
+The main advantage of this approach is that both the search and the index size are \emph{alphabet independent} (although the indexing time is not).
 
@@
 
-Our work is related to efforts to index the text for approximate matching such as~\cite{maass:text,amir00oneerror}.
-The structure presented is superficially very similar to the one presented in~\cite{DBLP:journals/tcs/ChattarajP05} as an \textit{inexact suffix tree} though that work is concerned with solving a different set of problems and the algorithms presented are fundamentally different.
+Our work is related to efforts to index the text for approximate matching~\cite{maass:text,amir00oneerror}.
+The structure presented is superficially very similar to the one presented in~\cite{DBLP:journals/tcs/ChattarajP05} as an \textit{inexact suffix tree}, although that work is concerned with solving a different set of problems and the algorithms presented are fundamentally different.

### 2005-09-16 paper/search.tex
Corrections by aml, the easy ones

@@
 
-When presenting the algorithm, we take the liberty of using $s+j$ where $s$ is a string and $j$ an integer to mean the suffix of $s$ starting at position $j$ as happens in the C~Programming Language.
+When presenting the algorithm, we take the liberty of using $s+j$ where $s$ is a string and $j$ an integer to mean the suffix of $s$ starting at position $j$ (as happens in the C~Programming Language).
 

### 2005-09-16 paper/space.tex
Corrections by aml, the easy ones

@@
 
-For a node $\overline{w}$, its suffix-depth is $|w|$. It is obvious that the $l$ is a maximum on the suffix-depth of nodes.
+For a node $\overline{w}$, its suffix-depth is $|w|$. It is obvious that $l$ is a maximum on the suffix-depth of nodes.
 
@@
 It is a known fact that $N_0 = \mathcal{O}(n)$.
-The algorithm for turning a k-error into a (k+1)-error dotted tree, can be looked at the following way~\footnote{We are here considering that the \textit{foreach} loop of Figure~\ref{algo:construct-k} runs in a specific order which is in fact difficult to code for.
+The algorithm for turning a k-error into a (k+1)-error dotted tree, can be looked at the following way~\footnote{We are here considering that the second \textit{foreach} loop of Figure~\ref{algo:construct-k} runs in a specific order which is in fact difficult to code for.
 However, as an analysis tool, it is a valid assumption.}: First it makes an almost exact copy of the k-error dotted tree and sets it as the root's error tree and it clears all the other error trees.

### 2005-09-16 paper/structure.tex
Corrections by aml, the easy ones

@@
 Suffix Trees have become a well known structure since their introduction in 1973~\cite{DBLP:conf/focs/Weiner73}.
-In Figure~\ref{fig:mississippi-0} we show the suffix tree for the string \putstring{mississippi}.
+The example in Figure~\ref{fig:mississippi-0} shows the suffix tree for the string \putstring{mississippi}.
 
-To extend the suffix tree, we consider that a node in the suffix tree corresponds to several points in the string at the same time (those points where the string on the path to the node occurs).
+To extend the suffix tree, we will consider that a node in the suffix tree corresponds to several points in the string at the same time (those points where the string on the path to the node occurs).
 We now add to each node an extra edge which corresponds to moving one character forward in all these points.
@@
 \begin{definition}[Character, string]
-Given a set $\Sigma$, we say that $s$ is a \emph{string over $\Sigma$} if $s$ is a possibly empty sequence of elements of $\Sigma$.
+Given a set $\Sigma$, we say that $s$ is a \emph{string over $\Sigma$} if $s$ is a (possibly empty) sequence of elements of $\Sigma$.
 The elements of the set $\Sigma$ are also called \emph{characters}.
@@
 
-The set of all string is denoted by $\Sigma^*$ and $\Sigma^+=\Sigma^*-\{\epsilon\}$.
+The set of all strings is denoted by $\Sigma^*$ and $\Sigma^+=\Sigma^*-\{\epsilon\}$.
 \end{definition}
 
-For denoting character we shall use letters from the beginning of the roman alphabet ($a$, $b$, $c$,\ldots) and for strings, we shall use letters from the end of the alphabet ($w$, $z$, \ldots).
+For denoting characters we shall use letters from the beginning of the roman alphabet ($a$, $b$, $c$,\ldots) and for strings, we shall use letters from the end of the alphabet ($w$, $z$, \ldots).
 
 \begin{definition}
-$wx$ will denote the string formed by all the character of $w$ followed by all the characters of $x$ (the usual concatenation operation).
-$aw$ ($wa$) shall the denote the strings by the character $a$ before (after) the characters of $w$.
+$wx$ will denote the string formed by all the characters of $w$ followed by all the characters of $x$ (the usual concatenation operation).
+$aw$ ($wa$) shall denote the string formed by the character $a$ before (after) the characters of $w$.
 
@@
 \begin{definition}[Patricia tree]
-$T$ is a Patricia tree if $T$ is a rooted tree with edge labels from $\Sigma^+$. For each $a \in \Sigma$ and every node $n$ in $T$, there is at most an edge leaving $n$ whose label is $aw$.
+$T$ is a Patricia tree if $T$ is a rooted tree with edge labels from $\Sigma^+$. For each $a \in \Sigma$ and every node $n$ in $T$, there is at most one edge leaving $n$ whose label is $aw$.
 
-Each node in a Patricia tree has a path leading to it which forms a string. If the node $n$ has the leading path $w$, we shall also refer to $n$ as $\overline{w}$.
+Each node in a Patricia tree has a path leading to it which forms a string. If the node $n$ has the leading path $w$, we shall also refer to $n$ as $\overline{w}$one .
 \end{definition}
 
-The name Patricia is sometimes spelled PATRICIA, since it was formed from the initial of Practical Algorithm to Retrieve Information Coded in Alphanumeric~\cite{DBLP:journals/jacm/Morrison68}.
-
+%The name Patricia is sometimes spelled PATRICIA, since it was formed from the initial of Practical Algorithm to Retrieve Information Coded in Alphanumeric~\cite{DBLP:journals/jacm/Morrison68}.
+%
 In what follows we will assume that there are two symbols (\$ and $.$) which are not part of $\Sigma$.
@@
 
-For example, in the node $\overline{issi}$ in the suffix tree of Figure~\ref{fig:mississippi-0}, the occurrence set is $\{5, 2\}$ and its position set is $\{6, 9\}$.
+For example, in the node $\overline{issi}$ in the suffix tree of Figure~\ref{fig:mississippi-0}, the occurrence set is $\{2, 5\}$ and its position set is $\{6, 9\}$.
 In a sense, one can say that being at node $\overline{issi}$ is being at positions 6 and~9 simultaneously.
@@
 \begin{lemma}[Position set at the suffix node]\label{lemma:suffix-error}
-Given two nodes $\overline{aw}$ and $\overline{w}$ (connected by a suffix link, labeled $a$), if one takes position set of $\overline{w}$, subtracts one to each element, one obtains a superset of the position set of $\overline{w}$.
+Given two nodes $\overline{aw}$ and $\overline{w}$ (connected by a suffix link, labeled $a$), if one takes position set of $\overline{w}$, subtracts one from each element, resulting a superset of the position set of $\overline{w}$.
 The items shared by both sets are those positions of the string which contain an $a$.
@@
 
-For example, in the above mentioned node $\overline{issi}$, the error tree is formed by taking the strings starting at positions $\{7, 10\}$ (ie \putstring{sippi\$} and \putstring{i\$}) in a Patricia tree.
+For example, in the above mentioned node $\overline{issi}$, the error tree is formed by taking the strings starting at positions $\{7, 10\}$ (ie \putstring{sippi\$} and \putstring{pi\$}) in a Patricia tree.
 

### 2005-09-17 scripts/get-fileportion.py
Make the python script do the repeats itself

@@ -1,5 +1,6 @@
 #!/usr/bin/env python
 from sys import argv
+import sys
 import random
 import os
 
@@ -9,23 +10,30 @@ filelen=os.stat(inputfile)[6]
 lenght=int(argv[3])
 k=int(argv[4])
 
-if argv[2] == '.':
-	begin = random.randrange(filelen-lenght)
-else:
-	begin=int(argv[2])
-	if begin + lenght > filelen:
-		raise Exception('File not large enough')
-
 f=open(inputfile,'r')
-f.seek(begin)
-string = f.read(lenght)
+repeats=1
+if len(argv) > 5:
+	repeats=int(argv[5])
 
-changed = []
-for point in random.sample(range(0,len(string)),k):
-	# '#' is a character which does not show up in any of our data files
-	if point > 0:
-		string = string[0:point-1] + '#' + string[point:]
+def get_string(f,begin,lenght,k):
+	f.seek(begin)
+	string = f.read(lenght)
+	for point in random.sample(range(0,len(string)),k):
+		# '#' is a character which does not show up in any of our data files
+		if point > 0:
+			string = string[0:point-1] + '#' + string[point:]
+		else:
+			string = '#' + string[1:]
+
+	print string
+
+
+for i in range(repeats):
+	if argv[2] == '.':
+		begin = random.randrange(filelen-lenght)
 	else:
-		string = '#' + string[1:]
+		begin=int(argv[2])
+		if begin + lenght > filelen:
+			raise Exception('File not large enough')
+	get_string(f,begin,lenght,k)
 
-print string

### 2005-09-17 src/timer.h
One should answer when asked for a name

@@ -23,6 +23,7 @@ class CummulativeTimer {
 		~CummulativeTimer();
 		void start();
 		void stop();
+		const char* name() const { return id.c_str(); }
 
 	private:
 		const std::string id;

### 2005-09-17 src/main.cpp
Measure time spent on searching Fix an important bug while reading from a file

@@ -5,6 +5,7 @@
 #include <stdlib.h>
 #include <string.h>
 #include <iostream>
+#include <exception>
 #include <boost/format.hpp>
 
 /**************************************************************************************************/
@@ -26,12 +27,13 @@ void PrintUsage()
 void find_substring( SUFFIX_TREE* tree, const char* string, CummulativeTimer* m = 0 ) {
 	Timer match( "match" );
 	if ( m ) m->start();
-	unsigned i = ST_FindSubstringWithErrors(tree, string, strlen(string));
+	unsigned i = ST_FindSubstringWithErrors(tree, string, strlen(string),
+			( m ? m->name(): static_cast<const char*>( 0 ) ) );
 	match.stop();
 	if ( m ) m->stop();
 
 	if(i == ST_ERROR)
-		printf("\nResults: String is not a substring.\n\n");
+		printf("\nResults: String[-%s-] is not a substring.\n\n", string);
 	else
 		std::cout << boost::format( "\nResults:   Substring exists in position %s.\n\n" ) % i;
 }
@@ -50,7 +52,7 @@ char* read_file( const char* fname )
 	fseek(file, 0, SEEK_END);
 	int len = ftell(file);
 	fseek(file, 0, SEEK_SET);
-	res = (char*)malloc(len);
+	res = (char*)malloc(len+1);
 	if ( !res ) 
 	{
 		printf("\nOut of memory.\n");
@@ -58,6 +60,7 @@ char* read_file( const char* fname )
 	}
 	fread(res, 1, len, file);
 	fclose( file );
+	res[len] = '\0';
 	return res;
 }
 
@@ -68,35 +71,47 @@ int main(int argc, char* argv[])
 	int k = atoi( argv[ 1 ] );
 	char* str = read_file( argv[ 2 ] );
 	
-	Timer full( "full-tree-construction" );
-	Timer dots( "add-dot-links" );
-	full.start();
-	tree = ST_CreateTree(str);
-	dots.start();
-	ST_AddDotLinks( tree, k );
-	full.stop();
-	dots.stop();
-	
-	
-	if ( strcmp( argv[ 3 ], "-" ) ) {
-		find_substring( tree, argv[ 3 ] );
-	} else {
-		std::string tmp;
-		CummulativeTimer* cummul = 0;
-		while ( std::getline( std::cin, tmp ) ) {
-			if ( tmp.substr( 0, 5 ) == "timer" ) {
-				delete cummul;
[…]

### 2005-09-17 src/suffix_tree.cpp

@@ -591,11 +591,11 @@ DBL_WORD ST_FindSubstring(
 
 
 namespace {
-DBL_WORD find_string( SUFFIX_TREE* tree, NODE* node, const char* W, DBL_WORD P, int d ) {
+DBL_WORD find_string( SUFFIX_TREE* tree, NODE* node, const char* W, DBL_WORD P, int d, const char* acc_name ) {
 	if ( !node ) return ST_ERROR;
 	if ( !P ) return node->path_position;
 	if ( d ) {
-		DBL_WORD follow_dot_link = find_string( tree, node->dot_link, W + 1, P - 1, d - 1 );
+		DBL_WORD follow_dot_link = find_string( tree, node->dot_link, W + 1, P - 1, d - 1, acc_name );
 		if ( follow_dot_link != ST_ERROR ) return follow_dot_link;
 	}
 	node = find_son( tree, node, W[ 0 ] );
@@ -606,10 +606,11 @@ DBL_WORD find_string( SUFFIX_TREE* tree, NODE* node, const char* W, DBL_WORD P, 
 	const unsigned len = std::min( P, get_node_label_length( tree, node ) );
 	while ( d >= 0 ) {
 		while ( j < len && W[ j ] == tree->tree_string[ start + j ] ) {
+			if ( acc_name ) stats::count_one( acc_name );
 			++j;
 		}
 		if ( j == len ) {
-			return find_string( tree, node, W + j, P - j, d );
+			return find_string( tree, node, W + j, P - j, d, acc_name );
 		}
 		--d;
 		++j;
@@ -629,9 +630,10 @@ DBL_WORD ST_FindSubstringWithErrors(
 		/* The substring to find */
 		const char*  W,         
 		/* The length of W */
-		DBL_WORD        P)         
+		DBL_WORD        P,
+		const char* acc_name )         
 {
-	return find_string( tree, tree->root, W, P, tree->k );
+	return find_string( tree, tree->root, W, P, tree->k, acc_name );
 }
 
 

### 2005-09-17 src/suffix_tree.h

@@ -140,7 +140,8 @@ DBL_WORD ST_FindSubstring(SUFFIX_TREE*      tree,   /* The suffix array */
                           DBL_WORD          P);     /* The length of W */
 DBL_WORD ST_FindSubstringWithErrors(SUFFIX_TREE*      tree,   /* The suffix array */
                           const char*    W,      /* The substring to find */
-                          DBL_WORD          P);     /* The length of W */
+                          DBL_WORD          P,     /* The length of W */
+			  const char* acc_name = 0 );
 
 /******************************************************************************/
 /*

### 2005-09-19 paper/conclusion.tex
Win space and other small changes

@@
 
-Along the same lines, the structure's definition might be extended to structures such as the suffix-DAG presented in~\cite[7.7]{gusfield:algorithms}.
+Along the same lines, the structure's definition might be extended to structures such as the suffix-DAG presented in~\cite[\S 7.7]{gusfield:algorithms}.
 An algorithm to efficiently construct it over these suffix-DAGs is another open problem.

### 2005-09-19 paper/construction.tex
Win space and other small changes

@@
 
-Constructing the tree is actually very simple. We will start with a suffix tree which includes suffix links and show how to add error trees to it.
-
-First, we construct the error tree for the root. This tree is almost a copy of the entire tree, except for two properties:
+Constructing the tree is very simple.
+We start with a suffix tree which includes suffix links and add error trees to it.
+First, we construct the error tree for the root.
+This tree is almost a copy of the entire tree, except for two properties:
 
@@
 It is easy to see why this is so.
-The leaf labeled 1 would have resulted in the string \putstring{.s\$} which does not occur in original string (it is one character too large).
+The leaf labeled 1 would have resulted in the pattern \putstring{.s\$} which does not occur in original string (it is one character too large).
 For any other leaf $\overline{w\$}$ which occurred at position $p$ in the string, we have a new leaf $\overline{.w\$}$ which occurs at position $p-1$ in the string.
 
-For any other node $\overline{aw}$, the error tree is a filtered copy of the error tree at node $\overline{w}$ (the node pointed to by node $\overline{aw}$'s suffix link).
+For any other node $\overline{aw}$, the error tree is a copy of the error tree at node $\overline{w}$ (the node pointed to by node $\overline{aw}$'s suffix link) with the following changes:
 
@@
 \item The leaf labeled 1 in the original error tree is not included
-\item A leaf labeled $p$ is included only if $s_{p-1}$ equals $a$ (the label of the suffix link)
-\item All leafs in the copy have a label which is the original label's value minus one
+\item A leaf labeled $p$ is included only if $s_{p-1}$ equals the label of the suffix link
+\item Leafs in the copy have a label which is the original label's value minus one
 \end{enumerate}
 
-These conditions are an expression of Lemma~\ref{lemma:suffix-error}.
-Note that these conditions are very similar to the conditions for the root and both can be implemented as shown in Figure~\ref{algo:copy-subtree}.
+These conditions are an expression of Lemma~\ref{lemma:suffix-error} and a an extension of the conditions for the root. Both are implemented by Algorithm~\ref{algo:copy-subtree}.
 
-\input{copy-subtree-algo}
-
+%\input{copy-subtree-algo}
+%
 The only point to note is line~\ref{algoline:copy-subtree:merge}.
-This is motivated by noting that since we filter some leafs, without a merging procedure, it would be possible to have nodes which possess only one child.
+Since we filter some leafs, without a merging procedure, it would be possible to have nodes which possess only one child.
 These need to be removed.
@@
 \centering
-\includegraphics[width=.5\textwidth]{merge}
+\includegraphics[width=.4\textwidth]{merge}
 \caption{Node Merging demonstration}%
@@
 It is very simple to remove these nodes.
-As discussed above, suffix trees are normally implemented such that each node is actually just a pair of indices into the original string as shown in Figure~\ref{fig:mississippi-nodes}.
+As discussed above, suffix trees are normally implemented such that each node is actually just a pair of indices into the original string (exemplified in Figure~\ref{fig:mississippi-nodes}).
 To merge a child with its parent, we just need to replace both by a node whose starting index is the starting index of the original child minus the length of the parent edge.
@@
 
-The construction of the tree using either Ukonnen's or McCreight's algorithm assures one that this operation is always correct due to the way that the indices are created.
+The construction of the tree using either Ukonnen's or McCreight's algorithm assures that this operation is correct due to the way that the indices are created.
 
@@
 One can construct the error tree for any node if the tree for the node pointed to by the current node's suffix link has an error tree.
-This leads to a recursive definition as presented in Figure~\ref{algo:addDotLink}.
+This leads to a recursive definition implemented as Algorithm~\ref{algo:addDotLink}.
 
@@
 To add dot links at every node, we need to add the dot link at the root and then for every node call the procedure addDotLink.
-This is shown in Figure~\ref{algo:construct-1}.
+This is implemented as Algorithm~\ref{algo:construct-1}.
 The \textit{foreach} loop at line~\ref{algoline:construct-1:foreach} can be implemented using Depth First Search on the tree.
@@
 Each of these copies need time proportional to the original error tree.
-Since there are at maximum $|\Sigma|$ incoming suffix links to a node~\cite{}, each error tree is \emph{looked at} a maximum of $|\Sigma|$ times.
+Since there are at maximum $|\Sigma|$ incoming suffix links to a node, each error tree is \emph{looked at} a maximum of $|\Sigma|$ times.
 The sum of all these operations is therefore bounded by $|\Sigma|N$.
@@
 
-The above algorithm can be used to construct trees with any number of errors by iterating it in the following way:
-
+The above algorithm can be used to construct trees with any number of errors by iterating it.
 To construct the (k+1)-error tree from the k-error tree, make an adjusted copy of the tree as above (adjusting leafs and filtering the leafs labeled one) and make this the new root error tree.
@@
 Finally, for every node except the root, construct its error tree as above (adjusting leafs, filtering the leafs labeled one or which do not correspond to positions in the suffix link's label).
-These error trees will now have the correct number of nodes.
+These error trees will now have the correct strings.
 
+\input{construct-k-algo}
 \input{copy-subtree-with-dot-link-algo}
 
-The code for the copy is shown in Figure~\ref{algo:copy-subtree-with-dot-link} which differs from Figure~\ref{algo:copy-subtree} in that the dot link is also followed while copying the tree.
+The code for the copy is Algorithm~\ref{algo:copy-subtree-with-dot-link} which differs from Algorithm~\ref{algo:copy-subtree} in that dot link are also followed while copying the tree.
 
-\input{construct-k-algo}
-
-The final algorithm for constructing k-error trees is shown in Figure~\ref{algo:construct-k}.
+The final procedure for constructing k-error trees is Algorithm~\ref{algo:construct-k}.
 
@@
 \begin{definition}
-We define $N_k$ to be the number of nodes of the k-error dotted tree.
-
-$N_0$ is the number of nodes of the suffix tree.
-
-For a known $k$, we set $N=N_k$.
+We define $N_k$ to be the number of nodes of the k-error dotted tree. $N_0$ is the number of nodes of the suffix tree. If $k$ is known from context we write $N$ for $N_k$.
 \end{definition}

### 2005-09-19 paper/experimental.tex
Corrections by aml, the easy ones
Win space and other small changes

@@
 
-The code was implemented in the C~Language, starting from an open-source implementation~\cite{ansicimplementation}.
-
-Three data sets were used:
+The code was implemented in the C~Language, starting from an open-source implementation of suffix trees~\cite{ansicimplementation}. Three data sets were used:
 
@@
 \item The English translation of ``Democracy in America'' by Alexis de Tocqueville as natural language example
-\item The DNA of S. Cervesiae (first chromosome)
+\item The DNA of S. Cerevesiae (first chromosome)
 \item Randomly generated text by a uniform distribution
@@
 
-The machine used for testing was an Intel Pentium~IV at 2.40GHz with 4Gb of main memory.
+The machine used for testing was an Intel Pentium~IV running at~2.40GHz with~4GB of main memory.
 
@@
 
-We report the indexing time and size (as a count of the nodes in the structure) as a function of the number of character (by using larger and larger suffixes of the datasets).
+We report the indexing time and size (as a count of the nodes in the structure) as a function of the number of character (by using larger and larger prefixes of the datasets).
 
@@
 \centering
-\subfigure[English Text]{
-\includegraphics[width=.25\textwidth]{ratio-in=english-0-1.eps}%
-\includegraphics[width=.25\textwidth]{ratio-in=english-1-2.eps}%
-\includegraphics[width=.25\textwidth]{ratio-in=english-2-3.eps}%
-}
-\\
-\subfigure[DNA]{
-\includegraphics[width=.25\textwidth]{ratio-in=dna-0-1.eps}%
-\includegraphics[width=.25\textwidth]{ratio-in=dna-0-1.eps}%
-\includegraphics[width=.25\textwidth]{ratio-in=dna-1-2.eps}%
-}
-\\
-\subfigure[Random Text]{
-\includegraphics[width=.25\textwidth]{ratio-in=random-0-1.eps}%
-\includegraphics[width=.25\textwidth]{ratio-in=random-0-1.eps}%
-\includegraphics[width=.25\textwidth]{ratio-in=random-1-2.eps}%
-}
+\includegraphics[width=.7\textwidth]{ratio-in=english-01-12.eps}%
 \caption{Ratios of k-error and (k+1)-error dotted trees}\label{fig:ratios}%
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{search-time-vary-m.k=2.in=english-100k.eps}
+\centering
+\includegraphics[width=.7\textwidth]{search-vary-N.in=all.eps}
 \caption{Search time versus pattern size}\label{fig:search-vary-m}

### 2005-09-19 paper/search.tex
Win space and other small changes

@@
 If we have a mismatch in the middle of an edge, we continue, decreasing the amount of errors permitted.
+
+Before introducing the pseudo-code, it is important to understand that suffix trees are normally implemented so that each node consists of only two indices into the original string which are the start and end position of its incoming edge.
+Figure~\ref{fig:mississippi-nodes} shows an example of this structure (for the now familiar \putstring{mississippi}).
 
@@
 
-To understand the pseudo-code, it is important to understand that suffix trees are normally implemented so that each node tracks the string on the edge leading to it by keeping the start and end indices on the original string (all strings on the trees presented here are substrings of the original one).
-On Figure~\ref{fig:mississippi-nodes} we show an example of this structure (for the now familiar \putstring{mississippi}) where the string each node references is shown explicitly.
-
 When presenting the algorithm, we take the liberty of using $s+j$ where $s$ is a string and $j$ an integer to mean the suffix of $s$ starting at position $j$ (as happens in the C~Programming Language).
@@
 The search algorithm is presented in Figure~\ref{algo:search}.
-Reporting of the leafs below a certain node can be done by a simple Depth First Search reporting any leafs found.
-If one is interested in knowing only whether the string occurs, then the lines which report the matches can be substituted by simple early exist returning true.
+Reporting of the leafs below a certain node can be done by a Depth First Search reporting any leafs found.
+If one is interested in knowing only whether the string occurs, then the lines which report the matches can be substituted by early exits returning true.
 

### 2005-09-19 paper/space.tex
Win space and other small changes

@@
 \centering
-\includegraphics[height=.2\textheight]{figures/aaaa}
+\includegraphics[height=.3\textheight]{figures/aaaa}
 \caption{$a^n$: worst case space consumption}%
@@
 For a larger number of errors $k$, it is easy to see that this leads to $\mathcal{O}(n^{k+1})$ nodes.
-
 We can do better, however, by considering the expected case.
@@
 \begin{definition}[Maximum Path Lenght]
-$l$ is the length of the deepest node in the tree (which isn't a leaf)
+$l$ is the length of the deepest branching node in the tree
 
@@
 \begin{definition}[Suffix-Depth]
-For any node $\overline{w}$ its suffix-depth is the number of nodes which must be followed to reach the root.
+For any node $\overline{w}$ its suffix-depth is the number of nodes which must be followed to reach the root, which is $|w|$. It is obvious that $l$ is a maximum on the suffix-depth of nodes.
 \end{definition}
-
-For a node $\overline{w}$, its suffix-depth is $|w|$. It is obvious that $l$ is a maximum on the suffix-depth of nodes.
 
@@
 The algorithm for turning a k-error into a (k+1)-error dotted tree, can be looked at the following way~\footnote{We are here considering that the second \textit{foreach} loop of Figure~\ref{algo:construct-k} runs in a specific order which is in fact difficult to code for.
-However, as an analysis tool, it is a valid assumption.}: First it makes an almost exact copy of the k-error dotted tree and sets it as the root's error tree and it clears all the other error trees.
+However, as an analysis tool, it is a valid assumption.}: First it makes an almost exact copy of the k-error dotted tree and as the error tree at the root and it clears all the other error trees.
 Then it proceeds in stages, making a (possibly incomplete) copy of this tree spread amongst the nodes at suffix-depth 1.
-It proceeds to the nodes at suffix-depth 2 and on to increasing suffix-depths up to $l$.
+It proceeds to the other nodes in increasing suffix-depths up to $l$.
 At each such step, the number of nodes is increased by a maximum of $N_k$.
-Therefore, we start with $N_k$ nodes, make an almost full copy, and copy that at most $l$ times.
-We have $N_{k+1}=\mathcal{O}(N_k(l+1))$.
+Therefore, we start with $N_k$ nodes, make an almost full copy, and copy that at most $l$ times, $N_{k+1}=\mathcal{O}(N_k(l+1))$.
 Assuming $N_k=\mathcal{O}(nl^k)$ by induction we conclude $N_{k+1}=\mathcal{O}(nl^{k+1})$.

### 2005-09-19 paper/structure.tex
Win space and other small changes

@@
 \begin{figure}
-\includegraphics[width=\textwidth]{figures/mississippi-0}
+\centering
+\includegraphics[width=.75\textwidth]{figures/mississippi-0}
 \caption{Suffix Tree for the string \putstring{mississippi}}%
@@
 
-To extend the suffix tree, we will consider that a node in the suffix tree corresponds to several points in the string at the same time (those points where the string on the path to the node occurs).
-We now add to each node an extra edge which corresponds to moving one character forward in all these points.
+To extend the suffix tree, we consider that a node in the suffix tree corresponds to several points in the string simultaneously (those points where the string on the path to the node occurs).
+To each node we add an extra edge corresponding to moving one character forward in all these points.
 
@@
 For denoting characters we shall use letters from the beginning of the roman alphabet ($a$, $b$, $c$,\ldots) and for strings, we shall use letters from the end of the alphabet ($w$, $z$, \ldots).
+In what follows we assume that there are two symbols (\$ and $.$) which are not part of $\Sigma$.
 
@@
 %
-In what follows we will assume that there are two symbols (\$ and $.$) which are not part of $\Sigma$.
 
-In this paper, we will be interested in substitution distance (aka Hamming distance):
+In this paper, we are interested in the substitution distance (aka Hamming distance):
 
 \begin{definition}[Aproximate Match]
-We say that the string $s$ matches the string $t$ at position $p$ with $k$ errors if we can make $k$ substitutions in $s$ such that $s$ is a substring of $t$ at position $p$.
+We say that the string $s$ matches the string $t$ at position $p$ with $k$ errors if we can make $k$ substitutions in $s$ to obtain $s'$ is a substring of $t$ at position $p$.
 \end{definition}
@@
 Both Ukkonen's~\cite{ukkonen92constructing} and McCreight's~\cite{mccreight} algorithms are linear time algorithms for constructing a suffix tree which includes suffix links.
-The suffix links are shown in Figure~\ref{fig:mississippi-0} as dotted links.
+The suffix links are shown in Figure~\ref{fig:mississippi-0} as dashed arrows with the label in italic.
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{figures/mississippi-1.eps}
+\caption{1-error dotted tree for \putstring{mississippi}}%
+\label{fig:mississippi-1}
+\end{figure}
 
@@
 In a sense, one can say that being at node $\overline{issi}$ is being at positions 6 and~9 simultaneously.
-Note that for each node, the tree below it is the Patricia tree of the suffixes starting at the positions in its position set.
-
+For each node, the tree below it is the Patricia tree of the suffixes starting at the positions in its position set.
 In a leaf, the occurrence set is a singleton, and we label the leaf by its element.
@@
 
-The lemma is fairly obvious given that the position set of $\overline{aw}$ contains all the positions where $aw$ occurs which will be exactly one minus those positions where $w$ occurs following $a$.
+The lemma is fairly obvious given that the position set of $\overline{aw}$ contains all the positions where $aw$ occurs which are exactly those positions where $w$ occurs preceded by $a$.
 
@@
 
-\begin{figure} 
-\includegraphics[width=\textwidth]{figures/mississippi-1.eps}
-% mississippi-1.eps: 300dpi, width=6.08cm, height=1.63cm, bb=0 0 718 193
-\caption{1-error dotted tree for \putstring{mississippi}}%
-\label{fig:mississippi-1}
-\end{figure}
-
-The one-error dotted tree for \putstring{mississippi} is shown in Figure~\ref{fig:mississippi-1}. The nodes are connected to their error trees by thick diagonal links.
+The 1-error dotted tree for \putstring{mississippi} is shown in Figure~\ref{fig:mississippi-1}. The nodes are connected to their error trees by thick diagonal links.
 

### 2005-09-19 paper/conclusion.tex
A submission ready version of the paper

--- 2005-09-16 paper/conclusion.tex
@@
 
-Along the same lines, the structure's definition might be extended to structures such as the suffix-DAG presented in~\cite[7.7]{gusfield:algorithms}.
+Along the same lines, the structure's definition might be extended to structures such as the suffix-DAG presented in~\cite[\S\,7.7]{gusfield:algorithms}.
 An algorithm to efficiently construct it over these suffix-DAGs is another open problem.

### 2005-09-19 paper/construction.tex
A submission ready version of the paper

@@
 \subsection{One-Error Dotted Tree}
-\subsubsection{The Algorithm}
 
@@
 \begin{enumerate}
-\item It does not have the leaf labeled 1 in the original tree.
-\item All other leafs are labeled with value which is the original leaf's value minus one.
+\item It does not have the leaf labeled 1 in the original tree. Keeping this leaf would have resulted in the \putstring{.s\$} which does not occur in original string (it is one character too large).
+\item For any other leaf $\overline{w\$}$ which occurred at position $p$ in the string, we have a new leaf $\overline{.w\$}$ which occurs at position $p-1$ in the string.
+Therefore, we adjust the labels.
+
 \end{enumerate}
-
-It is easy to see why this is so.
-The leaf labeled 1 would have resulted in the pattern \putstring{.s\$} which does not occur in original string (it is one character too large).
-For any other leaf $\overline{w\$}$ which occurred at position $p$ in the string, we have a new leaf $\overline{.w\$}$ which occurs at position $p-1$ in the string.
 
@@
 \begin{enumerate}
-\item The leaf labeled 1 in the original error tree is not included
-\item A leaf labeled $p$ is included only if $s_{p-1}$ equals the label of the suffix link
-\item Leafs in the copy have a label which is the original label's value minus one
+\item The leaf labeled 1 in the original error tree is not included.
+\item Leafs in the copy have a label which is the original label's value minus one.
+\item A leaf labeled $p$ is included only if $s_{p-1}$ equals the label of the suffix link.
 \end{enumerate}
@@
 
-%\input{copy-subtree-algo}
-%
+\input{copy-subtree-algo}
+
 The only point to note is line~\ref{algoline:copy-subtree:merge}.
 Since we filter some leafs, without a merging procedure, it would be possible to have nodes which possess only one child.
-These need to be removed.
+These are removed by merging a child with its parent.
 
@@
 
-It is very simple to remove these nodes.
-As discussed above, suffix trees are normally implemented such that each node is actually just a pair of indices into the original string (exemplified in Figure~\ref{fig:mississippi-nodes}).
+As discussed above, suffix trees are implemented such that each node is actually just a pair of indices into the original string (examplified in Figure~\ref{fig:mississippi-nodes}).
 To merge a child with its parent, we just need to replace both by a node whose starting index is the starting index of the original child minus the length of the parent edge.
@@
 
-To add dot links at every node, we need to add the dot link at the root and then for every node call the procedure addDotLink.
-This is implemented as Algorithm~\ref{algo:construct-1}.
-The \textit{foreach} loop at line~\ref{algoline:construct-1:foreach} can be implemented using Depth First Search on the tree.
+Algorithm~~\ref{algo:construct-1} constructs the 1-error dotted tree.
+The \textit{foreach} loop at line~\ref{algoline:construct-1:foreach} can be implemented using Depth First Search.
 In fact, the procedure addDotLink only needs to be called for certain nodes (those which do not possess an incoming suffix link).
@@
 
-\subsubsection{Time cost}
+\subsubsection*{Time cost}
 
@@
 
+\input{copy-subtree-with-dot-link-algo}
+
 The above algorithm can be used to construct trees with any number of errors by iterating it.
@@
 \input{construct-k-algo}
-\input{copy-subtree-with-dot-link-algo}
 
-The code for the copy is Algorithm~\ref{algo:copy-subtree-with-dot-link} which differs from Algorithm~\ref{algo:copy-subtree} in that dot link are also followed while copying the tree.
+The code for the subtree copy is Algorithm~\ref{algo:copy-subtree-with-dot-link} which differs from Algorithm~\ref{algo:copy-subtree} in that dot link are also followed while copying the tree.
 

### 2005-09-19 paper/experimental.tex
A submission ready version of the paper

@@
 
-\subsection{Indexing Time and Size}
+\subsection{Index Size}
 
-We report the indexing time and size (as a count of the nodes in the structure) as a function of the number of character (by using larger and larger prefixes of the datasets).
-
-\begin{figure}%
+\begin{figure*}%
 \centering
 \includegraphics[width=.7\textwidth]{ratio-in=english-01-12.eps}%
-\caption{Ratios of k-error and (k+1)-error dotted trees}\label{fig:ratios}%
-\end{figure}
+\caption{Ratio of number of nodes in k-error and (k+1)-error dotted trees}\label{fig:ratios}%
+\end{figure*}
 
-To experimentally verify the average case, we show in Figures~\ref{fig:ratios} the ratios between the k-error and the (k+1)-error dotted trees.
-We can easily see that for all three data sets, the experimental values do resemble a logarithm as predicted.
+To experimentally verify the average case prediction, we show in Figures~\ref{fig:ratios} the ratios between the k-error and the (k+1)-error dotted trees, for english text (the other datasets show similiar behaviour).
+We can easily see that the experimental values do resemble a logarithm as predicted.
 
@@
 For searching, the full texts were first indexed and then online searches were performed on top of the structure.
-To generate input patterns, randomly chosen bits of the text were taken.
+To generate input patterns, randomly chosen snippets of the text were taken.
 These patterns were then altered by substituting a character at random.
@@
 
-\begin{figure}
+\begin{figure*}
 \centering
-\includegraphics[width=.7\textwidth]{search-vary-N.in=all.eps}
-\caption{Search time versus pattern size}\label{fig:search-vary-m}
-\end{figure}
+\includegraphics[width=\textwidth]{search-vary-N.in=all.eps}
+\caption{Search time versus text size}\label{fig:search-vary-N}
+\end{figure*}
 
-Figure~\ref{fig:search-vary-m} shows the search time for growing pattern lenghts.
+%Figure~\ref{fig:search-vary-m} shows the search time for growing pattern lenghts.
 It shows the case where the pattern exists and where it does not (an extra error is added).
@@
 The value presented is the number of character comparisons the algorithm makes.
+
+Figure~\ref{fig:search-vary-N} shows searching for a 15~character long sequences on a 2-error dotted tree.
+Presented are the number of character comparisons performed (averaged over 10\,000 such searches) while varying the text size.
+We see that after an initial small growth which is explained by the increasing density of the tree, the search time is roughly constant.

### 2005-09-19 paper/search.tex
A submission ready version of the paper

@@
-To search a string with at maximum k errors in a k-error dotted tree, we need to descend the tree.
+To search a string with at maximum k errors in a k-error dotted tree, we descend the tree.
 Whenever we find a node we follow both the dot link and the edge which matches the current position of the string we are searching.
-If we have a mismatch in the middle of an edge, we continue, decreasing the amount of errors permitted.
-
-Before introducing the pseudo-code, it is important to understand that suffix trees are normally implemented so that each node consists of only two indices into the original string which are the start and end position of its incoming edge.
-Figure~\ref{fig:mississippi-nodes} shows an example of this structure (for the now familiar \putstring{mississippi}).
-
-\begin{figure}
-\includegraphics[width=\textwidth]{figures/mississippi-nodes}
-\caption{Suffix Tree for the string \putstring{mississippi}}%
-\label{fig:mississippi-nodes}
-\end{figure}
-
+If we find a mismatch in the middle of an edge, we continue, decreasing the amount of errors permitted.
 When presenting the algorithm, we take the liberty of using $s+j$ where $s$ is a string and $j$ an integer to mean the suffix of $s$ starting at position $j$ (as happens in the C~Programming Language).
@@
 
-The search algorithm is presented in Figure~\ref{algo:search}.
-Reporting of the leafs below a certain node can be done by a Depth First Search reporting any leafs found.
+The process just described is Algorithm~\ref{algo:search}.
+Reporting of the leafs below a certain node can be done by a Depth First Search.
 If one is interested in knowing only whether the string occurs, then the lines which report the matches can be substituted by early exits returning true.
 
-This algorithm runs in time $\mathcal{O}(m^{k+1})$:
-
-The recursive call at line~\ref{algoline:search:followDot} means that a dot link was followed.
-There are at most $\mathcal{O}(m^k)$ places where a dot link can be followed while descending a string~\footnote{For $k=1$ each point in the string.
-For $k>1$, each path for $k-1$, can now have at most $m$ extra branches.}.
+This algorithm runs in time $\mathcal{O}(m^{k+1})$: The recursive call at line~\ref{algoline:search:followDot} corresponds to following a dot link.
+There are at most ${m \choose k}=\mathcal{O}(m^k)$ different paths which must be followed.
 Following one path in the tree takes $\mathcal{O}(m)$ time.

### 2005-09-19 paper/space.tex
A submission ready version of the paper

@@
 
-How much space does a k-error dotted tree take?
-Suffix Trees have a number of nodes proportional to the size of the string (which we will refer to as $n$).
-One-error dotted trees have more nodes.
-The worst case presented in Figure~\ref{fig:aaaa} is a string of the form $a^n = aaa\dots$ which generates a one-dotted tree taking $\mathcal{O}(n^2)$ nodes.
-For a larger number of errors $k$, it is easy to see that this leads to $\mathcal{O}(n^{k+1})$ nodes.
+Suffix trees have a number of nodes proportional to the size of the string (which we will refer to as $n$).
+Dotted trees have more nodes.
+The worst case is a string of the form $a^n = aaa\dots$ which generates a one-dotted tree taking $\mathcal{O}(n^2)$ nodes (illustrated in Figure~\ref{fig:aaaa}).
+For a larger number of errors $k$, this string leads to $\mathcal{O}(n^{k+1})$ nodes.
 We can do better, however, by considering the expected case.
@@
 
-We can now show that $N_k = \mathcal{O}(nl^k)$ by induction.
+We now show $N_k = \mathcal{O}(nl^k)$ using induction.
 It is a known fact that $N_0 = \mathcal{O}(n)$.
-The algorithm for turning a k-error into a (k+1)-error dotted tree, can be looked at the following way~\footnote{We are here considering that the second \textit{foreach} loop of Figure~\ref{algo:construct-k} runs in a specific order which is in fact difficult to code for.
+The algorithm for turning a k-error into a (k+1)-error dotted tree, can be looked at the following way~\footnote{We are here considering that the second \textit{foreach} loop of Algorithm~\ref{algo:construct-k} runs in a specific order which is in fact difficult to code for.
 However, as an analysis tool, it is a valid assumption.}: First it makes an almost exact copy of the k-error dotted tree and as the error tree at the root and it clears all the other error trees.

### 2005-09-19 paper/structure.tex
A submission ready version of the paper

@@
 \centering
-\includegraphics[width=.75\textwidth]{figures/mississippi-0}
-\caption{Suffix Tree for the string \putstring{mississippi}}%
+\includegraphics[width=.45\textwidth]{figures/mississippi-0}
+\caption{Suffix tree representation for the string \putstring{mississippi}}%
 \label{fig:mississippi-0}
@@
 
-Each node in a Patricia tree has a path leading to it which forms a string. If the node $n$ has the leading path $w$, we shall also refer to $n$ as $\overline{w}$one .
+Each node in a Patricia tree has a path leading to it which forms a string. If the node $n$ has the leading path $w$, we shall also refer to $n$ as $\overline{w}$.
 \end{definition}
@@
 
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{figures/mississippi-1.eps}
-\caption{1-error dotted tree for \putstring{mississippi}}%
-\label{fig:mississippi-1}
-\end{figure}
+\begin{figure*}
+\includegraphics[width=\textwidth]{figures/mississippi-nodes}
+\caption{Suffix Tree for the string \putstring{mississippi}}%
+\label{fig:mississippi-nodes}
+\end{figure*}
+
+It is important to understand that suffix trees are normally implemented so that each node consists of only two indices into the original string which are the start and end position of its incoming edge.
+Figure~\ref{fig:mississippi-nodes} illustrates this.
 
@@
 
+\begin{figure*}
+\centering
+\includegraphics[width=\textwidth]{figures/mississippi-1.eps}
+\caption{1-error dotted tree for \putstring{mississippi}}%
+\label{fig:mississippi-1}
+\end{figure*}
+
 The 1-error dotted tree for \putstring{mississippi} is shown in Figure~\ref{fig:mississippi-1}. The nodes are connected to their error trees by thick diagonal links.

### 2005-09-19 scripts/get-fileportion.py

@@ -4,6 +4,11 @@ import sys
 import random
 import os
 
+if len(argv) < 5:
+	print "usage:"
+	print sys.argv[0], " inputfile [ begin | . ] lenght errors [ repeats ]"
+	sys.exit(0)
+
 inputfile=argv[1]
 filelen=os.stat(inputfile)[6]
 
@@ -12,8 +17,14 @@ k=int(argv[4])
 
 f=open(inputfile,'r')
 repeats=1
+
 if len(argv) > 5:
 	repeats=int(argv[5])
+
+if lenght > filelen:
+	print "File %s is too small!" % inputfile
+	sys.exit(1)
+	
 
 def get_string(f,begin,lenght,k):
 	f.seek(begin)

### 2005-09-19 paper/construction.tex
First submission

@@
 \begin{enumerate}
-\item It does not have the leaf labeled 1 in the original tree. Keeping this leaf would have resulted in the \putstring{.s\$} which does not occur in original string (it is one character too large).
-\item For any other leaf $\overline{w\$}$ which occurred at position $p$ in the string, we have a new leaf $\overline{.w\$}$ which occurs at position $p-1$ in the string.
-Therefore, we adjust the labels.
+\item It does not have the leaf labeled 1 in the original tree.
+Keeping this leaf would have resulted in a leaf labeled \putstring{.s\$} which does not occur in original string (it is one character too long).
+\item For any other leaf $\overline{w\$}$ occuring at position $p$ in the string, we have a new leaf $\overline{.w\$}$ which occurs at position $p-1$ in the string. Therefore, we adjust the labels.
 
@@
 
-These conditions are an expression of Lemma~\ref{lemma:suffix-error} and a an extension of the conditions for the root. Both are implemented by Algorithm~\ref{algo:copy-subtree}.
+These conditions are an expression of Lemma~\ref{lemma:suffix-error} and an extension of the conditions for the root. Both are implemented by Algorithm~\ref{algo:copy-subtree}.
 
@@
 
-As discussed above, suffix trees are implemented such that each node is actually just a pair of indices into the original string (examplified in Figure~\ref{fig:mississippi-nodes}).
+As discussed above, suffix trees are implemented such that each node is actually just a pair of indices into the original string (exemplified in Figure~\ref{fig:mississippi-nodes}).
 To merge a child with its parent, we just need to replace both by a node whose starting index is the starting index of the original child minus the length of the parent edge.
@@
 The code for the subtree copy is Algorithm~\ref{algo:copy-subtree-with-dot-link} which differs from Algorithm~\ref{algo:copy-subtree} in that dot link are also followed while copying the tree.
-
 The final procedure for constructing k-error trees is Algorithm~\ref{algo:construct-k}.

### 2005-09-19 paper/experimental.tex
First submission

@@
 
-To experimentally verify the average case prediction, we show in Figures~\ref{fig:ratios} the ratios between the k-error and the (k+1)-error dotted trees, for english text (the other datasets show similiar behaviour).
+To experimentally verify the average case prediction, we show in Figures~\ref{fig:ratios} the ratios between the k-error and the (k+1)-error dotted trees, for English text (the other datasets show similar behaviour).
 We can easily see that the experimental values do resemble a logarithm as predicted.
@@
 
-For searching, the full texts were first indexed and then online searches were performed on top of the structure.
+For searching, the texts were first indexed and then online searches were performed on top of the structure.
 To generate input patterns, randomly chosen snippets of the text were taken.
-These patterns were then altered by substituting a character at random.
-To generate mismatches, one character too many was altered~\footnote{Since the characters used were not present in the original text, there is no risk of generating a string which \emph{does appear} in the text.}.
+These patterns were then altered by substituting a few characters at random.
+To generate mismatches, one character too many was altered\,\footnote{Since the characters used were not present in the original text, there is no risk of generating a string which \emph{does appear} in the text.}.
 The search algorithm performed an early exit, without reporting of occurrences (ie, it only reported whether the string exists in the text).

### 2005-09-19 paper/search.tex
First submission

@@
 To search a string with at maximum k errors in a k-error dotted tree, we descend the tree.
-Whenever we find a node we follow both the dot link and the edge which matches the current position of the string we are searching.
+Whenever we find a node, we follow both the dot link and the edge which matches the current position of the string we are searching for.
 If we find a mismatch in the middle of an edge, we continue, decreasing the amount of errors permitted.
@@
 
-This algorithm runs in time $\mathcal{O}(m^{k+1})$: The recursive call at line~\ref{algoline:search:followDot} corresponds to following a dot link.
-There are at most ${m \choose k}=\mathcal{O}(m^k)$ different paths which must be followed.
+This algorithm runs in time $\mathcal{O}(m^{k+1})$: There are at most $\sum_{i=0}{k}{m \choose i}=\mathcal{O}(m^k)$ different paths which must be followed.
 Following one path in the tree takes $\mathcal{O}(m)$ time.

### 2005-09-19 paper/space.tex
First submission

@@
 \begin{definition}[Maximum Path Lenght]
-$l$ is the length of the deepest branching node in the tree
+$l$ is the string depth of the deepest branching node in the tree:
 
@@
 It is a known fact that $N_0 = \mathcal{O}(n)$.
-The algorithm for turning a k-error into a (k+1)-error dotted tree, can be looked at the following way~\footnote{We are here considering that the second \textit{foreach} loop of Algorithm~\ref{algo:construct-k} runs in a specific order which is in fact difficult to code for.
+The algorithm for turning a k-error into a (k+1)-error dotted tree, can be looked at the following way\,\footnote{We are here considering that the second \textit{foreach} loop of Algorithm~\ref{algo:construct-k} runs in a specific order which is in fact difficult to code for.
 However, as an analysis tool, it is a valid assumption.}: First it makes an almost exact copy of the k-error dotted tree and as the error tree at the root and it clears all the other error trees.

### 2005-09-19 paper/structure.tex
First submission

@@
 We shall write $s_i$ for the $i$th element of the string.
-For the empty string, we write $\epsilon$.
+For the empty string, we write $\varepsilon$.
 
-The set of all strings is denoted by $\Sigma^*$ and $\Sigma^+=\Sigma^*-\{\epsilon\}$.
+The set of all strings is denoted by $\Sigma^*$ and $\Sigma^+=\Sigma^*-\{\varepsilon\}$.
 \end{definition}
@@
 
-\begin{definition}
+\begin{definition}[Concatenation]
 $wx$ will denote the string formed by all the characters of $w$ followed by all the characters of $x$ (the usual concatenation operation).
@@
 \begin{definition}[Aproximate Match]
-We say that the string $s$ matches the string $t$ at position $p$ with $k$ errors if we can make $k$ substitutions in $s$ to obtain $s'$ is a substring of $t$ at position $p$.
+We say that the string $s$ matches the string $t$ at position $p$ with $k$ errors if we can make $k$ substitutions in $s$ to obtain $s'$ which is a substring of $t$ at position $p$.
 \end{definition}
@@
 
-It is important to understand that suffix trees are normally implemented so that each node consists of only two indices into the original string which are the start and end position of its incoming edge.
+Suffix trees are normally implemented so that each node consists of only two indices into the original string which are the start and end position of its incoming edge.
 Figure~\ref{fig:mississippi-nodes} illustrates this.
@@
 \begin{lemma}[Position set at the suffix node]\label{lemma:suffix-error}
-Given two nodes $\overline{aw}$ and $\overline{w}$ (connected by a suffix link, labeled $a$), if one takes position set of $\overline{w}$, subtracts one from each element, resulting a superset of the position set of $\overline{w}$.
+Given two nodes $\overline{aw}$ and $\overline{w}$ (connected by a suffix link, labeled $a$), if one takes position set of $\overline{w}$, subtracts one from each element, one obtains a superset of the position set of $\overline{w}$.
 The items shared by both sets are those positions of the string which contain an $a$.
@@
 \begin{definition}[Error Tree]
-For any node $w$ its error tree is the Patricia tree formed by taking its position set, adding one to each element and forming the Patricia tree of the suffixes of the suffixes starting at those positions.
+For any node $w$, its error tree is the Patricia tree formed by taking its position set, adding one to each element and forming the Patricia tree of the suffixes of the suffixes starting at those positions.
 If the position set includes the end of the string, that element is removed.

### 2005-09-20 paper/conclusion.tex
First submission
Don't use [REF] as word.
SUBMITTED

--- 2005-09-16 paper/conclusion.tex
@@
 This structure allows matches in $\mathcal{O}(m^{k+1})$ time and can be constructed in $\mathcal{O}(kN|\Sigma|)$ time ($N$ being the actual number of nodes).
-The structure and the algorithms to construct it are very simple and should be easily implemented.
+The structure and the algorithms to construct it are very simple and are easily implemented.
 
@@
 
-Along the same lines, the structure's definition might be extended to structures such as the suffix-DAG presented in~\cite[7.7]{gusfield:algorithms}.
+Along the same lines, the structure's definition might be extended to structures such as the suffix-DAG presented by Gusfield~\cite[\S\,7.7]{gusfield:algorithms}.
 An algorithm to efficiently construct it over these suffix-DAGs is another open problem.

### 2005-09-20 paper/introduction.tex
Don't use [REF] as word.
SUBMITTED

@@
 Several algorithms have been proposed to solve this problem.
-\cite{navarro:guided} presents a survey of the problem and existing approaches to solving it at the time that work was written.
-Recently there have been some further work in the field, notably~\cite{maass:text} which also contains an updated overview of methods for solving the problem in addition to presenting one itself.
+Navarro presents a survey of the problem and existing approaches to solving it at the time that work was written~\cite{navarro:guided}.
+Recently there have been some further work in the field, notably the work of M.~Maa\ss{}~\cite{maass:text} which also contains an updated overview of methods for solving the problem in addition to presenting one itself.
 
@@
 Our work is related to efforts to index the text for approximate matching~\cite{maass:text,amir00oneerror}.
-The structure presented is superficially very similar to the one presented in~\cite{DBLP:journals/tcs/ChattarajP05} as an \textit{inexact suffix tree}, although that work is concerned with solving a different set of problems and the algorithms presented are fundamentally different.
+The structure presented is superficially very similar to the one presented by Chattaraj~\cite{DBLP:journals/tcs/ChattarajP05} as an \textit{inexact suffix tree}, although that work is concerned with solving a different set of problems and the algorithms presented are fundamentally different.

### 2005-11-21 paper/search.tex
Correct the formula

@@
 
-This algorithm runs in time $\mathcal{O}(m^{k+1})$: There are at most $\sum_{i=0}{k}{m \choose i}=\mathcal{O}(m^k)$ different paths which must be followed.
+This algorithm runs in time $\mathcal{O}(m^{k+1})$: There are at most $\sum_{i=0}^{k}{m \choose i}=\mathcal{O}(m^k)$ different paths which must be followed.
 Following one path in the tree takes $\mathcal{O}(m)$ time.

### 2006-07-04 paper/abstract.tex
First submission
edition distance

@@
 We consider that we are given a text $\mathcal{T}$ which undergoes some preprocessing to generate an index.
-We can later query this index to identify the places where a string occurs up to a certain number of allowed substitutions $k$.
+We can later query this index to identify the places where a string occurs up to a certain number of allowed errors~$k$ (edition distance).
 We present a structure for indexing which occupies space $\mathcal{O}(n\log^kn)$ in the average case, independent of alphabet size, $n$ being the text size.
-This structure support searching in $\mathcal{O}(m^{k+1})$ time, for patterns of size $m$, again independent of alphabet size.
+This structure can be used to report the existence of a match with $k$ errors in $\mathcal{O}(3^k m^{k+1})$ and to report the occurrences in $\mathcal{O}(3^k m^{k+1} + \mbox{\it ed})$ time, where $m$ is the length of the pattern and where {\it ed} the number of matching edit scripts.
+These bounds are independent of alphabet size.
 The construction of the structure has time bound by $\mathcal{O}(kN|\Sigma|)$, where $N$ is the number of nodes in the index and $|\Sigma|$ the alphabet size.

### 2006-07-04 paper/conclusion.tex
edition scripts

@@
 This complexity was analyzed theoretically and observed experimentally.
-This structure allows matches in $\mathcal{O}(m^{k+1})$ time and can be constructed in $\mathcal{O}(kN|\Sigma|)$ time ($N$ being the actual number of nodes).
+This structure reports the existence of a match in  $\mathcal{O}(3^k m^{k+1})$ and reports the positions where the matches occur in  $\mathcal{O}(3^k m^{k+1} + {\it ed})$ time.
+It can be constructed in $\mathcal{O}(kN|\Sigma|)$ time ($N$ being the actual number of nodes).
 The structure and the algorithms to construct it are very simple and are easily implemented.
-
-The biggest drawback of this structure is the amount of space it takes which might limit its applicability. Future research along these lines might address this issue (see below).
+The fact that the structure uses $\mathcal{O}({\it ed})$ time (instead of $\mathcal{O}({\it occur})$) to report the occurrences of a pattern may a disadvantage in some applications, where one occurrence may lead to many edit scripts.
+In other applications (e.g., searching in DNA strings for degenerated occurrences of long strings), this will not be a problem since each occurrence with $k$ errors will, in general, correspond to only one edit script.
 
@@
 
-In the example for the string \putstring{mississippi}, presented in Figure~\ref{fig:mississippi-1}, one can see that the tree below \putstring{s.i} and \putstring{ssi} are exactly the same.
+One drawback of the structure we present is the amount of space it takes which might limit its applicability.
+One possible direction for tackling this problem is the following remark: In the example for the string \putstring{mississippi}, presented in Figure~\ref{fig:mississippi-1}, one can see that the tree below \putstring{s.i} and \putstring{ssi} are exactly the same.
 Whether such occurrences are the basis for a significant space saving and what algorithms might exploit them is an open question.
+Another limitation that should be addressed in future work is related with the fact that the complexity for reporting occurrences depends on the number of edit scripts, and not on the number of occurrences.
 
 Along the same lines, the structure's definition might be extended to structures such as the suffix-DAG presented by Gusfield~\cite[\S\,7.7]{gusfield:algorithms}.
-An algorithm to efficiently construct it over these suffix-DAGs is another open problem.
+An algorithm to efficiently construct such a structure over these suffix-DAGs is another open problem.
 
 \subsection{Acknowledgments}
-
-We thank Lu\'{\i}s~Russo and Sara~Madeira for several productive discussions.
+%
+We thank Lu\'{\i}s~Russo and Sara~Madeira for several productive discussions. This work was support in part by project \textsc{posi/sri/47778/2002 BioGrid}.

### 2006-07-04 paper/construction.tex
Merge most recent changes

@@
 
-Constructing the tree is very simple.
-We start with a suffix tree which includes suffix links and add error trees to it.
-First, we construct the error tree for the root.
-This tree is almost a copy of the entire tree, except for two properties:
+Starting with a suffix tree which includes suffix links we add error trees to it. First, we construct the error tree for the root which is almost a copy of the entire tree, except for two properties:
 
@@
 Keeping this leaf would have resulted in a leaf labeled \putstring{.s\$} which does not occur in original string (it is one character too long).
-\item For any other leaf $\overline{w\$}$ occuring at position $p$ in the string, we have a new leaf $\overline{.w\$}$ which occurs at position $p-1$ in the string. Therefore, we adjust the labels.
-
+\item For any other leaf $\overline{w\$}$ occurring at position $p$ in the string, we have a new leaf $\overline{.w\$}$ which occurs at position $p-1$ in the string. Therefore, we adjust the labels.
 \end{enumerate}
@@
 \item The leaf labeled 1 in the original error tree is not included.
-\item Leafs in the copy have a label which is the original label's value minus one.
-\item A leaf labeled $p$ is included only if $s_{p-1}$ equals the label of the suffix link.
+\item Leafs in the copy have a label which is the original value minus one.
+\item A leaf labeled $p$ is included only if $s_{p-1} = a$ ($a$ being the label of the suffix link).
 \end{enumerate}
 
-These conditions are an expression of Lemma~\ref{lemma:suffix-error} and an extension of the conditions for the root. Both are implemented by Algorithm~\ref{algo:copy-subtree}.
+\input{copy-subtree-algo}
+\input{adddotlink-algo}%
 
-\input{copy-subtree-algo}
-
+These conditions are an expression of Lemma~\ref{lemma:suffix-error} and an extension of the conditions for the root.
+Both are implemented by Algorithm~\ref{algo:copy-subtree}.
 The only point to note is line~\ref{algoline:copy-subtree:merge}.
-Since we filter some leafs, without a merging procedure, it would be possible to have nodes which possess only one child.
+Since we filter some leaves, without a merging procedure, it would be possible to have nodes which possess only one child.
 These are removed by merging a child with its parent.
-
 \begin{figure}
@@
 \includegraphics[width=.4\textwidth]{merge}
-\caption{Node Merging demonstration}%
+\caption{Node Merging example}%
 \label{fig:merge}
@@
 
-As discussed above, suffix trees are implemented such that each node is actually just a pair of indices into the original string (exemplified in Figure~\ref{fig:mississippi-nodes}).
+As discussed above, suffix trees are implemented in a way such that each node stores the start and end indices of the substring on its incoming edge.
 To merge a child with its parent, we just need to replace both by a node whose starting index is the starting index of the original child minus the length of the parent edge.
-This process is illustrated in Figure~\ref{fig:merge}.
-
 The construction of the tree using either Ukonnen's or McCreight's algorithm assures that this operation is correct due to the way that the indices are created.
-
-\input{adddotlink-algo}
 
@@
 This leads to a recursive definition implemented as Algorithm~\ref{algo:addDotLink}.
+This procedure needs to be called for all nodes, which can be done using a simple depth first search on the tree.
 
-\input{construct-algo}
+\subsubsection*{Time Cost}
 
-Algorithm~~\ref{algo:construct-1} constructs the 1-error dotted tree.
-The \textit{foreach} loop at line~\ref{algoline:construct-1:foreach} can be implemented using Depth First Search.
-In fact, the procedure addDotLink only needs to be called for certain nodes (those which do not possess an incoming suffix link).
-However, there is no readily available way to access only these nodes and the procedure above does little extra work.
-
-\subsubsection*{Time cost}
-
-If the number of nodes in the final tree is $N$ and the alphabet $\Sigma$, then the above algorithm runs in time $\mathcal{O}(N|\Sigma|)$.
-
-The error tree at the root is created in time proportional to the number of nodes it contains.
-Every other error tree is constructed by looking at an existing one and copying it.
-Each of these copies need time proportional to the original error tree.
-Since there are at maximum $|\Sigma|$ incoming suffix links to a node, each error tree is \emph{looked at} a maximum of $|\Sigma|$ times.
+Copying a tree takes time proportional to the number of nodes it contains.
+The error tree at the root is a straightforward copy of the whole tree and every other tree is a copy of an existing one.
+Since each node can have at most $|\Sigma|$ incoming suffix links, each error tree is \emph{looked at\/} to be copied at most $|\Sigma|$ times.
 The sum of all these operations is therefore bounded by $|\Sigma|N$.
+Therefore, if the number of nodes in the final tree is $N$, then the above algorithm runs in time $\mathcal{O}(N|\Sigma|)$.
 
@@
 
-\input{copy-subtree-with-dot-link-algo}
-
+\input{construct-k-algo}
 The above algorithm can be used to construct trees with any number of errors by iterating it.
-To construct the (k+1)-error tree from the k-error tree, make an adjusted copy of the tree as above (adjusting leafs and filtering the leafs labeled one) and make this the new root error tree.
+To construct the $(k+1)$-error tree from the $k$-error tree, make an adjusted copy of the tree as above (adjusting leaves and filtering the leaves with label~1) and make this the new root error tree.
 Then, for every other node, remove the current error tree (it has a level too few).
-Finally, for every node except the root, construct its error tree as above (adjusting leafs, filtering the leafs labeled one or which do not correspond to positions in the suffix link's label).
+Finally, for every node except the root, construct its error tree as above (adjusting leaves, filtering the leaves with label~1 or which do not correspond to positions in the suffix link's label).
 These error trees will now have the correct strings.
-
-\input{construct-k-algo}
-
-The code for the subtree copy is Algorithm~\ref{algo:copy-subtree-with-dot-link} which differs from Algorithm~\ref{algo:copy-subtree} in that dot link are also followed while copying the tree.
 The final procedure for constructing k-error trees is Algorithm~\ref{algo:construct-k}.
 
-We note that the method of removing the existing error trees can be improved in practice (the asymptotic bound remains the same) the following way: While making a copy of the tree for construction of the error tree at the root, instead of following a dot link to make a copy, move the error tree to its destination and make the necessary adjustments in place (this move will probably be just an adjustment of pointer depending on the exact implementation).
-Using this method only leafs labeled one need to be removed (as well as any internal node which becomes unnecessary).
+We note that the method of removing the existing error trees can be improved in practice (the asymptotic bound remains the same) the following way: While making a copy of the tree for construction of the error tree at the root, instead of following a dot link to make a copy, move the error tree to its destination and make the necessary adjustments in place.% (this move will probably be just an adjustment of pointer depending on the exact implementation).
+Using this method only leaves labeled one need to be removed (as well as any internal node which becomes unnecessary).
 
 \begin{definition}
-We define $N_k$ to be the number of nodes of the k-error dotted tree. $N_0$ is the number of nodes of the suffix tree. If $k$ is known from context we write $N$ for $N_k$.
+We define $N_k$ to be the number of nodes of the $k$-error dotted tree. $N_0$ is the number of nodes of the suffix tree. We will use $N$ for $N_k$ if $k$ is known from context.
 \end{definition}

### 2006-07-04 paper/experimental.tex
Count steps
Merge most recent changes

@@
 
-The code was implemented in the C~Language, starting from an open-source implementation of suffix trees~\cite{ansicimplementation}. Three data sets were used:
-
-\begin{enumerate}
-\item The English translation of ``Democracy in America'' by Alexis de Tocqueville as natural language example
-\item The DNA of S. Cerevesiae (first chromosome)
-\item Randomly generated text by a uniform distribution
-\end{enumerate}
-
-The machine used for testing was an Intel Pentium~IV running at~2.40GHz with~4GB of main memory.
+The code was implemented in the \CC~Language~\footnote{Code is available upon request.}.
+Three data sets were used: a set of English texts; the DNA of S.
+Cerevesiae; and randomly generated text (uniform distribution).
+For testing we used an Intel Pentium~IV running at~2.40GHz with~4GB of main memory.
 
@@
 \centering
-\includegraphics[width=.7\textwidth]{ratio-in=english-01-12.eps}%
+\subfigure[DNA]{\includegraphics[height=3cm,width=.45\textwidth]{ratio-in=dna-01-12.eps}}%
+\subfigure[English]{\includegraphics[height=3cm,width=.45\textwidth]{ratio-in=english-01-12.eps}}\\%
+\subfigure[Random]{\includegraphics[height=3cm,width=.45\textwidth]{ratio-in=random-01-12.eps}}%
 \caption{Ratio of number of nodes in k-error and (k+1)-error dotted trees}\label{fig:ratios}%
 \end{figure*}
-
-To experimentally verify the average case prediction, we show in Figures~\ref{fig:ratios} the ratios between the k-error and the (k+1)-error dotted trees, for English text (the other datasets show similar behaviour).
+%
+To experimentally verify the average case prediction, we show in Figures~\ref{fig:ratios} the ratios between the $k$-error and the $(k+1)$-error dotted trees, for all datasets.
 We can easily see that the experimental values do resemble a logarithm as predicted.
-
+%
 \subsection{Search Time}
-
-For searching, the texts were first indexed and then online searches were performed on top of the structure.
-To generate input patterns, randomly chosen snippets of the text were taken.
-These patterns were then altered by substituting a few characters at random.
-To generate mismatches, one character too many was altered\,\footnote{Since the characters used were not present in the original text, there is no risk of generating a string which \emph{does appear} in the text.}.
+%
+\begin{figure*}
+\centering
+\subfigure[k=1]{\includegraphics[width=.75\textwidth,height=5cm]{search-vary-N.in=all.k=1.steps.eps}}\\
+\subfigure[k=2]{\includegraphics[width=.75\textwidth,height=5cm]{search-vary-N.in=all.k=2.steps.eps}\hspace{-1cm}}
+\caption{Search time versus text size}\label{fig:search-vary-N}
+\end{figure*}
+%
+For searching, the texts were first indexed and then searches were performed on top of the structure.
 The search algorithm performed an early exit, without reporting of occurrences (ie, it only reported whether the string exists in the text).
 Therefore, the number of occurrences had no influence on the search time.
-
-\begin{figure*}
-\centering
-\includegraphics[width=\textwidth]{search-vary-N.in=all.eps}
-\caption{Search time versus text size}\label{fig:search-vary-N}
-\end{figure*}
-
-%Figure~\ref{fig:search-vary-m} shows the search time for growing pattern lenghts.
+%
+%Figure~\ref{fig:search-vary-m} shows the search time for growing pattern lengths.
 It shows the case where the pattern exists and where it does not (an extra error is added).
@@
 
-Figure~\ref{fig:search-vary-N} shows searching for a 15~character long sequences on a 2-error dotted tree.
-Presented are the number of character comparisons performed (averaged over 10\,000 such searches) while varying the text size.
+Figure~\ref{fig:search-vary-N} shows searching for a 15~character long sequences on a $k$-error dotted tree.
+Presented are the number of steps in the algorithm (averaged over 10\,000 such searches) while varying the text size.
 We see that after an initial small growth which is explained by the increasing density of the tree, the search time is roughly constant.
+
+\subsection{Varying Alphabet Size}
+%
+\begin{figure*}
+\centering
+\subfigure[Nodes]{\includegraphics[width=.35\textwidth]{nodes-vary-E.eps}}
+%\subfigure[Time]{\includegraphics[width=.25\textwidth]{construction-time-vary-E.eps}}
+\subfigure[Ratio]{\label{subfig:nodes-vary-E-ratio}\includegraphics[width=.35\textwidth]{construction-time-ratio-vary-E.eps}}
+\caption{Construction time versus $|\Sigma|$}\label{fig:nodes-vary-E}
+\end{figure*}%
+%
+\begin{figure*}
+\centering
+\subfigure[Time]{\includegraphics[width=.35\textwidth]{search-time-vary-E.eps}}%
+\subfigure[Steps]{\includegraphics[width=.35\textwidth]{search-steps-vary-E.eps}}
+\caption{Searching versus $|\Sigma|$}\label{fig:search-vary-E}
+\end{figure*}%
+%
+We generated 100\,KB long strings using a uniform distribution.
+We varied the alphabet from binary through 100 and constructed a~2-error tree.
+As above, searches were for 15~character long strings, averaged over 10\,000 repetitions.
+
+As predicted, the time to construct one node in the tree (shown in Figure~\ref{subfig:nodes-vary-E-ratio}) grows in a roughly linear fashion.
+The search time (shown in Figure~\ref{fig:search-vary-E}) does not grow significantly with alphabet size.

### 2006-07-04 paper/introduction.tex
Merge most recent changes

@@
 Since their introduction~\cite{DBLP:conf/focs/Weiner73}, suffix trees have been one of the methods of choice for text indexing.
-Suffix trees can be used solve the following problem: given a text which has been preprocessed, and a pattern, efficiently find the places in the text where this pattern occurs.
+Suffix trees can be used to solve the following problem: given a text which has been preprocessed, and a pattern, efficiently find the places in the text where this pattern occurs.
 However, for many real-life problems this is too restrictive.
@@
 Several algorithms have been proposed to solve this problem.
-Navarro presents a survey of the problem and existing approaches to solving it at the time that work was written~\cite{navarro:guided}.
-Recently there have been some further work in the field, notably the work of M.~Maa\ss{}~\cite{maass:text} which also contains an updated overview of methods for solving the problem in addition to presenting one itself.
+Navarro {\it et al} present a survey of the problem and existing approaches to solving it at the time that work was written~\cite{navarro:guided}.
 
@@
 The structure presented is superficially very similar to the one presented by Chattaraj~\cite{DBLP:journals/tcs/ChattarajP05} as an \textit{inexact suffix tree}, although that work is concerned with solving a different set of problems and the algorithms presented are fundamentally different.
+
+Amongst competing solutions for this problem, the most relevant is the
+work of M.~Maa\ss{}~\cite{maass:text}. His solution takes space
+$\bigO(|\Sigma|^k n\log^k n)$ (average case)~\footnote{Maa{\ss}
+considers $|\Sigma|$ and $k$ as constant and presents $\bigO(n\log^k
+n)$ as a complexity result. However, such analysis ignores the
+potentially large impact of alphabet size.} for a search time
+$\bigO(m)$.
+
+This work is also related to the work of Cole {\it el
+al}~\cite{cole:dictionary} whose initial intuitions resemble
+ours. However, they make different time and space tradeoffs to achieve
+$\bigO(m+\log\log n+\mathit{occ})$ searching (Hamming distance) with
+$\bigO(n\frac{log^k n}{k!})$ space. They also propose an algorithm for
+searching using edit distance at a time cost of $\bigO(m+\frac{\log^k
+n \log\log n}{k!} +3^k \mathit{occ})$. We improve on that result,
+especially for edit distance, by totally removing the dependency
+on~$n$ while searching.
+
+Very recently, Navarro and Chavez have presented a completely new
+approach based on metric spaces~\cite{navarro:metric}. This approach
+achieves, on average, $\mathcal{O}(m^{1+\epsilon})$ search time. Their
+structure takes however space $\bigO(m^{1+\sqrt{2}+\epsilon}n)$. In
+order to obtain such good bounds it also requires $m$ to be known at
+construction time. They have presented a variation which trades time
+for space to achieve $\bigO(n)$ space, but at a cost in construction
+and search time.

### 2006-07-04 paper/search.tex
Merge most recent changes

--- 2005-09-19 paper/search.tex
@@
-To search a string with at maximum k errors in a k-error dotted tree, we descend the tree.
-Whenever we find a node, we follow both the dot link and the edge which matches the current position of the string we are searching for.
-If we find a mismatch in the middle of an edge, we continue, decreasing the amount of errors permitted.
+Given a pattern to search for, we descend the tree point by point, following the pattern character by character.
+We represent this by keeping a node and an offset from the start of its incoming link.
+In an edge, we consider that there is an implicit dot link which goes forward one character.
+At each point, we can take four possible actions:
+
+\begin{description}
+\item[match] Where we descend the tree according to the pattern (may not be possible).
+\item[substitution] We follow the dot link (possibly implicit), moving in the pattern.
+\item[insertion] We follow the (possibly implicit) dot link, while not moving in the pattern.
+\item[deletion] We advance in the pattern, while not moving in the tree.
+\end{description}
+
+\input{search-algo-edit}
+%
+Going down the tree can therefore be seen as a sequence of edition and matching operations: For example, for the string \putstring{ixsi}: \operation{match(i)}, \operation{substitution(x)}, \operation{match(s)}, \operation{match(i)}, where there are at most $k$ elements which are not \operation{match(c)}.
+There are at most $\sum_{i=1}^k{m \choose i}=\mathcal{O}(m^k)$ ways to combine $k$ edit operations into a string of size $m$.
+Since there are $3$ different operations (substitution, insertion, and deletion), we have at most $\mathcal{O}(3^k m^k)$ different edition sequences.
+Each sequence has at most $m+k=\bigO(m)$ (given that $k<m$) elements and therefore the total time to find matches is $\mathcal{O}(3^k m^{k+1})$.
+Once a match has been found in the tree, a Depth First Search can be used to report all the leaves below the node.
+This takes time proportional to the number of leaves, ie.\ proportional to the number of edit scripts which can be used to match the pattern to a substring of the text (which can be greater than the number of occurences~\footnote{As often happens, strings of the form $a^m$ serve as examples of pathological behaviour as they can match any position of a string of form $a^n$ in a large number of ways.}).
+The total search time is therefore $\bigO(3^k m^{k+1} + \mbox{\it ed})$.
+
+The code needs to distinguish whether we are at the end of a branch (ie.\ at a node) or in the middle of one.
 When presenting the algorithm, we take the liberty of using $s+j$ where $s$ is a string and $j$ an integer to mean the suffix of $s$ starting at position $j$ (as happens in the C~Programming Language).
-
-\input{search-algo}
-
 The process just described is Algorithm~\ref{algo:search}.
-Reporting of the leafs below a certain node can be done by a Depth First Search.
-If one is interested in knowing only whether the string occurs, then the lines which report the matches can be substituted by early exits returning true.
-
-This algorithm runs in time $\mathcal{O}(m^{k+1})$: There are at most $\sum_{i=0}{k}{m \choose i}=\mathcal{O}(m^k)$ different paths which must be followed.
-Following one path in the tree takes $\mathcal{O}(m)$ time.
-Therefore, the total time is $\mathcal{O}(m^{k+1})$ without the reporting of occurrences (which can be done in time linear to the number of occurrences).

### 2006-07-04 paper/space.tex
Merge most recent changes

--- 2005-09-19 paper/space.tex
@@
 For a larger number of errors $k$, this string leads to $\mathcal{O}(n^{k+1})$ nodes.
-We can do better, however, by considering the expected case.
+An analysis of the average case, however, leads to a much better result.
 
-\begin{definition}[Maximum Path Lenght]
-$l$ is the length of the deepest branching node in the tree
+\begin{definition}[Maximum Path Length, Suffix Depth]
+$l$ is the string depth of the deepest branching node in the tree.
+%
+%\[ l = \max_{\overline{w} \in \mathcal{T}} |w| \mbox{ and $\overline{w}$ isn't a leaf} \]
+%\end{definition}
 
-\[ l = \max_{\overline{w} \in \mathcal{T}} |w| \mbox{ and $\overline{w}$ isn't a leaf} \]
-\end{definition}
-
-\begin{definition}[Suffix-Depth]
+%\begin{definition}[Suffix-Depth]
 For any node $\overline{w}$ its suffix-depth is the number of nodes which must be followed to reach the root, which is $|w|$. It is obvious that $l$ is a maximum on the suffix-depth of nodes.
@@
 It is a known fact that $N_0 = \mathcal{O}(n)$.
-The algorithm for turning a k-error into a (k+1)-error dotted tree, can be looked at the following way~\footnote{We are here considering that the second \textit{foreach} loop of Algorithm~\ref{algo:construct-k} runs in a specific order which is in fact difficult to code for.
-However, as an analysis tool, it is a valid assumption.}: First it makes an almost exact copy of the k-error dotted tree and as the error tree at the root and it clears all the other error trees.
+The algorithm for turning a $k$-error into a $(k+1)$-error dotted tree, can be looked at the following way\,\footnote{We are here considering that the second \textit{foreach} loop of Algorithm~\ref{algo:construct-k} runs in a specific order which is in fact difficult to code for.
+However, as an analysis tool, it is a valid assumption.}: First it makes an almost exact copy of the k-error dotted tree as the error tree at the root and it clears all the other error trees.
 Then it proceeds in stages, making a (possibly incomplete) copy of this tree spread amongst the nodes at suffix-depth 1.
@@
 
-So far, we have achieved little since in the worst case $l=n-1$ as shown above.
+So far, we have achieved little since in the worst case $l=n-1$ ($s=a^n$).
 However, the expected case is $l=\mathcal{O}(\log n)$~\cite{apostolico92selfalignments,szpankowski:unexpected} and we have $N_k=\mathcal{O}(n\log^k n)$.

### 2006-07-04 paper/structure.tex
Merge most recent changes

@@
-\subsection{Intuition}
-
-\begin{figure}
+\begin{figure*}
 \centering
-\includegraphics[width=.45\textwidth]{figures/mississippi-0}
+\includegraphics[width=.6\textwidth]{figures/mississippi-0}
 \caption{Suffix tree representation for the string \putstring{mississippi}}%
 \label{fig:mississippi-0}
-\end{figure}
-
+\end{figure*}
+\subsection{Intuition}
 Suffix Trees have become a well known structure since their introduction in 1973~\cite{DBLP:conf/focs/Weiner73}.
@@
 For denoting characters we shall use letters from the beginning of the roman alphabet ($a$, $b$, $c$,\ldots) and for strings, we shall use letters from the end of the alphabet ($w$, $z$, \ldots).
-In what follows we assume that there are two symbols (\$ and $.$) which are not part of $\Sigma$.
+In what follows we assume that there are two special symbols (\$ and $.$) which are not part of $\Sigma$.
 
 \begin{definition}[Concatenation]
-$wx$ will denote the string formed by all the characters of $w$ followed by all the characters of $x$ (the usual concatenation operation).
-$aw$ ($wa$) shall denote the string formed by the character $a$ before (after) the characters of $w$.
+$wx$ or $aw$ will denote the usual concatenation operation.
 
@@
 \begin{definition}[Patricia tree]
-$T$ is a Patricia tree if $T$ is a rooted tree with edge labels from $\Sigma^+$. For each $a \in \Sigma$ and every node $n$ in $T$, there is at most one edge leaving $n$ whose label is $aw$.
-
-Each node in a Patricia tree has a path leading to it which forms a string. If the node $n$ has the leading path $w$, we shall also refer to $n$ as $\overline{w}$.
+$T$ is a Patricia tree if $T$ is a rooted tree with edge labels from $\Sigma^+$.
+For each $a \in \Sigma$ and every node $n$ in $T$, there is at most one edge leaving $n$ whose label starts with $a$.
+Each node in a Patricia tree has a path leading to it which forms a string.
+If the node $n$ has the leading path $w$, we shall also refer to $n$ as $\overline{w}$.
 \end{definition}
@@
 %
-
-In this paper, we are interested in the substitution distance (aka Hamming distance):
+In this work we are interested in approximate matching using edition (or Levenshtein) distance:
 
 \begin{definition}[Aproximate Match]
-We say that the string $s$ matches the string $t$ at position $p$ with $k$ errors if we can make $k$ substitutions in $s$ to obtain $s'$ which is a substring of $t$ at position $p$.
+We say that the string $s$ matches the string $t$ at position $p$ with $k$ errors if we can make $k$ modifications in $s$ to obtain $s'$ which is a substring of $t$ at position $p$.
+A modification is either deletion, insertion or substitution of one character.
 \end{definition}
 
-\begin{definition}[Pattern]
-A pattern is a string over the extended alphabet $\Sigma\cup\{.\}$.
-We say that a pattern $s$ matches a string $t$ at position $p$, if for each dot in the pattern we can replace it by a character in the original alphabet to obtain $s'$ which is a substring of $t$ at position $p$.
+\begin{definition}[Suffix Tree, Suffix Link]
+A \emph{suffix tree} for a string $S$ is a Patricia tree whose leaf nodes (those without children) have paths corresponding to all suffixes of the string $S\$ $.
+A suffix link in a suffix tree is a link from the node $\overline{aw}$ to the node $\overline{w}$.
+This link has the label $a$.
 \end{definition}
 
-\begin{definition}[Suffix Tree]
-A \emph{suffix tree} for a string $S$ is a Patricia tree whose leaf nodes (those without children) have paths corresponding to all suffixes of the string $S\$ $.
-\end{definition}
-
-\begin{definition}[Suffix Link]
-A suffix link in a suffix tree is a link from the node $\overline{aw}$ to the node $\overline{w}$. This node has the label $a$.
-\end{definition}
-
-In a suffix tree, it is possible to define a suffix link for each internal node.
+In a suffix tree, all internal nodes have a well defined suffix link.
 Both Ukkonen's~\cite{ukkonen92constructing} and McCreight's~\cite{mccreight} algorithms are linear time algorithms for constructing a suffix tree which includes suffix links.
@@
 
-\begin{figure*}
-\includegraphics[width=\textwidth]{figures/mississippi-nodes}
-\caption{Suffix Tree for the string \putstring{mississippi}}%
-\label{fig:mississippi-nodes}
-\end{figure*}
+Suffix trees are normally implemented by associating with each node two indices into the original string which are the start and end position of its incoming edge.
 
-Suffix trees are normally implemented so that each node consists of only two indices into the original string which are the start and end position of its incoming edge.
-Figure~\ref{fig:mississippi-nodes} illustrates this.
-
-\begin{definition}[Occurence Set]
-Given a node $\overline{w}$ in a suffix tree, we call its \emph{occurrence set} the set of indexes in the original string where the string $w$ occurs.
+\begin{definition}[Point in a suffix tree]
+A \emph{point} in a suffix tree is either a node in the tree or a point in the middle of an edge.
 \end{definition}
 
-\begin{definition}[Position Set]
+\begin{definition}[Occurrence Set, Position Set]
+Given a node $\overline{w}$ in a suffix tree, we call its \emph{occurrence set} the set of indexes in the original string where the string $w$ occurs.
+
 Given a node $\overline{w}$ in a suffix tree, its \emph{position set} is the set formed by taking its occurrence set and adding the length of $w$ to each element.
@@
 \begin{definition}[Error Tree]
-For any node $w$, its error tree is the Patricia tree formed by taking its position set, adding one to each element and forming the Patricia tree of the suffixes of the suffixes starting at those positions.
+For any node $w$, its error tree is the Patricia tree formed by taking its position set, adding one to each element and forming the Patricia tree of the suffixes starting at those positions.
 If the position set includes the end of the string, that element is removed.
 
-The leafs are labeled by the position of the string in which their paths occur minus $|w| + 1$.
+The leaves are labeled by the position of the string in which their paths occur minus $|w| + 1$.
 \end{definition}
@@
 \begin{definition}[1-error dotted Tree]
-We define a \emph{1-error dotted tree} as the tree which is formed by adding to each node in a suffix tree, a new edge labeled by \putstring{.} which points to its error tree.
-The edge labeled \putstring{.} shall be called a \emph{dot link}.
+We define a \emph{1-error dotted tree} as the tree which is formed by adding to each node in a suffix tree, a new edge labeled by \putstring{$\cdot$} which points to its error tree.
+The edge labeled \putstring{$\cdot$} shall be called a \emph{dot link}.
 \end{definition}
@@
 \centering
-\includegraphics[width=\textwidth]{figures/mississippi-1.eps}
+\includegraphics[width=\textwidth]{figures/mississippi-1.eps}%
 \caption{1-error dotted tree for \putstring{mississippi}}%
-\label{fig:mississippi-1}
-\end{figure*}
-
+\label{fig:mississippi-1}%
+\end{figure*}%
 The 1-error dotted tree for \putstring{mississippi} is shown in Figure~\ref{fig:mississippi-1}. The nodes are connected to their error trees by thick diagonal links.
@@
 The paths in the dotted tree are paths in the extended alphabet $\Sigma\cup\{.,\$\}$.
-The notions of \emph{occurrence set}, \emph{position set} and \emph{error tree} are valid for all nodes in a dotted tree considering that a node path may contain dot elements which match any other character.
+The notions of \emph{occurrence set}, \emph{position set} and \emph{error tree} are valid for all nodes in a dotted tree.
 
 \begin{definition}[k-error dotted tree]
-We define a k-error dotted tree as the tree obtained by adding error trees to each node in the (k-1)-error dotted tree which does not already contain one.
+We define a $k$-error dotted tree as the tree obtained by adding error trees to each node in the $(k-1)$-error dotted tree which does not already contain one.
 \end{definition}

### 2006-07-04 paper/space.tex
slight corrections

@@
 \begin{definition}[Maximum Path Length, Suffix Depth]
-$l$ is the string depth of the deepest branching node in the tree.
-%
-%\[ l = \max_{\overline{w} \in \mathcal{T}} |w| \mbox{ and $\overline{w}$ isn't a leaf} \]
-%\end{definition}
+Let $l$ be the string depth of the deepest branching node in the tree~\footnote{For a node \Node, its string depth is $|w|$.}.
 
-%\begin{definition}[Suffix-Depth]
 For any node $\overline{w}$ its suffix-depth is the number of nodes which must be followed to reach the root, which is $|w|$. It is obvious that $l$ is a maximum on the suffix-depth of nodes.

### 2006-07-04 paper/structure.tex
slight corrections

@@
 If the node $n$ has the leading path $w$, we shall also refer to $n$ as $\overline{w}$.
+
+A \emph{compact} Patricia tree is one with no nodes with just one child.
 \end{definition}
@@
 \begin{definition}[Suffix Tree, Suffix Link]
-A \emph{suffix tree} for a string $S$ is a Patricia tree whose leaf nodes (those without children) have paths corresponding to all suffixes of the string $S\$ $.
+A \emph{suffix tree} for a string $S$ is a compact Patricia tree whose leaf nodes (those without children) have paths corresponding to all suffixes of the string $S\$ $.
 A suffix link in a suffix tree is a link from the node $\overline{aw}$ to the node $\overline{w}$.
@@
 Suffix trees are normally implemented by associating with each node two indices into the original string which are the start and end position of its incoming edge.
+This allows the structure to be kept in $\bigO(n)$ space.
 
